---
title: "Final_data_clean"
author: "Group2"
date: "2024-08-21"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
table-of-contents: true 
number-sections: true 
embed-resources: true 
editor_options: 
  markdown: 
    wrap: 72
---


# Library
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)      
library(readxl)    
library(janitor)   
library(stringr)   
library(tidyr)     
library(dplyr)
library(knitr)
library(ggplot2)
library(factoextra)
library(gt)
library(naniar)
library(ggcorrplot)
library(corrplot)
library(GGally)
library(randomForest)
library(pdp)
library(e1071)  
library(dplyr) 
library(caret)
library(plotly)
library(stargazer)
library(broom)
library(knitr)
library(kableExtra)
```
# Data Import
```{r data import}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
raw_bsp <- read.csv(here("ZIPallFiles","inp13bsp.csv"), header=TRUE)
raw_bcn <- read.csv(here("ZIPallFiles","inp13bcn.csv"), header=TRUE)
raw_bhh <- read.csv(here("ZIPallFiles","inp13bhh.csv"), header=TRUE)
```
# Rename and Created
```{r rename+Identity}
bp_bb <- inner_join(raw_bp, raw_bb, by = "ABSPID") %>%
  rename(HypertensionStatus = HYPBC,
         BodyMass = SABDYMS,
         SocialStatus = SF2SA1QN,
         PersonID = ABSPID,
         Age = AGEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2) %>%
   mutate(Identity = factor(0))

bsp_bhh<- left_join(raw_bsp, raw_bhh,by = c("ABSHID")) %>%
  rename(BodyMass = SABDYMS,
         SocialStatus = SF2SA1DB,
         HouseID = ABSHID,
         Age = AGEEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2)%>%
   mutate(Identity = factor(1))

bcn <- raw_bcn %>%
  rename(HouseID = ABSHID,
         MedicalCondition = ICD10ME)
```
# First Filter
```{r filter 1}
bp_bb <- bp_bb %>%
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    HypertensionStatus == 5,
    BodyMass != 4,
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))

bsp_bhh <- bsp_bhh %>% 
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    BodyMass != 4,
    !HouseID %in% bcn$HouseID[bcn$MedicalCondition %in% c(7, 20)],
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))
```
# Merge
```{r merge}
raw_data <- bind_rows(bp_bb,bsp_bhh)
```
# Variable Selection
```{r variable selection}
selected_data <- raw_data %>%
select(
    PersonID, HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, Identity,
    AlcoholPercentage_Day1, AlcoholPercentage_Day2,
    Potassium_Day1, Potassium_Day2, 
    Sodium_Day1, Sodium_Day2,
    FiberEnergy_Day1, FiberEnergy_Day2, 
    CarbonhydrateEnergy_Day1, CarbonhydrateEnergy_Day2
  )
```
# Type Check 1
```{r variable check before}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 1: Variable types table"))
```
# Na values + New Variables + Type Conversion
```{r navalue_new_typeconversion}
selected_data <- selected_data %>%
  mutate(
    across(c(Gender, SocialStatus, SmokeStatus), as.factor),
    BMI = na_if(BMI, 98) %>% na_if(99),
    Systolic = na_if(Systolic, 0) %>% na_if(998) %>% na_if(999),
    Diastolic = na_if(Diastolic, 0) %>% na_if(998) %>% na_if(999),
    Potassium = (Potassium_Day1 + Potassium_Day2) / 2, 
    Sodium = (Sodium_Day1 + Sodium_Day2) / 2,
    FiberEnergy = (FiberEnergy_Day1 + FiberEnergy_Day2)/2,
    CarbonhydrateEnergy = (CarbonhydrateEnergy_Day1 + CarbonhydrateEnergy_Day2) / 2,
    AlcoholPercentage = (AlcoholPercentage_Day1 + AlcoholPercentage_Day2) /2,
    PotassiumSodiumRatio = Sodium / Potassium
    ) %>% 
  select(PersonID,HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, AlcoholPercentage, PotassiumSodiumRatio, FiberEnergy, CarbonhydrateEnergy, Identity)
```
# Type Check 2
```{r variable check after}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 2: Variable types table after correction"))
```
# Duplicate Values
```{r duplicate values}
selected_data %>% 
  select(-PersonID) %>% 
  distinct() %>% 
  dim() %>% 
  tibble::tibble(Dimension = c("Rows", "Columns"), Count = .) %>% 
  gt() %>%
  tab_header(title = "Dimensions of Unique Tibble") %>%
  tab_caption(caption = md("Table 3: Dimensions of Unique Data Table"))
```

# Categrocial Distribution
```{r categrocial distribution bar plot}
plot_categorical_distribution <- function(data) {
  categorical_vars <- data %>%
    select(where(is.factor) | where(is.character)) %>%
    select(-PersonID, -HouseID)
  
  for (var in names(categorical_vars)) {
    non_na_data <- data %>%
      filter(!is.na(.data[[var]]))
    
    if (nrow(non_na_data) == 0) next
    
    p <- ggplot(non_na_data, aes(x = .data[[var]])) +
      geom_bar(aes(y = (..count..) / sum(..count..), fill = .data[[var]]), color = "black") +
      scale_fill_brewer(palette = "Set3") + 
      labs(title = paste("Proportion of Categories in", var), 
           x = "Category", 
           y = "Proportion") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  
    
    print(p)
  }
}

plot_categorical_distribution(selected_data) 
```
# Merge Variable
```{r merge variable}
selected_data <- selected_data %>% 
  mutate(SmokeStatus = recode_factor(SmokeStatus, "2" = "1", "3" = "1","4" = "2","5" = "3"))
```

# Outliers and New Variables
```{r variable creation and outliers}
normal_ranges <- list(Systolic = c(90, 180),
                      Diastolic = c(60, 120),
                      BMI = c(16, 47.5), 
                      PotassiumSodiumRatio = c(0, 4), 
                      AlcoholPercentage = c(0, 85), 
                      CarbonhydrateEnergy = c(0, 100), 
                      FiberEnergy = c(0, 100))

selected_data <- selected_data %>%
  filter(
    (is.na(Systolic) | (Systolic >= normal_ranges$Systolic[1] & Systolic <= normal_ranges$Systolic[2])),
    (is.na(Diastolic) | (Diastolic >= normal_ranges$Diastolic[1] & Diastolic <= normal_ranges$Diastolic[2])),
    (is.na(BMI) | (BMI >= normal_ranges$BMI[1] & BMI <= normal_ranges$BMI[2])),
    (is.na(AlcoholPercentage) | (AlcoholPercentage >= normal_ranges$AlcoholPercentage[1] & AlcoholPercentage <= normal_ranges$AlcoholPercentage[2])),
    (is.na(CarbonhydrateEnergy) | (CarbonhydrateEnergy >= normal_ranges$CarbonhydrateEnergy[1] & CarbonhydrateEnergy <= normal_ranges$CarbonhydrateEnergy[2])),
    (is.na(FiberEnergy) | (FiberEnergy >= normal_ranges$FiberEnergy[1] & FiberEnergy <= normal_ranges$FiberEnergy[2])),
    (is.na(PotassiumSodiumRatio) | (PotassiumSodiumRatio >= normal_ranges$PotassiumSodiumRatio[1] & PotassiumSodiumRatio <= normal_ranges$PotassiumSodiumRatio[2]))
  ) %>%
  mutate(
    Hypertension = as.factor(case_when(
      Systolic >= 120 | Diastolic >= 80 ~ 1,
      TRUE ~ 0
    ))
  )
 # select(-Sodium, -Potassium)
```

# Missing Values
```{r missing value, warning = FALSE}
selected_data %>% 
  select(-PersonID, -HouseID) %>% 
  miss_var_summary() %>% 
  gt() %>%
  tab_header(title = "Missing Value") %>%
  tab_caption(caption = md("Table 4: Missing Values in Selected Data")) %>%
  cols_label(variable = "Variable Name", n_miss = "Missing Count", pct_miss = "Percentage Missing") %>%
  fmt_number(columns = everything(), decimals = 2)
```

```{r heat map}
corr_matrix <- selected_data%>%
  select(where(is.numeric)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  cor(use = "complete.obs")

ggcorrplot(corr_matrix, lab = TRUE, 
           title = "Correlation Heatmap for ALL", 
           colors = c("blue", "white", "red"),
           tl.cex = 9, lab_size = 4) + theme(legend.position = "none")
```

# Regression Imputation
```{r regression imputation, fig.height = 5, fig.width= 11}
selected_data$BMI[is.na(selected_data$BMI)] <- lm(BMI ~ Gender + SocialStatus + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$BMI), ])

selected_data$Systolic[is.na(selected_data$Systolic)] <- lm(Systolic ~ Gender + SocialStatus + Age + BMI + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$Systolic), ])

selected_data$Diastolic[is.na(selected_data$Diastolic)] <- lm(Diastolic ~ Gender + SocialStatus + Age + BMI + Systolic + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>%
  predict(newdata = selected_data[is.na(selected_data$Diastolic), ])

vis_miss(selected_data %>% 
           select(-PersonID, -HouseID))
```
# Normalization
```{r normalization}
#numeric_cols <- sapply(selected_data, is.numeric)  
#normalized_data <- selected_data
#normalized_data[numeric_cols] <- scale(selected_data[numeric_cols])



numeric_data <- selected_data %>% select_if(is.numeric)

preprocess_params <- preProcess(numeric_data, method = c("center", "scale"))

scaled_numeric_data <- predict(preprocess_params, numeric_data)

normalized_data <- bind_cols(scaled_numeric_data, selected_data %>% select_if(Negate(is.numeric)))
```
# Multicollinearity
```{r multicollinearity, fig.height=10, fig.width=10}
selected_data %>% 
  select(where(is.numeric), -Systolic, -Diastolic) %>% 
  ggscatmat()
```

# Logistic Regression
```{r logistic regression}
lr <- glm(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
           data = normalized_data, 
           family = binomial())
summary(lr)
```

# Random Forest
```{r random forest}
set.seed(3888)

rf <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = normalized_data, ntree = 500)
```

# Support Vector Machine
```{r svm}
final_data <- cbind(as.data.frame(predict(dummyVars(~ . - Hypertension - PersonID - HouseID - Systolic - Diastolic, data = normalized_data), 
                                          newdata = normalized_data)), 
                    Hypertension = normalized_data$Hypertension)

set.seed(3888)

svm <- svm(Hypertension ~ ., 
           data = final_data,  
           kernel = "radial", cost = 1, gamma = 0.1)  
```

# Model comparison (5-fold cross validation)
```{r }
set.seed(3888)
k_folds <- 5
n_repeats <- 3

logistic_accuracy <- numeric(k_folds * n_repeats)
logistic_f1 <- numeric(k_folds * n_repeats)
logistic_sensitivity <- numeric(k_folds * n_repeats)
logistic_specificity <- numeric(k_folds * n_repeats)
logistic_precision <- numeric(k_folds * n_repeats)

rf_accuracy <- numeric(k_folds * n_repeats)
rf_f1 <- numeric(k_folds * n_repeats)
rf_sensitivity <- numeric(k_folds * n_repeats)
rf_specificity <- numeric(k_folds * n_repeats)
rf_precision <- numeric(k_folds * n_repeats)

svm_accuracy <- numeric(k_folds * n_repeats)
svm_f1 <- numeric(k_folds * n_repeats)
svm_sensitivity <- numeric(k_folds * n_repeats)
svm_specificity <- numeric(k_folds * n_repeats)
svm_precision <- numeric(k_folds * n_repeats)

ensure_factor_levels <- function(predictions, reference) {
  factor(predictions, levels = levels(reference))
}

for (repeat_idx in 1:n_repeats) {
  folds <- createFolds(normalized_data$Hypertension, k = k_folds, list = TRUE, returnTrain = FALSE)
  
  for (i in 1:k_folds) {
    fold_idx <- (repeat_idx - 1) * k_folds + i

    test_indices <- folds[[i]]
    test_data <- normalized_data[test_indices, ]
    train_data <- normalized_data[-test_indices, ]

    log_model <- glm(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                     data = train_data, family = binomial())
    log_predictions <- predict(log_model, test_data, type = "response")
    log_class <- ifelse(log_predictions > 0.5, 1, 0)
    log_class <- ensure_factor_levels(log_class, test_data$Hypertension)

    log_conf_matrix <- confusionMatrix(log_class, test_data$Hypertension, positive = "1")
    logistic_accuracy[fold_idx] <- log_conf_matrix$overall['Accuracy']
    logistic_f1[fold_idx] <- log_conf_matrix$byClass['F1']
    logistic_sensitivity[fold_idx] <- log_conf_matrix$byClass['Sensitivity']
    logistic_specificity[fold_idx] <- log_conf_matrix$byClass['Specificity']
    logistic_precision[fold_idx] <- log_conf_matrix$byClass['Precision']

    rf_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = train_data, ntree = 500)
    
    rf_predictions <- predict(rf_model, test_data)
    rf_predictions <- ensure_factor_levels(rf_predictions, test_data$Hypertension)

    rf_conf_matrix <- confusionMatrix(rf_predictions, test_data$Hypertension, positive = "1")
    rf_accuracy[fold_idx] <- rf_conf_matrix$overall['Accuracy']
    rf_f1[fold_idx] <- rf_conf_matrix$byClass['F1']
    rf_sensitivity[fold_idx] <- rf_conf_matrix$byClass['Sensitivity']
    rf_specificity[fold_idx] <- rf_conf_matrix$byClass['Specificity']
    rf_precision[fold_idx] <- rf_conf_matrix$byClass['Precision']

    dummy_model <- dummyVars(~ . - Hypertension - PersonID - HouseID - Systolic - Diastolic, data = train_data)
    encoded_train_data <- as.data.frame(predict(dummy_model, newdata = train_data))
    encoded_test_data <- as.data.frame(predict(dummy_model, newdata = test_data))
    encoded_train_data$Hypertension <- train_data$Hypertension
    encoded_test_data$Hypertension <- test_data$Hypertension

    svm_model <- svm(Hypertension ~ ., data = encoded_train_data, kernel = "radial", cost = 1, gamma = 0.1)
    svm_predictions <- predict(svm_model, encoded_test_data)
    svm_predictions <- ensure_factor_levels(svm_predictions, encoded_test_data$Hypertension)

    svm_conf_matrix <- confusionMatrix(svm_predictions, encoded_test_data$Hypertension, positive = "1")
    svm_accuracy[fold_idx] <- svm_conf_matrix$overall['Accuracy']
    svm_f1[fold_idx] <- svm_conf_matrix$byClass['F1']
    svm_sensitivity[fold_idx] <- svm_conf_matrix$byClass['Sensitivity']
    svm_specificity[fold_idx] <- svm_conf_matrix$byClass['Specificity']
    svm_precision[fold_idx] <- svm_conf_matrix$byClass['Precision']
  }
}

ave_logistic_accuracy <- mean(logistic_accuracy)
ave_logistic_f1 <- mean(logistic_f1)
ave_logistic_sensitivity <- mean(logistic_sensitivity)
ave_logistic_specificity <- mean(logistic_specificity)
ave_logistic_precision <- mean(logistic_precision)

ave_rf_accuracy <- mean(rf_accuracy)
ave_rf_f1 <- mean(rf_f1)
ave_rf_sensitivity <- mean(rf_sensitivity)
ave_rf_specificity <- mean(rf_specificity)
ave_rf_precision <- mean(rf_precision)

ave_svm_accuracy <- mean(svm_accuracy)
ave_svm_f1 <- mean(svm_f1)
ave_svm_sensitivity <- mean(svm_sensitivity)
ave_svm_specificity <- mean(svm_specificity)
ave_svm_precision <- mean(svm_precision)
```

```{r}
results_df <- data.frame(
    Model = c("Logistic Regression", "Random Forest", "SVM"),
    Average_Accuracy = round(c(ave_logistic_accuracy, ave_rf_accuracy, ave_svm_accuracy), 3),
    Average_F1 = round(c(ave_logistic_f1, ave_rf_f1, ave_svm_f1), 3),
    Average_Sensitivity = round(c(ave_logistic_sensitivity, ave_rf_sensitivity, ave_svm_sensitivity), 3),
    Average_Specificity = round(c(ave_logistic_specificity, ave_rf_specificity, ave_svm_specificity), 3),
    Average_Precision = round(c(ave_logistic_precision, ave_rf_precision, ave_svm_precision), 3)
)

# Display the table with rounded values
kable(results_df, format = "markdown", caption = "Model Performance Metrics") %>%
    kable_styling("striped", full_width = F)
```








```{r}
results_long <- results_df %>%
  pivot_longer(cols = starts_with("Average_"), 
               names_to = "Metric", 
               values_to = "Value") %>%
  mutate(Metric = sub("Average_", "", Metric)) 

colors <- c('#90F1EF', '#ffd6e0', '#ffef9f')

fig <- plot_ly(results_long, x = ~Metric, y = ~round(Value, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(Value, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "Model Performance Metrics Comparison",
    yaxis = list(title = 'Score', range = c(0.5, 0.75)), 
    barmode = 'group', # Group bars by model
    paper_bgcolor = 'rgba(245, 246, 249, 1)', # Background color
    showlegend = TRUE, # Show legend for models
    margin = list(l = 50, r = 50, b = 100, t = 100, pad = 4), 
    plot_bgcolor = 'rgba(0, 0, 0, 0)' 
  )

fig
```




# Final model chosed - Random forest

## Feature selection


```{r group by age}
scaled_data_young <- selected_data %>%
  filter(Age < 50)
scaled_data_old <- selected_data %>%
  filter(Age >= 50)


numeric_data_young <- scaled_data_young %>% select_if(is.numeric)
preprocess_params_young <- preProcess(numeric_data_young, method = c("center", "scale"))
scaled_numeric_data_young <- predict(preprocess_params_young, numeric_data_young)
scaled_data_young <- bind_cols(scaled_numeric_data_young, scaled_data_young %>% select_if(Negate(is.numeric)))

numeric_data_old <- scaled_data_old %>% select_if(is.numeric)
preprocess_params_old <- preProcess(numeric_data_old, method = c("center", "scale"))
scaled_numeric_data_old <- predict(preprocess_params_old, numeric_data_old)
scaled_data_old <- bind_cols(scaled_numeric_data_old, scaled_data_old %>% select_if(Negate(is.numeric)))
```



### All population

```{r}
set.seed(3888)

scaled_data$Hypertension <- as.factor(scaled_data$Hypertension)

rf_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + 
                         AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio,
                         data = scaled_data, ntree = 500, type = "classification", probability = TRUE)

rf_model
```



### Young Population
```{r}
set.seed(3888)
rf_young_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_young, ntree = 500)

rf_young_model
```

### Old Population

```{r}
set.seed(3888)
rf_old_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_old, ntree = 500)

rf_old_model
```


### Importance comparison between three models
```{r fig.width=15, fig.height= 6}
importance_young <- as.data.frame(importance(rf_young_model))
importance_overall <- as.data.frame(importance(rf_model))
importance_old <- as.data.frame(importance(rf_old_model))


importance_young$Variable <- rownames(importance_young)
importance_overall$Variable <- rownames(importance_overall)
importance_old$Variable <- rownames(importance_old)


all_vars <- unique(c(importance_young$Variable, importance_overall$Variable, importance_old$Variable))

p2 <- ggplot(importance_young, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "skyblue", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Young Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p1 <- ggplot(importance_overall, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "lightcoral", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Overall Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p3 <- ggplot(importance_old, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "mediumseagreen", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Old Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

grid.arrange(p1, p2, p3, ncol = 3)
```


# PDP Plot

```{r}
rf1 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)
rf2 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio,data = scaled_data_young, ntree = 500, importance = TRUE, proximity = TRUE)
rf3 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data_old, ntree = 500, importance = TRUE, proximity = TRUE)
rf4 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)
rf5 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_young, ntree = 500, importance = TRUE, proximity = TRUE)
rf6 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_old, ntree = 500, importance = TRUE, proximity = TRUE)
```


```{r }
models <- list(rf1, rf2, rf3, rf4, rf5, rf6)
model_names <- c("rf1", "rf2", "rf3", "rf4", "rf5", "rf6")
variables <- c("CarbonhydrateEnergy", "FiberEnergy")

# Create a color palette for the classes
pal <- c( "#3b9ad0", "#f4978e")  # Adjust the colors as needed

# Loop through each variable to generate and print the PDPs
for (variable in variables) {
  # Initialize an empty list to store plots
  plots <- list()
  
  # Loop through each model to generate the PDPs for the current variable
  for (i in seq_along(models)) {
    model <- models[[i]]
    model_name <- model_names[i]
    
    # Create partial dependence for both classes
    pdp_class0 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 1, rug = TRUE)
    pdp_class1 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 2, rug = TRUE)
    
    # Create the plot
    p <- ggplot() +
      geom_line(data = pdp_class0, aes_string(x = variable, y = "yhat"), color = pal[1], size = 1) +
      geom_line(data = pdp_class1, aes_string(x = variable, y = "yhat"), color = pal[2], size = 1) +
      labs(title = paste("PDP for", variable, "Predicting Hypertension\n(Model:", model_name, ")"),
           y = "Partial Dependence", x = variable) +
      theme_minimal() +
      scale_y_continuous(limits = c(0, 1))  # Set y limits between 0 and 1
    
    # Store the plot in the list
    plots[[i]] <- p
  }
  
  # Combine all plots into one figure for the current variable
  combined_plot <- wrap_plots(plots, ncol = 2)  # Arrange plots in 2 columns
  
  # Print the combined plot
  print(combined_plot)
}  
```



```{r fig.width= 15, fig.height= 18}
models <- list(rf1, rf4, rf2, rf5, rf3, rf6)
model_names <- c("A", "a", "B", "b", "C", "c")
variables <- c("CarbonhydrateEnergy", "FiberEnergy")

# Create a color palette for the classes
pal <- c("#3b9ad0", "#f4978e")  # Adjust the colors as needed

# Extract the legend function
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[[which(sapply(g$grobs, function(x) x$name) == "guide-box")]]
  return(legend)
}

# Loop through each variable to generate and print the PDPs
for (variable in variables) {
  plots <- list()  # Initialize an empty list to store plots
  
  # Loop through each model to generate PDPs for the current variable
  for (i in seq_along(models)) {
    model <- models[[i]]
    model_name <- model_names[i]
    
    # Create partial dependence for both classes
    pdp_class0 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 1, rug = TRUE)
    pdp_class1 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 2, rug = TRUE)
    
    # Create the plot with a shared legend
    p <- ggplot() +
      geom_line(data = pdp_class0, aes_string(x = variable, y = "yhat", color = "'Class 0'"), size = 1) +
      geom_line(data = pdp_class1, aes_string(x = variable, y = "yhat", color = "'Class 1'"), size = 1) +
      labs(title = paste("Model:", model_name), y = "Partial Dependence", x = variable) +
      theme_minimal() +
      scale_y_continuous(limits = c(0, 1)) +  # Set y-axis limits between 0 and 1
      scale_color_manual(values = pal)  # Use the custom color palette
    
    # Store the plot in the list
    plots[[i]] <- p
  }
  
  # Extract the legend from one of the plots
  legend <- get_legend(plots[[1]])

  # Combine all plots into a grid (first column: A, B, C; second column: D, E, F)
  combined_plot <- wrap_plots(plots, ncol = 2, byrow = TRUE)  # Arrange in 2 columns

  # Add the legend below the combined plot
  final_plot <- combined_plot + plot_layout(guides = "collect") & theme(legend.position = "bottom")

  # Print the final plot with the outside legend
  print(final_plot)
}

```
















