---
title: "Project proposal"
author: "Loki"
date: "2024-08-25"
output: html_document
---

```{r setup, include=FALSE}
# Default knitting options
knitr::opts_chunk$set(echo=TRUE, # Echo the code
                      tidy=TRUE, # Nicely dity up code
                      warning=FALSE, # No warnings please 
                      message=FALSE) # No messages please

options(warn=-1) 

# Suppress start up warnings when loading libraries
library <- function(...) {
  suppressPackageStartupMessages(base::library(...))
}
```


```{r libraries}
# Load in all libraries
library(tidyverse)
library(here)      # directory referencing
library(readxl)    # reading Excel files
library(janitor)   # data cleaning 
library(stringr)   # string manimpuation
library(tidyr)     # new tidy functions
```
# Feasibility Analysis
## 1. Research Questions:
Question 1: Is there a dependence of grains and cereals intake (serves/g) on whether individuals have diabetes mellitus?
Question 2: What is the correlation between grains and cereals intake (serves/g) and fasting plasma glucose?
Question 3: What is the correlation between grains and cereals intake (serves/g) and HbA1c?


```{r read_in_raw_data}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
```

```{r}
# Subset raw_bp to include relevant variables
selected_bp <- raw_bp %>%
  select(ABSHID, ABSPID, LEVELP, GRAINS1N, GRAINS2N, DIABBC)

# View the first few rows to ensure the subset is correct
head(selected_bp)
```

```{r}
# Subset raw_bb to include relevant variables
selected_bb <- raw_bb %>%
  select(ABSHID, ABSPID, ABSBID, GLUCFPD, GLUCFREB, HBA1MREB, HBA1PREB, DIABFPGE, DIABPRVE)

# View the first few rows to ensure the subset is correct
head(selected_bb)
```
```{r}
# Merge the datasets on ABSHID and ABSPID
merged_data <- merge(selected_bp, selected_bb, by = c("ABSHID", "ABSPID"))

# View the first few rows of the merged dataset
head(merged_data)
```
```{r}
# Create a data frame with variable names and their meanings
variable_info <- data.frame(
  Variable = c("GRAINS1N", "GRAINS2N", "DIABBC", "GLUCFPD", "GLUCFREB", 
               "HBA1MREB", "HBA1PREB", "DIABFPGE", "DIABPRVE", "ABSHID", 
               "ABSPID", "LEVELP", "ABSBID"),
  Description = c("Grains and cereals intake (serves) Day 1",
                  "Grains and cereals intake (serves) Day 2",
                  "Diabetes mellitus status",
                  "Fasting plasma glucose status (mmol/L)",
                  "Fasting plasma glucose - ranged (mmol/L)",
                  "HbA1c - ranged (mmol/mol)",
                  "HbA1c - ranged (%)",
                  "Diabetes prevalence - fasting plasma glucose (mmol/L)",
                  "Diabetes prevalence - HbA1c (%)",
                  "Household identifier",
                  "Selected person identifier",
                  "Record level identifier for Persons",
                  "Biomedical record identifier")
)

# View the summary table
print(variable_info)
```
## 2. Response Variables:
For Question 1: DIABBC (Diabetes mellitus status) is a binary categorical response variable.
For Question 2: GLUCFPD (Fasting plasma glucose status) is a continuous response variable.
For Question 3: HBA1MREB (HbA1c - ranged in mmol/mol) or HBA1PREB (HbA1c - ranged in %) are continuous response variables.

## 3. Predictor Variables:
Primary Predictor: GRAINS1N and GRAINS2N (Grains and cereals intake in serves, Day 1 and Day 2).


## 4. Potential Confounding Variables:
```{r}
#Find Demographic Variables

```
Demographic Variables: Age, Sex, BMI (AGEC, SEX, BMISC).
Other Dietary Factors: Total calorie intake, intake of other macronutrients or micronutrients, physical activity levels.
Biomarkers: Cholesterol levels, blood pressure.

## 5. Data Summary:
### Response Variable Summaries
```{r}

```

### Predictor Variable Summary:
```{r}

```

# 6. Initial Analyses
```{r}
#Chi-Square Test for Dependence (Question 1):
table_data <- table(merged_data$DIABBC, merged_data$GRAINS1N)
chisq.test(table_data)
```
```{r}
#Correlation Analysis (Questions 2 & 3)
cor(merged_data$GRAINS1N, merged_data$GLUCFPD, use = "complete.obs")
cor(merged_data$GRAINS1N, merged_data$HBA1MREB, use = "complete.obs")
```
# 7. Visualization
```{r}
#Boxplot for Question 1
boxplot(GRAINS1N ~ DIABBC, data = merged_data, 
        main = "Grains and Cereals Intake by Diabetes Status",
        xlab = "Diabetes Status", ylab = "Grains and Cereals Intake (Serves)")
```
```{r}
#Scatter Plot for Questions 2 & 3
plot(merged_data$GRAINS1N, merged_data$GLUCFPD, 
     main = "Correlation between Grains Intake and Fasting Plasma Glucose",
     xlab = "Grains and Cereals Intake (Serves)", ylab = "Fasting Plasma Glucose (mmol/L)")

plot(merged_data$GRAINS1N, merged_data$HBA1MREB, 
     main = "Correlation between Grains Intake and HbA1c",
     xlab = "Grains and Cereals Intake (Serves)", ylab = "HbA1c (mmol/mol)")

```
# What's Next
## 1. Further Analysis and Model Building:
•	Logistic Regression (for Question 1):
o	Use logistic regression to model the probability of diabetes as a function of grains and cereals intake, controlling for confounding variables.

•	Linear Regression (for Questions 2 & 3):
o	Use linear regression to model the relationship between grains and cereals intake and fasting plasma glucose/HbA1c.

## 2. Addressing Potential Confounders:
•	Perform sensitivity analysis to check the robustness of your findings against potential confounders.
•	Consider stratification or matching if confounding is severe.
## 3. Statistical Methods to Consider:
•	Logistic Regression: To examine the dependence of categorical outcomes (like diabetes status) on predictors.
•	Linear Regression: To model the relationship between continuous outcomes (like glucose levels, HbA1c) and predictors.
•	ANOVA: To test for significant differences in means across multiple groups, especially if you expand to include more dietary categories.
•	Multivariate Analysis: If you plan to examine multiple outcomes simultaneously.
## 4. Data Cleaning and Preparation:
•	Handle missing data using imputation or complete case analysis.
•	Standardize or normalize variables if needed, particularly for biomarkers with different scales.
## 5. Final Analysis and Reporting:
•	Conduct the final statistical analysis using the models outlined above.
•	Interpret the results, highlighting significant findings and their implications.
•	Prepare visualizations that effectively communicate your results.




```{r}
# Save the summary table as a CSV file
#write.csv(variable_info, file = "variable_info_summary.csv", row.names = FALSE)
```




