---
title: "Result_OriginalData"
author: "group2"
date: "2024-10-17"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
# Library

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)      
library(readxl)    
library(janitor)   
library(stringr)   
library(tidyr)     
library(dplyr)
library(knitr)
library(ggplot2)
library(factoextra)
library(gt)
library(naniar)
library(ggcorrplot)
library(corrplot)
library(GGally)
library(randomForest)
library(e1071)  
library(caret)
library(plotly)
library(stargazer)
library(broom)
library(kableExtra)
library(gridExtra)
library(patchwork)
library(minerva)
library(mgcv)
library(MASS)
library(glmnet)
library(purrr)
library(htmltools)
library(insight)
```

# Data Import

```{r data import}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
raw_bsp <- read.csv(here("ZIPallFiles","inp13bsp.csv"), header=TRUE)
raw_bcn <- read.csv(here("ZIPallFiles","inp13bcn.csv"), header=TRUE)
raw_bhh <- read.csv(here("ZIPallFiles","inp13bhh.csv"), header=TRUE)
```

# Rename and Created

```{r rename+Identity}
bp_bb <- inner_join(raw_bp, raw_bb, by = "ABSPID") %>%
  rename(HypertensionStatus = HYPBC,
         BodyMass = SABDYMS,
         SocialStatus = SF2SA1QN,
         PersonID = ABSPID,
         Age = AGEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FibreEnergy_Day1 = FIBRPER1,
         FibreEnergy_Day2 = FIBRPER2,
         CarbohydrateEnergy_Day1 = CHOPER1,
         CarbohydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2) %>%
   mutate(Identity = factor(0))

bsp_bhh<- left_join(raw_bsp, raw_bhh,by = c("ABSHID")) %>%
  rename(BodyMass = SABDYMS,
         SocialStatus = SF2SA1DB,
         HouseID = ABSHID,
         Age = AGEEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FibreEnergy_Day1 = FIBRPER1,
         FibreEnergy_Day2 = FIBRPER2,
         CarbohydrateEnergy_Day1 = CHOPER1,
         CarbohydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2)%>%
   mutate(Identity = factor(1))

bcn <- raw_bcn %>%
  rename(HouseID = ABSHID,
         MedicalCondition = ICD10ME)
```

# First Filter

```{r filter 1}
bp_bb <- bp_bb %>%
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    HypertensionStatus == 5,
    BodyMass != 4,
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))

bsp_bhh <- bsp_bhh %>% 
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    BodyMass != 4,
    !HouseID %in% bcn$HouseID[bcn$MedicalCondition %in% c(7, 20)],
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))
```

# Merge

```{r merge}
raw_data <- bind_rows(bp_bb,bsp_bhh)
```

# Variable Selection

```{r variable selection}
selected_data <- raw_data %>%
  dplyr::select(PersonID, HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, Identity,
         AlcoholPercentage_Day1, AlcoholPercentage_Day2,
         Potassium_Day1, Potassium_Day2, Sodium_Day1, Sodium_Day2,
         FibreEnergy_Day1, FibreEnergy_Day2, 
         CarbohydrateEnergy_Day1, CarbohydrateEnergy_Day2)
```

# Type Check 1

```{r variable check before(Table1)}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 1: Variable types table"))
```

# Na values + New Variables + Type Conversion

```{r navalue_new_typeconversion}
selected_data <- selected_data %>%
  mutate(
    across(c(Gender, SocialStatus, SmokeStatus), as.factor),
    BMI = na_if(BMI, 98) %>% na_if(99),
    Systolic = na_if(Systolic, 0) %>% na_if(998) %>% na_if(999),
    Diastolic = na_if(Diastolic, 0) %>% na_if(998) %>% na_if(999),
    Potassium = (Potassium_Day1 + Potassium_Day2) / 2, 
    Sodium = (Sodium_Day1 + Sodium_Day2) / 2,
    FibreEnergy = (FibreEnergy_Day1 + FibreEnergy_Day2)/2,
    CarbohydrateEnergy = (CarbohydrateEnergy_Day1 + CarbohydrateEnergy_Day2) / 2,
    AlcoholPercentage = (AlcoholPercentage_Day1 + AlcoholPercentage_Day2) /2,
    SodiumPotassiumRatio = Sodium / Potassium
    ) %>% 
  dplyr::select(PersonID,HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, AlcoholPercentage, SodiumPotassiumRatio, FibreEnergy, CarbohydrateEnergy, Identity)
```

# Type Check 2

```{r variable check after(Table2)}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 2: Variable types table after correction"))
```

# Duplicate Values

```{r duplicate values(Table3)}
selected_data %>% 
  dplyr::select(-PersonID) %>% 
  distinct() %>% 
  dim() %>% 
  tibble::tibble(Dimension = c("Rows", "Columns"), Count = .) %>% 
  gt() %>%
  tab_header(title = "Dimensions of Unique Tibble") %>%
  tab_caption(caption = md("Table 3: Dimensions of Unique Data Table"))
```

# Categrocial Distribution

```{r categrocial distribution bar plot(Figure1), warning=FALSE}
plot_categorical_distribution <- function(data) {
  categorical_vars <- data %>%
    dplyr::select(where(is.factor) | where(is.character)) %>%
    dplyr::select(-PersonID, -HouseID)
  figure_counter <- 1
  for (var in names(categorical_vars)) {
    non_na_data <- data %>%
      filter(!is.na(.data[[var]]))
    if (nrow(non_na_data) == 0) next
    
    figure_title <- paste("Figure 3.", figure_counter, ": Proportion of Categories in", var, sep = "")
    
    p <- ggplot(non_na_data, aes(x = .data[[var]])) +       
      geom_bar(aes(y = (..count..) / sum(..count..), fill = .data[[var]]), color = "black") +       
      scale_fill_brewer(palette = "Set3") +        
      labs(title = figure_title, x = "Category", y = "Proportion") +       
      theme_minimal() +       
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    print(p)
    figure_counter <- figure_counter + 1
    }
  }

plot_categorical_distribution(selected_data) 
```

# Merge Variable

```{r merge variable}
selected_data <- selected_data %>% 
  mutate(SmokeStatus = recode_factor(SmokeStatus, "2" = "1", "3" = "1","4" = "2","5" = "3"))
```

# Outliers and New Variables

```{r variable creation and outliers}
normal_ranges <- list(Systolic = c(90, 180), Diastolic = c(60, 120), BMI = c(16, 47.5), 
                      SodiumPotassiumRatio = c(0, 4), AlcoholPercentage = c(0, 85), 
                      CarbohydrateEnergy = c(0, 100), FibreEnergy = c(0, 100))

selected_data <- selected_data %>%
  filter(
    (is.na(Systolic) | (Systolic >= normal_ranges$Systolic[1] & Systolic <= normal_ranges$Systolic[2])),
    (is.na(Diastolic) | (Diastolic >= normal_ranges$Diastolic[1] & Diastolic <= normal_ranges$Diastolic[2])),
    (is.na(BMI) | (BMI >= normal_ranges$BMI[1] & BMI <= normal_ranges$BMI[2])),
    (is.na(AlcoholPercentage) | (AlcoholPercentage >= normal_ranges$AlcoholPercentage[1] & AlcoholPercentage <= normal_ranges$AlcoholPercentage[2])),
    (is.na(CarbohydrateEnergy) | (CarbohydrateEnergy >= normal_ranges$CarbohydrateEnergy[1] & CarbohydrateEnergy <= normal_ranges$CarbohydrateEnergy[2])),
    (is.na(FibreEnergy) | (FibreEnergy >= normal_ranges$FibreEnergy[1] & FibreEnergy <= normal_ranges$FibreEnergy[2])),
    (is.na(SodiumPotassiumRatio) | (SodiumPotassiumRatio >= normal_ranges$SodiumPotassiumRatio[1] & SodiumPotassiumRatio <= normal_ranges$SodiumPotassiumRatio[2]))
  ) %>%
  mutate(Hypertension = as.factor(case_when(Systolic >= 120 | Diastolic >= 80 ~ 1,TRUE ~ 0)))
```

# Variable check
```{r variable descriptions(Table4)}
variable_data <- data.frame(
  Variable = c("PersonID", "HouseID", "SocialStatus", "Age", "Gender", "BMI", 
               "Systolic", "Diastolic", "SmokeStatus", "AlcoholPercentage", 
               "SodiumPotassiumRatio", "FibreEnergy", "CarbohydrateEnergy", 
               "Identity"),
  Description = c("Selected person identifier", "Household identifier",
                  "ICD10 Conditions - Mini classification",
                  "Age of person", "Sex of person",
                  "Body Mass Index (BMI) - score measured",
                  "Systolic Blood Pressure (mmHg)",
                  "Diastolic Blood Pressure (mmHg)", "Smoking Status",
                  "Average of Percentage of energy from alcohol (Day1) and Percentage of energy from alcohol (Day2)",
                  "The ratio of the average of Sodium (supplement) Day 1 mg and Sodium (supplement) Day 2 mg over the average of Potassium (total) Day 1 mg and Potassium (total) Day 2 mg",
                  "Average of Percentage of energy from fibre (Day1) and Percentage of energy from fibre (Day2)",
                  "Average of Percentage of energy from carbohydrate (Day1) and Percentage of energy from carbohydrate (Day2)",
                  "Indicates whether the individual is part of the indigenous population"))

variable_table <- variable_data %>%
  gt() %>%
  tab_header(title = "All related variables") %>%
  cols_label(Variable = "Variable",Description = "Description") %>%
  tab_caption(caption = md("**Table 4:** Variable Descriptions"))%>%
  fmt_markdown(columns = everything()) %>%
  tab_options(table.width = pct(100))

variable_table
```

```{r predictor(Table5)}
selected_predictors <- selected_data %>%
  dplyr::select(FibreEnergy, CarbohydrateEnergy)

predictor_table <- data.frame(Variable = names(selected_predictors), Type = sapply(selected_predictors, class)) %>%
  gt() %>%
  tab_header(title = "Variable Names and Their Types") %>%
  tab_caption(caption = md("**Table 5**: Predictor variables and their data types"))

predictor_table
```

```{r confounding(Table6)}
confounding_vars <- selected_data %>%
  dplyr::select(-FibreEnergy, -CarbohydrateEnergy,-PersonID, -HouseID)

confounding_table <- data.frame(Variable = names(confounding_vars), Type = sapply(confounding_vars, class)) %>%
  gt() %>%
  tab_header(title = "Confounding Variables and Their Types") %>%
  tab_caption(caption = md("**Table 6**: Confounding variables and their data types"))

confounding_table
```

```{r demographic and health characteristics by hypertension status(Table 7), warning=FALSE}
format_counts <- function(count, total) {
  paste(count, "(", round(count / total * 100, 1), "%)", sep = "")
}

format_counts_list <- function(values, data, column) {
  sapply(values, function(x) format_counts(sum(data[[column]] == x), nrow(data)))
}

format_mean_sd <- function(mean_val, sd_val) {
  paste(round(mean_val, 2), "(", round(sd_val, 2), ")", sep = "")
}

format_mean_sd_list <- function(data, columns) {
  sapply(columns, function(col) format_mean_sd(mean(data[[col]], na.rm = TRUE), sd(data[[col]], na.rm = TRUE)))
}

get_social_status <- function(data) {
  format_counts_list(1:5, data, "SocialStatus")
}

get_gender <- function(data) {
  format_counts_list(1:2, data, "Gender")
}

get_smoke_status <- function(data) {
  format_counts_list(1:3, data, "SmokeStatus")
}

get_identity <- function(data) {
  format_counts_list(0:1, data, "Identity")
}

get_age <- function(data) {
  c(
    format_counts(sum(data$Age >= 18 & data$Age <= 49), nrow(data)),
    format_counts(sum(data$Age > 50), nrow(data))
  )
}

get_health_metrics <- function(data) {
  format_mean_sd_list(data, c("BMI", "Systolic", "Diastolic", "AlcoholPercentage", "PotassiumSodiumRatio", "FiberEnergy", "CarbonhydrateEnergy"))
}

riskdata <- selected_data %>% filter(Hypertension == 1)
noriskdata <- selected_data %>% filter(Hypertension == 0)

df <- data.frame(
  Variable = c(
    "Lowest 20%, N(%)", "Second quintile, N(%)", "Third quintile, N(%)", "Fourth quintile, N(%)", "Highest 20%*, N(%)",
    "Male, N(%)", "Female, N(%)",
    "Currently Smoker, N(%)", "Ex-Smoker, N(%)", "Never smoked, N(%)",
    "Non Indigenous, N(%)", "Indigenous, N(%)",
    "18-49, N(%)", ">50, N(%)",
    "BMI, kg/m² , mean(SD)", 
    "Systolic, mmHg, mean(SD)", "Diastolic, mmHg, mean(SD)", 
    "Alcohol Percentage, mean(SD)", 
    "Potassium Sodium Ratio, mean(SD)", 
    "Fiber Energy, mean(SD)", 
    "Carbohydrate Energy, mean(SD)"
  ),
  Group = c(
    "SocialStatus", "SocialStatus", "SocialStatus", "SocialStatus", "SocialStatus",
    "Gender", "Gender",
    "SmokeStatus", "SmokeStatus", "SmokeStatus",
    "Identity", "Identity",
    "Age", "Age",
    NA, NA, NA, NA, NA, NA, NA
  ),
  Total = c(
    get_social_status(selected_data),
    get_gender(selected_data),
    get_smoke_status(selected_data),
    get_identity(selected_data),
    get_age(selected_data),
    get_health_metrics(selected_data)
  ),
  HypertensionRisk = c(
    get_social_status(riskdata),
    get_gender(riskdata),
    get_smoke_status(riskdata),
    get_identity(riskdata),
    get_age(riskdata),
    get_health_metrics(riskdata)
  ),
  NoHypertensionRisk = c(
    get_social_status(noriskdata),
    get_gender(noriskdata),
    get_smoke_status(noriskdata),
    get_identity(noriskdata),
    get_age(noriskdata),
    get_health_metrics(noriskdata)
  )
)

total_count <- nrow(selected_data)
risk_count <- nrow(riskdata)
no_risk_count <- nrow(noriskdata)

df %>%
  format_table(ci_brackets = c("(", ")")) %>%
  export_table(format = "html", title = "Table 7: Demographic and Health Characteristics by Hypertension Status")
```

# Missing Values

```{r missing value(Table8), warning = FALSE}
selected_data %>% 
  dplyr::select(-PersonID, -HouseID) %>% 
  miss_var_summary() %>% 
  gt() %>%
  tab_header(title = "Missing Value") %>%
  tab_caption(caption = md("Table 8: Missing Values in Selected Data")) %>%
  cols_label(variable = "Variable Name", n_miss = "Missing Count", pct_miss = "Percentage Missing") %>%
  fmt_number(columns = everything(), decimals = 2)
```

```{r heat map(Figure2)}
corr_matrix <- selected_data%>%
  dplyr::select(where(is.numeric)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  cor(use = "complete.obs")

ggcorrplot(corr_matrix, lab = TRUE, 
           title = "Figure 2: Correlation Heatmap for ALL", 
           colors = c("blue", "white", "red"),
           tl.cex = 9, lab_size = 4) + theme(legend.position = "none")
```

# Regression Imputation

```{r regression imputation(Figure3), fig.height = 5, fig.width= 11}
selected_data$BMI[is.na(selected_data$BMI)] <- lm(BMI ~ Gender + SocialStatus + Age + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$BMI), ])

selected_data$Systolic[is.na(selected_data$Systolic)] <- lm(Systolic ~ Gender + SocialStatus + Age + BMI + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$Systolic), ])

selected_data$Diastolic[is.na(selected_data$Diastolic)] <- lm(Diastolic ~ Gender + SocialStatus + Age + BMI + Systolic + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$Diastolic), ])

vis_miss(selected_data %>% dplyr::select(-PersonID, -HouseID)) +
  ggtitle("Figure 3: Missing Data Visualization of Selected Variables")
```

# Normalization

```{r normalization}
numeric_data <- selected_data %>% dplyr::select_if(is.numeric)
preprocess_params <- preProcess(numeric_data, method = c("center", "scale"))
scaled_numeric_data <- predict(preprocess_params, numeric_data)
normalized_data <- bind_cols(scaled_numeric_data, selected_data %>% dplyr::select_if(Negate(is.numeric)))
```

# Multicollinearity

```{r multicollinearity matrix plot(Figure4), fig.height=10, fig.width=10}
selected_data %>%
  dplyr::select(where(is.numeric), -Systolic, -Diastolic) %>%
  ggscatmat() +
  ggtitle("Figure 4: Scatterplot Matrix of Selected Numeric Variables")
```


# Correlation

## Scatter plot

```{r scatter plot(Figure 5), warning=FALSE, message = FALSE, fig.width = 15}
fiber_syst_plot <- ggplot(selected_data, aes(x = FibreEnergy, y = Systolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +
  labs(title = "Figure 5.1: Fiber Energy vs Systolic", x = "Percentage of Energy from Fiber", y = "Systolic(mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5))

fiber_dias_plot <- ggplot(selected_data, aes(x = FibreEnergy, y = Diastolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +
  labs(title = "Figure 5.2: Fiber Energy vs Diastolic", x = "Percentage of Energy from Fiber", y = "Diastolic(mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5))

carbo_syst_plot <- ggplot(selected_data, aes(x = CarbohydrateEnergy, y = Systolic)) +
  geom_point(color = "#2CA02C", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +
  labs(title = "Figure 5.3: Carbohydrate Energy vs Systolic", x = "Percentage of Energy from Carbohydrate", y = "Systolic(mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5))

carbo_dias_plot <- ggplot(selected_data, aes(x = CarbohydrateEnergy, y = Diastolic)) +
  geom_point(color = "#2CA02C", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +
  labs(title = "Figure 5.4: Carbohydrate Energy vs Diastolic", x = "Percentage of Energy from Carbohydrate", y = "Diastolic(mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fiber_syst_plot, fiber_dias_plot, carbo_syst_plot, carbo_dias_plot, nrow = 2)
```

## Linear correlation table

```{r linar correlation table(Table9)}
fiber_syst_pearson <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "pearson")
fiber_syst_spearman <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "spearman")
fiber_dias_pearson <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "pearson")
fiber_dias_spearman <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "spearman")

carbo_syst_pearson <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "pearson")
carbo_syst_spearman <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "spearman")
carbo_dias_pearson <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "pearson")
carbo_dias_spearman <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "spearman")

correlation_table <- data.frame(
  Measure = c("FibreEnergy vs Systolic", "FibreEnergy vs Diastolic",
              "CarbohydrateEnergy vs Systolic", "CarbohydrateEnergy vs Diastolic"),
  Pearson_Correlation = c(fiber_syst_pearson, fiber_dias_pearson, carbo_syst_pearson, carbo_dias_pearson),
  Spearman_Correlation = c(fiber_syst_spearman, fiber_dias_spearman, carbo_syst_spearman, carbo_dias_spearman))

correlation_table %>%
  gt() %>%
  tab_header(
    title = "Table 9: Correlation Between Nutrient Energy Intake and Blood Pressure",
    subtitle = "Correlation values for both Systolic and Diastolic Blood Pressure") %>%
  fmt_number(columns = c(Pearson_Correlation, Spearman_Correlation), decimals = 2)
```

## Non-linear correlation table

```{r non-linear correlation table(Table10)}
fiber_syst_kendall <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "kendall")
fiber_syst_mic <- mine(selected_data$FibreEnergy, selected_data$Systolic)$MIC
fiber_dias_kendall <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "kendall")
fiber_dias_mic <- mine(selected_data$FibreEnergy, selected_data$Diastolic)$MIC
carbo_syst_kendall <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "kendall")
carbo_syst_mic <- mine(selected_data$CarbohydrateEnergy, selected_data$Systolic)$MIC
carbo_dias_kendall <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "kendall")
carbo_dias_mic <- mine(selected_data$CarbohydrateEnergy, selected_data$Diastolic)$MIC

correlation_table_mic_kendall <- data.frame(
  Measure = c("FibreEnergy vs Systolic", "FibreEnergy vs Diastolic",
              "CarbohydrateEnergy vs Systolic", "CarbohydrateEnergy vs Diastolic"),
  Kendall_Correlation = c(fiber_syst_kendall, fiber_dias_kendall, carbo_syst_kendall, carbo_dias_kendall),
  MIC_Correlation = c(fiber_syst_mic, fiber_dias_mic, carbo_syst_mic, carbo_dias_mic))

correlation_table_mic_kendall %>%
  gt() %>%
  tab_header(
    title = "Table 10: Kendall's Tau and MIC Correlation between Nutrient Energy Intake and Blood Pressure") %>%
  fmt_number(columns = c(Kendall_Correlation, MIC_Correlation), decimals = 3)
```

## Hypothesis testing

```{r}
median_fiber <- median(selected_data$FibreEnergy)
selected_data$FiberGroup <- ifelse(selected_data$FibreEnergy > median_fiber, "High Fiber", "Low Fiber")

median_carb <- median(selected_data$CarbohydrateEnergy)
selected_data$CarboGroup <- ifelse(selected_data$CarbohydrateEnergy > median_carb, "High Carbo", "Low Carbo")
```

```{r Anova table(Table11)}
fiber_systolic_anova <- aov(Systolic ~ FiberGroup, data = selected_data)
fiber_diastolic_anova <- aov(Diastolic ~ FiberGroup, data = selected_data)
carbo_systolic_anova <- aov(Systolic ~ CarboGroup, data = selected_data)
carbo_diastolic_anova <- aov(Diastolic ~ CarboGroup, data = selected_data)

anova_results <- data.frame(
  Measure = c("Fiber vs Systolic", "Fiber vs Diastolic", 
              "Carbohydrate vs Systolic", "Carbohydrate vs Diastolic"),
  F_Statistic = c(summary(fiber_systolic_anova)[[1]][["F value"]][1],
                  summary(fiber_diastolic_anova)[[1]][["F value"]][1],
                  summary(carbo_systolic_anova)[[1]][["F value"]][1],
                  summary(carbo_diastolic_anova)[[1]][["F value"]][1]),
  P_Value = c(summary(fiber_systolic_anova)[[1]][["Pr(>F)"]][1],
              summary(fiber_diastolic_anova)[[1]][["Pr(>F)"]][1],
              summary(carbo_systolic_anova)[[1]][["Pr(>F)"]][1],
              summary(carbo_diastolic_anova)[[1]][["Pr(>F)"]][1])
)

anova_results %>%
  gt() %>%
  tab_header(
    title = "Table 11: ANOVA Test Results: Nutrient Energy Intake vs Blood Pressure"
  ) %>%
  fmt_number(columns = c(F_Statistic, P_Value), decimals = 3) %>%
  cols_label(
    Measure = "Nutrient vs Blood Pressure",
    F_Statistic = "F-Statistic",
    P_Value = "P-Value"
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.align = "center"
  )
```

# Model comparison

## Systolic
```{r model comparision metric[Systolic]}
set.seed(3888)
k <- 5
repeats <- 3

gam_rmse_systolic <- numeric()
gam_mae_systolic <- numeric()
gam_r2_systolic <- numeric()

mlr_rmse_systolic <- numeric()
mlr_mae_systolic <- numeric()
mlr_r2_systolic <- numeric()

robust_rmse_systolic <- numeric()
robust_mae_systolic <- numeric()
robust_r2_systolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Systolic
    #GAM
    gam_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))
    gam_mae <- mean(abs(gam_predictions - actual))
    gam_r2 <- cor(gam_predictions, actual)^2
    
    gam_rmse_systolic <- c(gam_rmse_systolic, gam_rmse)
    gam_mae_systolic <- c(gam_mae_systolic, gam_mae)
    gam_r2_systolic <- c(gam_r2_systolic, gam_r2)
    #Multiple Linear Regression
    mlr_model <- lm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                      Identity + SodiumPotassiumRatio, data = data_train)
    
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))
    mlr_mae <- mean(abs(mlr_predictions - actual))
    mlr_r2 <- cor(mlr_predictions, actual)^2
    
    mlr_rmse_systolic <- c(mlr_rmse_systolic, mlr_rmse)
    mlr_mae_systolic <- c(mlr_mae_systolic, mlr_mae)
    mlr_r2_systolic <- c(mlr_r2_systolic, mlr_r2)
    #Robust Regression
    robust_model <- rlm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                          AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                          Identity + SodiumPotassiumRatio, data = data_train)
    
    robust_predictions <- predict(robust_model, newdata = data_test)
    
    robust_rmse <- sqrt(mean((robust_predictions - actual)^2))
    robust_mae <- mean(abs(robust_predictions - actual))
    robust_r2 <- cor(robust_predictions, actual)^2
    
    robust_rmse_systolic <- c(robust_rmse_systolic, robust_rmse)
    robust_mae_systolic <- c(robust_mae_systolic, robust_mae)
    robust_r2_systolic <- c(robust_r2_systolic, robust_r2)
  }
}

mean_systolic_gam_rmse <- mean(gam_rmse_systolic)
mean_systolic_gam_mae <- mean(gam_mae_systolic)
mean_systolic_gam_r2 <- mean(gam_r2_systolic)

mean_systolic_mlr_rmse <- mean(mlr_rmse_systolic)
mean_systolic_mlr_mae <- mean(mlr_mae_systolic)
mean_systolic_mlr_r2 <- mean(mlr_r2_systolic)

mean_systolic_robust_rmse <- mean(robust_rmse_systolic)
mean_systolic_robust_mae <- mean(robust_mae_systolic)
mean_systolic_robust_r2 <- mean(robust_r2_systolic)
```

```{r model performance[Systolic](Figure6), fig.width=8}
Systolic_df <- data.frame(
    Model = c("GAM", "Multiple Linear Regression", "Robust"),
    RMSE = c(mean_systolic_gam_rmse, mean_systolic_mlr_rmse, mean_systolic_robust_rmse),
    MAE = c(mean_systolic_gam_mae, mean_systolic_mlr_mae, mean_systolic_robust_mae),
    R2 = c(mean_systolic_gam_r2, mean_systolic_mlr_r2, mean_systolic_robust_r2)
)

# Define custom colors for the bar plots
colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 6: RMSE [Left], MAE [Middle], R² [Right] for Full Model and Reduced Model(Systolic)", 
                      font = list(size = 12)))

SystolicFig
```
## Diastolic
```{r model comparision metric[Diastolic]}
set.seed(3888)
gam_rmse_diastolic <- numeric()
gam_mae_diastolic <- numeric()
gam_r2_diastolic <- numeric()

mlr_rmse_diastolic <- numeric()
mlr_mae_diastolic <- numeric()
mlr_r2_diastolic <- numeric()

robust_rmse_diastolic <- numeric()
robust_mae_diastolic <- numeric()
robust_r2_diastolic <- numeric()

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Diastolic
    #GAM
    gam_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))
    gam_mae <- mean(abs(gam_predictions - actual))
    gam_r2 <- cor(gam_predictions, actual)^2
    
    gam_rmse_diastolic <- c(gam_rmse_diastolic, gam_rmse)
    gam_mae_diastolic <- c(gam_mae_diastolic, gam_mae)
    gam_r2_diastolic <- c(gam_r2_diastolic, gam_r2)
    #Multiple Linear Regression
    mlr_model <- lm(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                      Identity + SodiumPotassiumRatio, data = data_train)
    
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))
    mlr_mae <- mean(abs(mlr_predictions - actual))
    mlr_r2 <- cor(mlr_predictions, actual)^2
    
    mlr_rmse_diastolic <- c(mlr_rmse_diastolic, mlr_rmse)
    mlr_mae_diastolic <- c(mlr_mae_diastolic, mlr_mae)
    mlr_r2_diastolic <- c(mlr_r2_diastolic, mlr_r2)
    #Robust Regression
    robust_model <- rlm(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                          AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                          Identity + SodiumPotassiumRatio, data = data_train)
    
    robust_predictions <- predict(robust_model, newdata = data_test)
    
    robust_rmse <- sqrt(mean((robust_predictions - actual)^2))
    robust_mae <- mean(abs(robust_predictions - actual))
    robust_r2 <- cor(robust_predictions, actual)^2
    
    robust_rmse_diastolic <- c(robust_rmse_diastolic, robust_rmse)
    robust_mae_diastolic <- c(robust_mae_diastolic, robust_mae)
    robust_r2_diastolic <- c(robust_r2_diastolic, robust_r2)
  }
}

mean_diastolic_gam_rmse <- mean(gam_rmse_diastolic)
mean_diastolic_gam_mae <- mean(gam_mae_diastolic)
mean_diastolic_gam_r2 <- mean(gam_r2_diastolic)

mean_diastolic_mlr_rmse <- mean(mlr_rmse_diastolic)
mean_diastolic_mlr_mae <- mean(mlr_mae_diastolic)
mean_diastolic_mlr_r2 <- mean(mlr_r2_diastolic)

mean_diastolic_robust_rmse <- mean(robust_rmse_diastolic)
mean_diastolic_robust_mae <- mean(robust_mae_diastolic)
mean_diastolic_robust_r2 <- mean(robust_r2_diastolic)
```

```{r model performance[Diastolic](Figure7), fig.width=8}
Diastolic_df <- data.frame(
    Model = c("GAM", "Multiple Linear Regression", "Robust"),
    RMSE = c(mean_diastolic_gam_rmse, mean_diastolic_mlr_rmse, mean_diastolic_robust_rmse),
    MAE = c(mean_diastolic_gam_mae, mean_diastolic_mlr_mae, mean_diastolic_robust_mae),
    R2 = c(mean_diastolic_gam_r2, mean_diastolic_mlr_r2, mean_diastolic_robust_r2)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(
  Diastolic_rmse, 
  Diastolic_mae, 
  Diastolic_r2, 
  nrows = 1, 
  shareY = FALSE
) %>%
  layout(title = list(text = "Figure 7: RMSE [Left], MAE [Middle], R² [Right] for Full Model and Reduced Model(Diastolic)", 
                      font = list(size = 12)))

DiastolicFig
```

# Full model vs Reduced model
```{r systolic full vs reduced}
set.seed(3888)
k <- 5
repeats <- 3

gam_full_rmse_systolic <- numeric()
gam_full_mae_systolic <- numeric()
gam_full_r2_systolic <- numeric()

gam_reduced_rmse_systolic <- numeric()
gam_reduced_mae_systolic <- numeric()
gam_reduced_r2_systolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Systolic
    #GAM full
    gam_full_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_full_predictions <- predict(gam_full_model, newdata = data_test)
    
    gam_full_rmse <- sqrt(mean((gam_full_predictions - actual)^2))
    gam_full_mae <- mean(abs(gam_full_predictions - actual))
    gam_full_r2 <- cor(gam_full_predictions, actual)^2
    
    gam_full_rmse_systolic <- c(gam_full_rmse_systolic, gam_full_rmse)
    gam_full_mae_systolic <- c(gam_full_mae_systolic, gam_full_mae)
    gam_full_r2_systolic <- c(gam_full_r2_systolic, gam_full_r2)
    
    #GAM reduced
    gam_reduced_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + Identity + s(SodiumPotassiumRatio), data = data_train)
    gam_reduced_predictions <- predict(gam_reduced_model, newdata = data_test)
    
    gam_reduced_rmse <- sqrt(mean((gam_reduced_predictions - actual)^2))
    gam_reduced_mae <- mean(abs(gam_reduced_predictions - actual))
    gam_reduced_r2 <- cor(gam_reduced_predictions, actual)^2
    
    gam_reduced_rmse_systolic <- c(gam_reduced_rmse_systolic, gam_reduced_rmse)
    gam_reduced_mae_systolic <- c(gam_reduced_mae_systolic, gam_reduced_mae)
    gam_reduced_r2_systolic <- c(gam_reduced_r2_systolic, gam_reduced_r2)
  }
}

mean_systolic_gam_full_rmse <- mean(gam_full_rmse_systolic)
mean_systolic_gam_full_mae <- mean(gam_full_mae_systolic)
mean_systolic_gam_full_r2 <- mean(gam_full_r2_systolic)

mean_systolic_gam_reduced_rmse <- mean(gam_reduced_rmse_systolic)
mean_systolic_gam_reduced_mae <- mean(gam_reduced_mae_systolic)
mean_systolic_gam_reduced_r2 <- mean(gam_reduced_r2_systolic)
```

```{r systolic full vs reduced(Figure 8), fig.width=8}
Systolic_df <- data.frame(
    Model = c("GAM Full", "GAM Reduced"),
    RMSE = c(mean_systolic_gam_full_rmse, mean_systolic_gam_reduced_rmse),
    MAE = c(mean_systolic_gam_full_mae, mean_systolic_gam_reduced_mae),
    R2 = c(mean_systolic_gam_full_r2, mean_systolic_gam_reduced_r2)
)

# Define custom colors for the bar plots
colors <- c('#FF5179', '#BF59D9')

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE, titleX = TRUE) %>%
  layout(title = list(text = "Figure 8: RMSE [Left], MAE [Middle], R² [Right] for Full and Reduced GAM (Systolic)", 
                      font = list(size = 12)))

SystolicFig
```

```{r diastolic full vs reduced}
set.seed(3888)
k <- 5
repeats <- 3

gam_full_rmse_diastolic <- numeric()
gam_full_mae_diastolic <- numeric()
gam_full_r2_diastolic <- numeric()

gam_reduced_rmse_diastolic <- numeric()
gam_reduced_mae_diastolic <- numeric()
gam_reduced_r2_diastolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Diastolic
    #GAM full
    gam_full_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_full_predictions <- predict(gam_full_model, newdata = data_test)
    
    gam_full_rmse <- sqrt(mean((gam_full_predictions - actual)^2))
    gam_full_mae <- mean(abs(gam_full_predictions - actual))
    gam_full_r2 <- cor(gam_full_predictions, actual)^2
    
    gam_full_rmse_diastolic <- c(gam_full_rmse_diastolic, gam_full_rmse)
    gam_full_mae_diastolic <- c(gam_full_mae_diastolic, gam_full_mae)
    gam_full_r2_diastolic <- c(gam_full_r2_diastolic, gam_full_r2)
    
    #GAM reduced
    gam_reduced_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + Identity + s(SodiumPotassiumRatio), data = data_train)
    gam_reduced_predictions <- predict(gam_reduced_model, newdata = data_test)
    
    gam_reduced_rmse <- sqrt(mean((gam_reduced_predictions - actual)^2))
    gam_reduced_mae <- mean(abs(gam_reduced_predictions - actual))
    gam_reduced_r2 <- cor(gam_reduced_predictions, actual)^2
    
    gam_reduced_rmse_diastolic <- c(gam_reduced_rmse_diastolic, gam_reduced_rmse)
    gam_reduced_mae_diastolic <- c(gam_reduced_mae_diastolic, gam_reduced_mae)
    gam_reduced_r2_diastolic <- c(gam_reduced_r2_diastolic, gam_reduced_r2)
  }
}

mean_diastolic_gam_full_rmse <- mean(gam_full_rmse_diastolic)
mean_diastolic_gam_full_mae <- mean(gam_full_mae_diastolic)
mean_diastolic_gam_full_r2 <- mean(gam_full_r2_diastolic)

mean_diastolic_gam_reduced_rmse <- mean(gam_reduced_rmse_diastolic)
mean_diastolic_gam_reduced_mae <- mean(gam_reduced_mae_diastolic)
mean_diastolic_gam_reduced_r2 <- mean(gam_reduced_r2_diastolic)
```

```{r diastolic full vs reduced(Figure 9), fig.width=8}
Diastolic_df <- data.frame(
    Model = c("GAM Full", "GAM Reduced"),
    RMSE = c(mean_diastolic_gam_full_rmse, mean_diastolic_gam_reduced_rmse),
    MAE = c(mean_diastolic_gam_full_mae, mean_diastolic_gam_reduced_mae),
    R2 = c(mean_diastolic_gam_full_r2, mean_diastolic_gam_reduced_r2)
)

# Define custom colors for the bar plots
colors <- c('#FF5179', '#BF59D9')

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE, titleX = TRUE) %>%
  layout(title = list(text = "Figure 9: RMSE [Left], MAE [Middle], R² [Right] for Full and Reduced GAM (Diastolic)", 
                      font = list(size = 12)))

DiastolicFig
```

# Stratify

## Function for CV

```{r function for systiolic cv}
k <- 5
repeats <- 3

sys_cv <- function(data) {
  n <- nrow(data)

  folds <- map(1:repeats, ~ sample(rep(1:k, length.out = n)))
  
  results <- map_dfr(folds, function(fold) {
    map_dfr(1:k, function(i) {
      train_idx <- which(fold != i)
      test_idx <- which(fold == i)
      
      data_train <- data[train_idx, ]
      data_test <- data[test_idx, ]
      
      actual <- data_test$Systolic
      
      gam_model <- gam(Systolic ~ s(FibreEnergy) + s(CarbohydrateEnergy), data = data_train)
      
      gam_predictions <- predict(gam_model, newdata = data_test)
      
      tibble(
        rmse = sqrt(mean((gam_predictions - actual)^2)),
        mae = mean(abs(gam_predictions - actual)),
        r2 = cor(gam_predictions, actual)^2
      )
    })
  })

  summary_results <- results %>%
    summarize(
      mean_rmse = mean(rmse),
      mean_mae = mean(mae),
      mean_r2 = mean(r2)
    )
  
  return(summary_results)
}
```

```{r function for diastolic cv}
dia_cv <- function(data) {
  n <- nrow(data)

  folds <- map(1:repeats, ~ sample(rep(1:k, length.out = n)))
  
  results <- map_dfr(folds, function(fold) {
    map_dfr(1:k, function(i) {
      train_idx <- which(fold != i)
      test_idx <- which(fold == i)
      
      data_train <- data[train_idx, ]
      data_test <- data[test_idx, ]
      
      actual <- data_test$Diastolic
      
      gam_model <- gam(Diastolic ~ s(FibreEnergy) + s(CarbohydrateEnergy), data = data_train)
      
      gam_predictions <- predict(gam_model, newdata = data_test)
      
      tibble(
        rmse = sqrt(mean((gam_predictions - actual)^2)),
        mae = mean(abs(gam_predictions - actual)),
        r2 = cor(gam_predictions, actual)^2
      )
    })
  })

  summary_results <- results %>%
    summarize(
      mean_rmse = mean(rmse),
      mean_mae = mean(mae),
      mean_r2 = mean(r2)
    )
  
  return(summary_results)
}
```

## By Age
### Group
```{r age group, warning = FALSE, message=FALSE}
age1 <- selected_data %>% filter(Age < 35) %>% select_if(is.numeric)

age2 <- selected_data %>%
  filter(Age >= 35 & Age < 65) %>% select_if(is.numeric)

age3 <- selected_data %>% filter(Age >= 65) %>% select_if(is.numeric)
```

### Systolic
```{r age systolic result}
set.seed(3888)
results_age1 <- sys_cv(age1)
results_age2 <- sys_cv(age2)
results_age3 <- sys_cv(age3)

gam_rmse_diastolic_ag1 <- results_age1$mean_rmse
gam_mae_diastolic_ag1 <- results_age1$mean_mae
gam_r2_diastolic_ag1 <- results_age1$mean_r2

gam_rmse_diastolic_ag2 <- results_age2$mean_rmse
gam_mae_diastolic_ag2 <- results_age2$mean_mae
gam_r2_diastolic_ag2 <- results_age2$mean_r2

gam_rmse_diastolic_ag3 <- results_age3$mean_rmse
gam_mae_diastolic_ag3 <- results_age3$mean_mae
gam_r2_diastolic_ag3 <- results_age3$mean_r2
```

```{r age systolic(Figure10), fig.width=8}
Systolic_df <- data.frame(
    Model = c("18-34", "35-64","65-85"),
    RMSE = c(gam_rmse_diastolic_ag1, gam_rmse_diastolic_ag2, gam_rmse_diastolic_ag3),
    MAE = c(gam_mae_diastolic_ag1, gam_mae_diastolic_ag2, gam_mae_diastolic_ag3),
    R2 = c(gam_r2_diastolic_ag1, gam_r2_diastolic_ag2, gam_r2_diastolic_ag3)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

AgeSystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 10: RMSE [Left], MAE [Middle], R² [Right] for Age Groups (Systolic)"), 
                      font = list(size = 12))

AgeSystolicFig
```

### Diastolic

```{r age diastolic result}
set.seed(3888)
results_age1 <- dia_cv(age1)
results_age2 <- dia_cv(age2)
results_age3 <- dia_cv(age3)

gam_rmse_diastolic_ag1 <- results_age1$mean_rmse
gam_mae_diastolic_ag1 <- results_age1$mean_mae
gam_r2_diastolic_ag1 <- results_age1$mean_r2

gam_rmse_diastolic_ag2 <- results_age2$mean_rmse
gam_mae_diastolic_ag2 <- results_age2$mean_mae
gam_r2_diastolic_ag2 <- results_age2$mean_r2

gam_rmse_diastolic_ag3 <- results_age3$mean_rmse
gam_mae_diastolic_ag3 <- results_age3$mean_mae
gam_r2_diastolic_ag3 <- results_age3$mean_r2
```


```{r age diastolic(Figure11), fig.width=8}
Diastolic_df <- data.frame(
    Model = c("18-34", "35-64","65-85"),
    RMSE = c(gam_rmse_diastolic_ag1, gam_rmse_diastolic_ag2, gam_rmse_diastolic_ag3),
    MAE = c(gam_mae_diastolic_ag1, gam_mae_diastolic_ag2, gam_mae_diastolic_ag3),
    R2 = c(gam_r2_diastolic_ag1, gam_r2_diastolic_ag2, gam_r2_diastolic_ag3)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 11: RMSE [Left], MAE [Middle], R² [Right] for Age Groups (Diastolic)"), font = list(size = 12))

DiastolicFig
```

### Partial effect plots - align with the model summary

```{r age model}
age_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age1)
age_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age2)
age_model_3 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age3)

age_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age1)
age_model_5 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age2)
age_model_6 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age3)

```

#### FibreEnergy
```{r age FibreEnergy(Figure12), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2))

fiber_color <- "blue"

plot(age_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.1 18-34 Group - Systolic - FibreEnergy")
plot(age_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.4 18-34 Group - Diastolic - FibreEnergy")
plot(age_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.2 35-64 Group - Systolic - FibreEnergy")
plot(age_model_5, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.5 35-64 Group - Diastolic - FibreEnergy")
plot(age_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.3 65-85 Group - Systolic - FibreEnergy")
plot(age_model_6, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 12.6 65-85 Group - Diastolic - FibreEnergy")

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r age carbohydrateenergy(Figure13), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2))
carb_color <- "red"

plot(age_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.1 18-34 Group - Systolic - CarbohydrateEnergy")
plot(age_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.2 18-34 Group - Diastolic - CarbohydrateEnergy")
plot(age_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.3 35-64 Group - Systolic - CarbohydrateEnergy")
plot(age_model_5, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.4 35-64 Group - Diastolic - CarbohydrateEnergy")
plot(age_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.5 65-85 Group - Systolic - CarbohydrateEnergy")
plot(age_model_6, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 13.6 65-85 Group - Diastolic - CarbohydrateEnergy")
par(mfrow = c(1, 1))
```

## By BMI
### Group
```{r bmi group}
light <- selected_data %>% filter(BMI < 24.9)

normal <- selected_data %>% filter(BMI >= 24.9 & BMI <= 29.9)

heavy <- selected_data %>% filter(BMI > 29.9)
```

### Systolic
```{r bmi result}
set.seed(3888)
results_light <- sys_cv(light)
results_normal <- sys_cv(normal)
results_heavy <- sys_cv(heavy)

gam_rmse_systolic_light <- results_light$mean_rmse
gam_mae_systolic_light <- results_light$mean_mae
gam_r2_systolic_light <- results_light$mean_r2

gam_rmse_systolic_normal <- results_normal$mean_rmse
gam_mae_systolic_normal <- results_normal$mean_mae
gam_r2_systolic_normal <- results_normal$mean_r2

gam_rmse_systolic_heavy <- results_heavy$mean_rmse
gam_mae_systolic_heavy <- results_heavy$mean_mae
gam_r2_systolic_heavy <- results_heavy$mean_r2
```

```{r bmi result(Figure14), fig.width=8}
Systolic_df <- data.frame(
    Group = c("Light BMI", "Normal BMI", "Heavy BMI"),
    RMSE = c(gam_rmse_systolic_light, gam_rmse_systolic_normal, gam_rmse_systolic_heavy),
    MAE = c(gam_mae_systolic_light, gam_mae_systolic_normal, gam_mae_systolic_heavy),
    R2 = c(gam_r2_systolic_light, gam_r2_systolic_normal, gam_r2_systolic_heavy)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 14: RMSE [Left], MAE [Middle], R² [Right] for BMI Groups (Systolic)"), font = list(size = 12))

SystolicFig
```

### Diastolic
```{r bmi diastolic result}
set.seed(3888)
results_light_dias <- dia_cv(light)
results_normal_dias <- dia_cv(normal)
results_heavy_dias <- dia_cv(heavy)

gam_rmse_diastolic_light <- results_light_dias$mean_rmse
gam_mae_diastolic_light <- results_light_dias$mean_mae
gam_r2_diastolic_light <- results_light_dias$mean_r2

gam_rmse_diastolic_normal <- results_normal_dias$mean_rmse
gam_mae_diastolic_normal <- results_normal_dias$mean_mae
gam_r2_diastolic_normal <- results_normal_dias$mean_r2

gam_rmse_diastolic_heavy <- results_heavy_dias$mean_rmse
gam_mae_diastolic_heavy <- results_heavy_dias$mean_mae
gam_r2_diastolic_heavy <- results_heavy_dias$mean_r2
```

```{r bmi diastolic(Figure15), fig.width=8}
Diastolic_df <- data.frame(
    Group = c("Light BMI", "Normal BMI", "Heavy BMI"),
    RMSE = c(gam_rmse_diastolic_light, gam_rmse_diastolic_normal, gam_rmse_diastolic_heavy),
    MAE = c(gam_mae_diastolic_light, gam_mae_diastolic_normal, gam_mae_diastolic_heavy),
    R2 = c(gam_r2_diastolic_light, gam_r2_diastolic_normal, gam_r2_diastolic_heavy)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 15: RMSE [Left], MAE [Middle], R² [Right] for BMI Groups (Diastolic)"), font = list(size = 12))

DiastolicFig
```

### Partial effect plots - align with the model summary

```{r bmi model}
bmi_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = light)
bmi_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = normal)
bmi_model_3 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = heavy)

bmi_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = light)
bmi_model_5 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = normal)
bmi_model_6 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = heavy)

```

#### FibreEnergy
```{r age FibreEnergy(Figure16), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2))

fiber_color <- "blue"

plot(bmi_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.1 Low BMI Group - Systolic - FibreEnergy")
plot(bmi_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.4 Low BMI Group - Diastolic - FibreEnergy")
plot(bmi_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.2 Normal BMI Group - Systolic - FibreEnergy")
plot(bmi_model_5, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.5 Normal BMI Group - Diastolic - FibreEnergy")
plot(bmi_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.3 High BMI Group - Systolic - FibreEnergy")
plot(bmi_model_6, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 16.6 High BMI Group - Diastolic - FibreEnergy")
par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r age carbohydrateenergy(Figure17), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2))
carb_color <- "red"

plot(bmi_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.1 Low BMI Group - Systolic - CarbohydrateEnergy")
plot(bmi_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.2 Low BMI Group - Diastolic - CarbohydrateEnergy")
plot(bmi_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.3 Normal BMI Group - Systolic - CarbohydrateEnergy")
plot(bmi_model_5, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.4 Normal BMI Group - Diastolic - CarbohydrateEnergy")
plot(bmi_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.5 High BMI Group - Systolic - CarbohydrateEnergy")
plot(bmi_model_6, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 17.6 High BMI Group - Diastolic - CarbohydrateEnergy")
par(mfrow = c(1, 1))
```

## By AlcoholPercentage
### Group
```{r alcohol group}
nonalcohol <- selected_data %>% filter(AlcoholPercentage == 0)

alcohol <- selected_data %>% filter(AlcoholPercentage > 0)
```

### Systolic

```{r alcohol systolic result}
set.seed(3888)
results_nonalcohol <- sys_cv(nonalcohol)
results_alcohol <- sys_cv(alcohol)

gam_rmse_systolic_nonalcohol <- results_nonalcohol$mean_rmse
gam_mae_systolic_nonalcohol <- results_nonalcohol$mean_mae
gam_r2_systolic_nonalcohol <- results_nonalcohol$mean_r2

gam_rmse_systolic_alcohol <- results_alcohol$mean_rmse
gam_mae_systolic_alcohol <- results_alcohol$mean_mae
gam_r2_systolic_alcohol <- results_alcohol$mean_r2
```

```{r alcohol systolic result(Figure18), fig.width=8}
Systolic_alcohol_df <- data.frame(
    Group = c("Non-Alcohol", "Alcohol"),
    RMSE = c(gam_rmse_systolic_nonalcohol, gam_rmse_systolic_alcohol),
    MAE = c(gam_mae_systolic_nonalcohol, gam_mae_systolic_alcohol),
    R2 = c(gam_r2_systolic_nonalcohol, gam_r2_systolic_alcohol)
)

colors_alcohol <- c('#FF5179', '#BF59D9')

rmse_range_alcohol <- c(min(Systolic_alcohol_df$RMSE) * 0.95, max(Systolic_alcohol_df$RMSE) * 1.05)
mae_range_alcohol <- c(min(Systolic_alcohol_df$MAE) * 0.95, max(Systolic_alcohol_df$MAE) * 1.05)
r2_range_alcohol <- c(min(Systolic_alcohol_df$R2) * 0.95, max(Systolic_alcohol_df$R2) * 1.05)

Systolic_rmse_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig_alcohol <- subplot(Systolic_rmse_alcohol, Systolic_mae_alcohol, Systolic_r2_alcohol, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 18: RMSE [Left], MAE [Middle], R² [Right] for Alcohol Groups (Systolic)"), 
                      font = list(size = 12))

SystolicFig_alcohol
```

### Diastolic
```{r alcohol diastolic result}
set.seed(3888)
results_nonalcohol_dias <- dia_cv(nonalcohol)
results_alcohol_dias <- dia_cv(alcohol)

gam_rmse_diastolic_nonalcohol <- results_nonalcohol_dias$mean_rmse
gam_mae_diastolic_nonalcohol <- results_nonalcohol_dias$mean_mae
gam_r2_diastolic_nonalcohol <- results_nonalcohol_dias$mean_r2

gam_rmse_diastolic_alcohol <- results_alcohol_dias$mean_rmse
gam_mae_diastolic_alcohol <- results_alcohol_dias$mean_mae
gam_r2_diastolic_alcohol <- results_alcohol_dias$mean_r2
```

```{r alcohol diastolic result(Figure19), fig.width=8}
Diastolic_alcohol_df <- data.frame(
    Group = c("Non-Alcohol", "Alcohol"),
    RMSE = c(gam_rmse_diastolic_nonalcohol, gam_rmse_diastolic_alcohol),
    MAE = c(gam_mae_diastolic_nonalcohol, gam_mae_diastolic_alcohol),
    R2 = c(gam_r2_diastolic_nonalcohol, gam_r2_diastolic_alcohol)
)

colors_alcohol <- c('#FF5179', '#BF59D9')

rmse_range_alcohol <- c(min(Diastolic_alcohol_df$RMSE) * 0.95, max(Diastolic_alcohol_df$RMSE) * 1.05)
mae_range_alcohol <- c(min(Diastolic_alcohol_df$MAE) * 0.95, max(Diastolic_alcohol_df$MAE) * 1.05)
r2_range_alcohol <- c(min(Diastolic_alcohol_df$R2) * 0.95, max(Diastolic_alcohol_df$R2) * 1.05)

Diastolic_rmse_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors_alcohol,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig_alcohol <- subplot(Diastolic_rmse_alcohol, Diastolic_mae_alcohol, Diastolic_r2_alcohol, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 19: RMSE [Left], MAE [Middle], R² [Right] for Alcohol Groups (Diastolic)"), 
                      font = list(size = 12))

DiastolicFig_alcohol
```

### Partial effect plots - align with the model summary

```{r alcohol model}
alc_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = nonalcohol)
alc_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = alcohol)
alc_model_3 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = nonalcohol)
alc_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = alcohol)
```

#### FibreEnergy
```{r alcohol FibreEnergy(Figure20), fig.width= 12, fig.height=8}
par(mfrow = c(2, 2))

fiber_color <- "blue"

plot(alc_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 20.1 non-alcohol Group - Systolic - FibreEnergy")
plot(alc_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 20.4 non-alcohol Group - Diastolic - FibreEnergy")
plot(alc_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 20.2 alcohol Group - Systolic - FibreEnergy")
plot(alc_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 20.5 alcohol Group - Diastolic - FibreEnergy")

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r alcohol carbohydrateenergy(Figure21), fig.width= 12, fig.height=8}
par(mfrow = c(2, 2))
carb_color <- "red"

plot(alc_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 21.1 non-alcohol Group - Systolic - CarbohydrateEnergy")
plot(alc_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 21.2 non-alcohol Group - Diastolic - CarbohydrateEnergy")
plot(alc_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 21.3 alcohol Group - Systolic - CarbohydrateEnergy")
plot(alc_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 21.4 alcohol Group - Diastolic - CarbohydrateEnergy")
par(mfrow = c(1, 1))
```

## By Gender
### Group
```{r gender group, warning = FALSE, message=FALSE}
male <- selected_data %>% filter(Gender == 1) %>% select_if(is.numeric)

female <- selected_data %>% filter(Gender == 2) %>% select_if(is.numeric)
```


### Systolic
```{r gender systolic}
set.seed(3888)
results_male <- sys_cv(male)
results_female <- sys_cv(female)

gam_rmse_diastolic_male <- results_male$mean_rmse
gam_mae_diastolic_male <- results_male$mean_mae
gam_r2_diastolic_male <- results_male$mean_r2

gam_rmse_diastolic_female <- results_female$mean_rmse
gam_mae_diastolic_female <- results_female$mean_mae
gam_r2_diastolic_female <- results_female$mean_r2
```

```{r gender systolic result(Figure22), fig.width=8}
Systolic_gender_df <- data.frame(
    Model = c("Male", "Female"),
    RMSE = c(gam_rmse_diastolic_male, gam_rmse_diastolic_female),
    MAE = c(gam_mae_diastolic_male, gam_mae_diastolic_female),
    R2 = c(gam_r2_diastolic_male, gam_r2_diastolic_female)
)

colors <- c('#FF5179', '#BF59D9')

rmse_range <- c(min(Systolic_gender_df$RMSE) * 0.95, max(Systolic_gender_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_gender_df$MAE) * 0.95, max(Systolic_gender_df$MAE) * 1.05)
r2_range <- c(min(Systolic_gender_df$R2) * 0.95, max(Systolic_gender_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

GenderSystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 22: RMSE [Left], MAE [Middle], R² [Right] for Gender Groups (Systolic)"), 
                      font = list(size = 12))

GenderSystolicFig
```

### Diastolic
```{r gender diastolic result}
set.seed(3888)

results_male <- dia_cv(male)
results_female <- dia_cv(female)

gam_rmse_diastolic_male <- results_male$mean_rmse
gam_mae_diastolic_male <- results_male$mean_mae
gam_r2_diastolic_male <- results_male$mean_r2

gam_rmse_diastolic_female <- results_female$mean_rmse
gam_mae_diastolic_female <- results_female$mean_mae
gam_r2_diastolic_female <- results_female$mean_r2
```


```{r gender diastolic result(Figure23), fig.width=8}
Diastolic_gender_df <- data.frame(
    Model = c("Male", "Female"),
    RMSE = c(gam_rmse_diastolic_male, gam_rmse_diastolic_female),
    MAE = c(gam_mae_diastolic_male, gam_mae_diastolic_female),
    R2 = c(gam_r2_diastolic_male, gam_r2_diastolic_female)
)

colors <- c('#FF5179', '#BF59D9')

rmse_range <- c(min(Diastolic_gender_df$RMSE) * 0.95, max(Diastolic_gender_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_gender_df$MAE) * 0.95, max(Diastolic_gender_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_gender_df$R2) * 0.95, max(Diastolic_gender_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_mae <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_r2 <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(title = list(text = "Figure 23: RMSE [Left], MAE [Middle], R² [Right] for Gender Groups (Diastolic)"), 
                      font = list(size = 12))

DiastolicFig
```

### Partial effect plots - align with the model summary

```{r gender model}
systolic_male <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = male)
systolic_female <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = female)

diastolic_male <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = male)
diastolic_female <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = female)
```

#### FibreEnergy
```{r gender FibreEnergy(Figure24), fig.width= 12, fig.height=8}
par(mfrow = c(2, 2))

fiber_color <- "blue"

plot(systolic_male, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 24.1 systolic_male Group - Systolic - FibreEnergy")
plot(diastolic_male, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 24.2  diastolic_male Group - Diastolic - FibreEnergy")
plot(systolic_female, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 24.3 systolic_female Group - Systolic - FibreEnergy")
plot(diastolic_female, select = 2, seWithMean = TRUE, shade = TRUE, col = fiber_color, main = "Figure 24.4 distolic_female Group - Diastolic - FibreEnergy")

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r gender carbonhydrateenergy(Figure25), fig.width= 12, fig.height=8}
par(mfrow = c(2, 2))
carb_color <- "red"

plot(systolic_male, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 25.1 systolic_male Group - Systolic - CarbohydrateEnergy")
plot(diastolic_male, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 25.2 diastolic_male Group - Systolic - CarbohydrateEnergy")
plot(systolic_female, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 25.3 systolic_female Group - Diastolic - CarbohydrateEnergy")
plot(diastolic_female, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "Figure 25.4 diastolic_female Group - Diastolic - CarbohydrateEnergy")

par(mfrow = c(1, 1))
```