---
title: "Codes"
author: "mia"
date: "2024-10-10"
output: html_document
---
# Library
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)      
library(readxl)    
library(janitor)   
library(stringr)   
library(tidyr)     
library(dplyr)
library(knitr)
library(ggplot2)
library(factoextra)
library(gt)
library(naniar)
library(ggcorrplot)
library(corrplot)
library(GGally)
library(randomForest)
library(pdp)
library(e1071)  
library(dplyr) 
library(caret)
library(plotly)
library(stargazer)
library(broom)
library(knitr)
library(kableExtra)
library(gridExtra)
library(patchwork)
```
# Data Import
```{r data import}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
raw_bsp <- read.csv(here("ZIPallFiles","inp13bsp.csv"), header=TRUE)
raw_bcn <- read.csv(here("ZIPallFiles","inp13bcn.csv"), header=TRUE)
raw_bhh <- read.csv(here("ZIPallFiles","inp13bhh.csv"), header=TRUE)
```
# Rename and Created
```{r rename+Identity}
bp_bb <- inner_join(raw_bp, raw_bb, by = "ABSPID") %>%
  rename(HypertensionStatus = HYPBC,
         BodyMass = SABDYMS,
         SocialStatus = SF2SA1QN,
         PersonID = ABSPID,
         Age = AGEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2) %>%
   mutate(Identity = factor(0))

bsp_bhh<- left_join(raw_bsp, raw_bhh,by = c("ABSHID")) %>%
  rename(BodyMass = SABDYMS,
         SocialStatus = SF2SA1DB,
         HouseID = ABSHID,
         Age = AGEEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2)%>%
   mutate(Identity = factor(1))

bcn <- raw_bcn %>%
  rename(HouseID = ABSHID,
         MedicalCondition = ICD10ME)
```
# First Filter
```{r filter 1}
bp_bb <- bp_bb %>%
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    HypertensionStatus == 5,
    BodyMass != 4,
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))

bsp_bhh <- bsp_bhh %>% 
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    BodyMass != 4,
    !HouseID %in% bcn$HouseID[bcn$MedicalCondition %in% c(7, 20)],
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))
```
# Merge
```{r merge}
raw_data <- bind_rows(bp_bb,bsp_bhh)
```
# Variable Selection
```{r variable selection}
selected_data <- raw_data %>%
select(
    PersonID, HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, Identity,
    AlcoholPercentage_Day1, AlcoholPercentage_Day2,
    Potassium_Day1, Potassium_Day2, 
    Sodium_Day1, Sodium_Day2,
    FiberEnergy_Day1, FiberEnergy_Day2, 
    CarbonhydrateEnergy_Day1, CarbonhydrateEnergy_Day2
  )
```
# Type Check 1
```{r variable check before}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 1: Variable types table"))
```
# Na values + New Variables + Type Conversion
```{r navalue_new_typeconversion}
selected_data <- selected_data %>%
  mutate(
    across(c(Gender, SocialStatus, SmokeStatus), as.factor),
    BMI = na_if(BMI, 98) %>% na_if(99),
    Systolic = na_if(Systolic, 0) %>% na_if(998) %>% na_if(999),
    Diastolic = na_if(Diastolic, 0) %>% na_if(998) %>% na_if(999),
    Potassium = (Potassium_Day1 + Potassium_Day2) / 2, 
    Sodium = (Sodium_Day1 + Sodium_Day2) / 2,
    FiberEnergy = (FiberEnergy_Day1 + FiberEnergy_Day2)/2,
    CarbonhydrateEnergy = (CarbonhydrateEnergy_Day1 + CarbonhydrateEnergy_Day2) / 2,
    AlcoholPercentage = (AlcoholPercentage_Day1 + AlcoholPercentage_Day2) /2,
    PotassiumSodiumRatio = Sodium / Potassium
    ) %>% 
  select(PersonID,HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, AlcoholPercentage, PotassiumSodiumRatio, FiberEnergy, CarbonhydrateEnergy, Identity)
```
# Type Check 2
```{r variable check after}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 2: Variable types table after correction"))
```
# Duplicate Values
```{r duplicate values}
selected_data %>% 
  select(-PersonID) %>% 
  distinct() %>% 
  dim() %>% 
  tibble::tibble(Dimension = c("Rows", "Columns"), Count = .) %>% 
  gt() %>%
  tab_header(title = "Dimensions of Unique Tibble") %>%
  tab_caption(caption = md("Table 3: Dimensions of Unique Data Table"))
```

# Categrocial Distribution
```{r categrocial distribution bar plot}
plot_categorical_distribution <- function(data) {
  categorical_vars <- data %>%
    select(where(is.factor) | where(is.character)) %>%
    select(-PersonID, -HouseID)
  
  for (var in names(categorical_vars)) {
    non_na_data <- data %>%
      filter(!is.na(.data[[var]]))
    
    if (nrow(non_na_data) == 0) next
    
    p <- ggplot(non_na_data, aes(x = .data[[var]])) +
      geom_bar(aes(y = (..count..) / sum(..count..), fill = .data[[var]]), color = "black") +
      scale_fill_brewer(palette = "Set3") + 
      labs(title = paste("Proportion of Categories in", var), 
           x = "Category", 
           y = "Proportion") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  
    
    print(p)
  }
}

plot_categorical_distribution(selected_data) 
```
# Merge Variable
```{r merge variable}
selected_data <- selected_data %>%
  mutate(SmokeStatus = recode_factor(SmokeStatus, `2` = "1", `3` = "1", `4` = "2", `5` = "3"))
```

# Outliers and New Variables
```{r variable creation and outliers}
normal_ranges <- list(Systolic = c(90, 180),
                      Diastolic = c(60, 120),
                      BMI = c(16, 47.5), 
                      PotassiumSodiumRatio = c(0, 4), 
                      AlcoholPercentage = c(0, 85), 
                      CarbonhydrateEnergy = c(0, 100), 
                      FiberEnergy = c(0, 100))

selected_data <- selected_data %>%
  filter(
    (is.na(Systolic) | (Systolic >= normal_ranges$Systolic[1] & Systolic <= normal_ranges$Systolic[2])),
    (is.na(Diastolic) | (Diastolic >= normal_ranges$Diastolic[1] & Diastolic <= normal_ranges$Diastolic[2])),
    (is.na(BMI) | (BMI >= normal_ranges$BMI[1] & BMI <= normal_ranges$BMI[2])),
    (is.na(AlcoholPercentage) | (AlcoholPercentage >= normal_ranges$AlcoholPercentage[1] & AlcoholPercentage <= normal_ranges$AlcoholPercentage[2])),
    (is.na(CarbonhydrateEnergy) | (CarbonhydrateEnergy >= normal_ranges$CarbonhydrateEnergy[1] & CarbonhydrateEnergy <= normal_ranges$CarbonhydrateEnergy[2])),
    (is.na(FiberEnergy) | (FiberEnergy >= normal_ranges$FiberEnergy[1] & FiberEnergy <= normal_ranges$FiberEnergy[2])),
    (is.na(PotassiumSodiumRatio) | (PotassiumSodiumRatio >= normal_ranges$PotassiumSodiumRatio[1] & PotassiumSodiumRatio <= normal_ranges$PotassiumSodiumRatio[2]))
  ) %>%
  mutate(
    Hypertension = as.factor(case_when(
      Systolic >= 120 | Diastolic >= 80 ~ 1,
      TRUE ~ 0
    ))
  )
 # select(-Sodium, -Potassium)
```
```{r}
str(selected_data)
```

# Missing Values
```{r missing value, warning = FALSE}
selected_data %>% 
  select(-PersonID, -HouseID) %>% 
  miss_var_summary() %>% 
  gt() %>%
  tab_header(title = "Missing Value") %>%
  tab_caption(caption = md("Table 4: Missing Values in Selected Data")) %>%
  cols_label(variable = "Variable Name", n_miss = "Missing Count", pct_miss = "Percentage Missing") %>%
  fmt_number(columns = everything(), decimals = 2)
```

```{r heat map}
corr_matrix <- selected_data%>%
  select(where(is.numeric)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  cor(use = "complete.obs")

ggcorrplot(corr_matrix, lab = TRUE, 
           title = "Correlation Heatmap for ALL", 
           colors = c("blue", "white", "red"),
           tl.cex = 9, lab_size = 4) + theme(legend.position = "none")
```

# Regression Imputation
```{r regression imputation, fig.height = 5, fig.width= 11}
selected_data$BMI[is.na(selected_data$BMI)] <- lm(BMI ~ Gender + SocialStatus + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$BMI), ])

selected_data$Systolic[is.na(selected_data$Systolic)] <- lm(Systolic ~ Gender + SocialStatus + Age + BMI + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$Systolic), ])

selected_data$Diastolic[is.na(selected_data$Diastolic)] <- lm(Diastolic ~ Gender + SocialStatus + Age + BMI + Systolic + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>%
  predict(newdata = selected_data[is.na(selected_data$Diastolic), ])

vis_miss(selected_data %>% 
           select(-PersonID, -HouseID))
```
# Normalization
```{r normalization}
#numeric_cols <- sapply(selected_data, is.numeric)  

#normalized_data <- selected_data
#normalized_data[numeric_cols] <- scale(selected_data[numeric_cols])

numeric_data <- selected_data %>% select_if(is.numeric)

preprocess_params <- preProcess(numeric_data, method = c("center", "scale"))

scaled_numeric_data <- predict(preprocess_params, numeric_data)

scaled_data <- bind_cols(scaled_numeric_data, selected_data %>% select_if(Negate(is.numeric)))
```
# Multicollinearity
```{r multicollinearity, fig.height=10, fig.width=10}
selected_data %>% 
  select(where(is.numeric), -Systolic, -Diastolic) %>% 
  ggscatmat()
```



# Model comparison (5-fold cross validation)
```{r}
set.seed(3888)
k_folds <- 5
n_repeats <- 3

logistic_accuracy <- numeric(k_folds * n_repeats)
logistic_f1 <- numeric(k_folds * n_repeats)
logistic_sensitivity <- numeric(k_folds * n_repeats)
logistic_specificity <- numeric(k_folds * n_repeats)
logistic_precision <- numeric(k_folds * n_repeats)

rf_accuracy <- numeric(k_folds * n_repeats)
rf_f1 <- numeric(k_folds * n_repeats)
rf_sensitivity <- numeric(k_folds * n_repeats)
rf_specificity <- numeric(k_folds * n_repeats)
rf_precision <- numeric(k_folds * n_repeats)

svm_accuracy <- numeric(k_folds * n_repeats)
svm_f1 <- numeric(k_folds * n_repeats)
svm_sensitivity <- numeric(k_folds * n_repeats)
svm_specificity <- numeric(k_folds * n_repeats)
svm_precision <- numeric(k_folds * n_repeats)

ensure_factor_levels <- function(predictions, reference) {
  factor(predictions, levels = levels(reference))
}

for (repeat_idx in 1:n_repeats) {
  folds <- createFolds(scaled_data$Hypertension, k = k_folds, list = TRUE, returnTrain = FALSE)
  
  for (i in 1:k_folds) {
    fold_idx <- (repeat_idx - 1) * k_folds + i

    test_indices <- folds[[i]]
    test_data <- scaled_data[test_indices, ]
    train_data <- scaled_data[-test_indices, ]

    log_model <- glm(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                     data = train_data, family = binomial())
    log_predictions <- predict(log_model, test_data, type = "response")
    log_class <- ifelse(log_predictions > 0.5, 1, 0)
    log_class <- ensure_factor_levels(log_class, test_data$Hypertension)

    log_conf_matrix <- confusionMatrix(log_class, test_data$Hypertension, positive = "1")
    logistic_accuracy[fold_idx] <- log_conf_matrix$overall['Accuracy']
    logistic_f1[fold_idx] <- log_conf_matrix$byClass['F1']
    logistic_sensitivity[fold_idx] <- log_conf_matrix$byClass['Sensitivity']
    logistic_specificity[fold_idx] <- log_conf_matrix$byClass['Specificity']
    logistic_precision[fold_idx] <- log_conf_matrix$byClass['Precision']

    rf_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = train_data, ntree = 500)
    
    rf_predictions <- predict(rf_model, test_data)
    rf_predictions <- ensure_factor_levels(rf_predictions, test_data$Hypertension)

    rf_conf_matrix <- confusionMatrix(rf_predictions, test_data$Hypertension, positive = "1")
    rf_accuracy[fold_idx] <- rf_conf_matrix$overall['Accuracy']
    rf_f1[fold_idx] <- rf_conf_matrix$byClass['F1']
    rf_sensitivity[fold_idx] <- rf_conf_matrix$byClass['Sensitivity']
    rf_specificity[fold_idx] <- rf_conf_matrix$byClass['Specificity']
    rf_precision[fold_idx] <- rf_conf_matrix$byClass['Precision']

    dummy_model <- dummyVars(~ . - Hypertension - PersonID - HouseID - Systolic - Diastolic, data = train_data)
    encoded_train_data <- as.data.frame(predict(dummy_model, newdata = train_data))
    encoded_test_data <- as.data.frame(predict(dummy_model, newdata = test_data))
    encoded_train_data$Hypertension <- train_data$Hypertension
    encoded_test_data$Hypertension <- test_data$Hypertension

    svm_model <- svm(Hypertension ~ ., data = encoded_train_data, kernel = "radial", cost = 1, gamma = 0.1)
    svm_predictions <- predict(svm_model, encoded_test_data)
    svm_predictions <- ensure_factor_levels(svm_predictions, encoded_test_data$Hypertension)

    svm_conf_matrix <- confusionMatrix(svm_predictions, encoded_test_data$Hypertension, positive = "1")
    svm_accuracy[fold_idx] <- svm_conf_matrix$overall['Accuracy']
    svm_f1[fold_idx] <- svm_conf_matrix$byClass['F1']
    svm_sensitivity[fold_idx] <- svm_conf_matrix$byClass['Sensitivity']
    svm_specificity[fold_idx] <- svm_conf_matrix$byClass['Specificity']
    svm_precision[fold_idx] <- svm_conf_matrix$byClass['Precision']
  }
}

ave_logistic_accuracy <- mean(logistic_accuracy)
ave_logistic_f1 <- mean(logistic_f1)
ave_logistic_sensitivity <- mean(logistic_sensitivity)
ave_logistic_specificity <- mean(logistic_specificity)
ave_logistic_precision <- mean(logistic_precision)

ave_rf_accuracy <- mean(rf_accuracy)
ave_rf_f1 <- mean(rf_f1)
ave_rf_sensitivity <- mean(rf_sensitivity)
ave_rf_specificity <- mean(rf_specificity)
ave_rf_precision <- mean(rf_precision)

ave_svm_accuracy <- mean(svm_accuracy)
ave_svm_f1 <- mean(svm_f1)
ave_svm_sensitivity <- mean(svm_sensitivity)
ave_svm_specificity <- mean(svm_specificity)
ave_svm_precision <- mean(svm_precision)
```

```{r}
results_df <- data.frame(
    Model = c("Logistic Regression", "Random Forest", "SVM"),
    Average_Accuracy = c(ave_logistic_accuracy, ave_rf_accuracy, ave_svm_accuracy),
    Average_F1 = c(ave_logistic_f1, ave_rf_f1, ave_svm_f1),
    Average_Sensitivity = c(ave_logistic_sensitivity, ave_rf_sensitivity, ave_svm_sensitivity),
    Average_Specificity = c(ave_logistic_specificity, ave_rf_specificity, ave_svm_specificity),
    Average_Precision = c(ave_logistic_precision, ave_rf_precision, ave_svm_precision)
)

kable(results_df, format = "markdown", caption = "Model Performance Metrics") %>%
    kable_styling("striped", full_width = F)
```
```{r model selection bar plot}
results_long <- results_df %>%
  pivot_longer(cols = starts_with("Average_"), 
               names_to = "Metric", 
               values_to = "Value") %>%
  mutate(Metric = sub("Average_", "", Metric)) 

colors <- c('#90F1EF', '#ffd6e0', '#ffef9f')

fig <- plot_ly(results_long, x = ~Metric, y = ~round(Value, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(Value, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "Figure 5: Model Performance Metrics Comparison",
    yaxis = list(title = 'Score', range = c(0.5, 0.8)), 
    barmode = 'group',
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    showlegend = TRUE, # Show legend for models
    margin = list(l = 50, r = 50, b = 100, t = 100, pad = 4), 
    plot_bgcolor = 'rgba(0, 0, 0, 0)' 
  )

fig
```
# Random Forest Analysis
## Complete models
# Feature Selection
## Group by Age

```{r group by age}
scaled_data_young <- selected_data %>%
  filter(Age < 65)
scaled_data_old <- selected_data %>%
  filter(Age >= 65)

# 对年龄小于 50 岁的组进行标准化
numeric_data_young <- scaled_data_young %>% select_if(is.numeric)
preprocess_params_young <- preProcess(numeric_data_young, method = c("center", "scale"))
scaled_numeric_data_young <- predict(preprocess_params_young, numeric_data_young)

# 重新组合数值列和非数值列
scaled_data_young <- bind_cols(scaled_numeric_data_young, scaled_data_young %>% select_if(Negate(is.numeric)))

# 对年龄大于等于 50 岁的组进行标准化
numeric_data_old <- scaled_data_old %>% select_if(is.numeric)
preprocess_params_old <- preProcess(numeric_data_old, method = c("center", "scale"))
scaled_numeric_data_old <- predict(preprocess_params_old, numeric_data_old)

# 重新组合数值列和非数值列
scaled_data_old <- bind_cols(scaled_numeric_data_old, scaled_data_old %>% select_if(Negate(is.numeric)))
```

```{r}
summary(scaled_data_old)
```

### All population

```{r}
set.seed(3888)

scaled_data$Hypertension <- as.factor(scaled_data$Hypertension)

rf_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + 
                         AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio,
                         data = scaled_data, ntree = 500, type = "classification", probability = TRUE)

rf_model
```




### Indigious Population
```{r}
set.seed(3888)
rf_young_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_young, ntree = 500)

rf_young_model
```

### Non Indigious Population

```{r}
set.seed(3888)
rf_old_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_old, ntree = 500)

rf_old_model
```

### Importance comparison between three models
```{r fig.width=15, fig.height= 6}
importance_young <- as.data.frame(importance(rf_young_model))
importance_overall <- as.data.frame(importance(rf_model))
importance_old <- as.data.frame(importance(rf_old_model))


importance_young$Variable <- rownames(importance_young)
importance_overall$Variable <- rownames(importance_overall)
importance_old$Variable <- rownames(importance_old)


all_vars <- unique(c(importance_young$Variable, importance_overall$Variable, importance_old$Variable))

p2 <- ggplot(importance_young, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "skyblue", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Young Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p1 <- ggplot(importance_overall, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "lightcoral", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Overall Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p3 <- ggplot(importance_old, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "mediumseagreen", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Old Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

grid.arrange(p1, p2, p3, ncol = 3)
```


# PDP Plot

# PDP Plot

```{r}
rf1 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)
rf2 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio,data = scaled_data_young, ntree = 500, importance = TRUE, proximity = TRUE)
rf3 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data_old, ntree = 500, importance = TRUE, proximity = TRUE)
rf4 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)
rf5 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_young, ntree = 500, importance = TRUE, proximity = TRUE)
rf6 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_old, ntree = 500, importance = TRUE, proximity = TRUE)
```


```{r fig.width= 9, fig.height= 11.5}
models <- list(rf1, rf4, rf2, rf5, rf3, rf6)
model_names <- c("A", "a", "B", "b", "C", "c")
variables <- c("CarbonhydrateEnergy", "FiberEnergy")

# Function to rescale the x-axis in the PDP
rescale_pdp <- function(pdp_data, variable_name, original_mean, original_sd) {
  pdp_data[[variable_name]] <- pdp_data[[variable_name]] * original_sd + original_mean
  return(pdp_data)
}

# Create a color palette for the classes
pal <- c("#0496ff", "#d00000")

# Extract the legend function
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[[which(sapply(g$grobs, function(x) x$name) == "guide-box")]]
  return(legend)
}

# Store the original scaling parameters for rescaling
mean_carb_energy <- mean(selected_data$CarbonhydrateEnergy)
sd_carb_energy <- sd(selected_data$CarbonhydrateEnergy)
mean_fiber_energy <- mean(selected_data$FiberEnergy)
sd_fiber_energy <- sd(selected_data$FiberEnergy)

# Modify the PDP plotting code
for (variable in variables) {
  plots <- list()  # Initialize an empty list to store plots
  
  # Loop through each model to generate PDPs for the current variable
  for (i in seq_along(models)) {
    model <- models[[i]]
    model_name <- model_names[i]
    
    # Create partial dependence for both classes
    pdp_class0 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 1, rug = TRUE)
    pdp_class1 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 2, rug = TRUE)
    
    # Rescale the x-axis back to the original scale
    if (variable == "CarbonhydrateEnergy") {
      pdp_class0 <- rescale_pdp(pdp_class0, "CarbonhydrateEnergy", mean_carb_energy, sd_carb_energy)
      pdp_class1 <- rescale_pdp(pdp_class1, "CarbonhydrateEnergy", mean_carb_energy, sd_carb_energy)
    } else if (variable == "FiberEnergy") {
      pdp_class0 <- rescale_pdp(pdp_class0, "FiberEnergy", mean_fiber_energy, sd_fiber_energy)
      pdp_class1 <- rescale_pdp(pdp_class1, "FiberEnergy", mean_fiber_energy, sd_fiber_energy)
    }
    
    # Create the plot with rescaled x-axis
    p <- ggplot() +
      geom_line(data = pdp_class0, aes_string(x = variable, y = "yhat", color = "'Non-Risk'"), size = 1) +
      geom_line(data = pdp_class1, aes_string(x = variable, y = "yhat", color = "'Risk'"), size = 1) +
      labs(title = paste("Model:", model_name), y = "Partial Dependence", x = variable) +
      theme_minimal() +
      scale_y_continuous(limits = c(0, 1)) +  # Set y-axis limits between 0 and 1
      scale_color_manual(values = pal) +  # Use the custom color palette
      theme(
        plot.title = element_text(hjust = 0, size = 14),  # Larger plot title
        axis.title = element_text(size = 12),  # Larger axis titles
        axis.text = element_text(size = 10),   # Larger axis text
        legend.title = element_text(size = 12),  # Larger legend title
        legend.text = element_text(size = 10)    # Larger legend text
      )
    
    # Store the plot in the list
    plots[[i]] <- p
  }
  
  # Extract the legend from one of the plots
  legend <- get_legend(plots[[1]])

  # Combine all plots into a grid (first column: A, B, C; second column: D, E, F)
  combined_plot <- wrap_plots(plots, ncol = 2, byrow = TRUE)  # Arrange in 2 columns

  # Add the legend below the combined plot
  final_plot <- combined_plot + plot_layout(guides = "collect") & theme(legend.position = "bottom")

  # Print the final plot with the outside legend
  print(final_plot)
}
```



```{r}
library(pdp)
library(lattice)  # for 3D plots

# Define the models and the variables of interest
models <- list(rf1, rf2, rf3, rf4, rf5, rf6)
model_names <- c("rf1", "rf2", "rf3", "rf4", "rf5", "rf6")

# Create an empty list to store all the plots
plots <- list()

# Loop through each model to generate 3D PDP plots
for (i in seq_along(models)) {
  model <- models[[i]]
  model_name <- model_names[i]
  
  # Create 3D partial dependence for the two variables
  pdp_3d <- pdp::partial(model, pred.var = c("CarbonhydrateEnergy", "FiberEnergy"), prob = TRUE, 
                         plot = FALSE, type = "classification", rug = TRUE)
  
  # Create the 3D plot and store it in the list
  p <- plotPartial(pdp_3d, levelplot = FALSE, zlab = "yhat", drape = TRUE,
                   colorkey = TRUE, screen = list(z = 220, x = -40),
                   main = paste("3D PDP for", model_name))
  
  # Save the plot to the list
  plots[[i]] <- p
}

# Combine the plots in a 3x2 layout (or 2x3, depending on interpretation)
print(plots[[1]], split = c(1, 1, 3, 2), more = TRUE)  # First row, first column
print(plots[[4]], split = c(1, 2, 3, 2), more = TRUE)  # Second row, first column
print(plots[[2]], split = c(2, 1, 3, 2), more = TRUE)  # First row, second column
print(plots[[5]], split = c(2, 2, 3, 2), more = TRUE)  # Second row, second column
print(plots[[3]], split = c(3, 1, 3, 2), more = TRUE)  # First row, third column
print(plots[[6]], split = c(3, 2, 3, 2))               # Second row, third column
```
## Group by BMI

```{r group by BMI}
# Group by BMI categories
scaled_data_light <- selected_data %>%
  filter(BMI < 24.9)  # Group with BMI < 18.5

scaled_data_normal <- selected_data %>%
  filter(BMI >= 24.9 & BMI <= 29.9)  # Group with BMI between 18.5 and 24.9

scaled_data_heavy <- selected_data %>%
  filter(BMI > 29.9)  # Group with BMI > 24.9

# For the group with BMI < 18.5
numeric_data_light <- scaled_data_light %>% select_if(is.numeric)
preprocess_params_light <- preProcess(numeric_data_light, method = c("center", "scale"))
scaled_numeric_data_light <- predict(preprocess_params_light, numeric_data_light)
scaled_data_light <- bind_cols(scaled_numeric_data_light, scaled_data_light %>% select_if(Negate(is.numeric)))

# For the group with BMI between 18.5 and 24.9
numeric_data_normal <- scaled_data_normal %>% select_if(is.numeric)
preprocess_params_normal <- preProcess(numeric_data_normal, method = c("center", "scale"))
scaled_numeric_data_normal <- predict(preprocess_params_normal, numeric_data_normal)
scaled_data_normal <- bind_cols(scaled_numeric_data_normal, scaled_data_normal %>% select_if(Negate(is.numeric)))

# For the group with BMI > 24.9
numeric_data_heavy <- scaled_data_heavy %>% select_if(is.numeric)
preprocess_params_heavy <- preProcess(numeric_data_heavy, method = c("center", "scale"))
scaled_numeric_data_heavy <- predict(preprocess_params_heavy, numeric_data_heavy)
scaled_data_heavy <- bind_cols(scaled_numeric_data_heavy, scaled_data_heavy %>% select_if(Negate(is.numeric)))
```

```{r}
summary(scaled_data)
```
```{r}
summary(selected_data)
```


### light Population
```{r}
set.seed(3888)
rf_light_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_light, ntree = 500)

```

### Normal Population
```{r}
set.seed(3888)
rf_normal_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_normal, ntree = 500)
```

### Obesity Population
```{r}
set.seed(3888)
rf_heavy_model <- randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, 
                             data = scaled_data_heavy, ntree = 500)
```

### Importance comparison between four models
```{r fig.width=15, fig.height= 6}
importance_overall <- as.data.frame(importance(rf_model))
importance_light <- as.data.frame(importance(rf_light_model))
importance_normal <- as.data.frame(importance(rf_normal_model))
importance_heavy <- as.data.frame(importance(rf_heavy_model))

importance_overall$Variable <- rownames(importance_overall)
importance_light$Variable <- rownames(importance_light)
importance_normal$Variable <- rownames(importance_normal)
importance_heavy$Variable <- rownames(importance_heavy)


all_vars <- unique(c(importance_overall$Variable, importance_light$Variable, importance_normal$Variable, importance_heavy$Variable))

p1 <- ggplot(importance_overall, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "lightcoral", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Overall Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p2 <- ggplot(importance_light, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "skyblue", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Light Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p3 <- ggplot(importance_normal, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "yellow", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Normal Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

p4 <- ggplot(importance_heavy, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_point(color = "mediumseagreen", size = 3) +
  coord_flip() +
  scale_x_discrete(limits = all_vars) +  # 确保变量顺序一致
  labs(x = "Variable", y = "Mean Decrease Gini") +  # 添加清晰的 x 和 y 轴标签
  ggtitle("Heavy Group") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12))

grid.arrange(p1, p2, p3, p4, ncol = 3)
```

# PDP Plot
```{r}
rf21 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)

rf22 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio,data = scaled_data_light, ntree = 500, importance = TRUE, proximity = TRUE)

rf23 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data_normal, ntree = 500, importance = TRUE, proximity = TRUE)

rf24 = randomForest(Hypertension ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = scaled_data_heavy, ntree = 500, importance = TRUE, proximity = TRUE)

rf25 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data, ntree = 500, importance = TRUE, proximity = TRUE)

rf26 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_light, ntree = 500, importance = TRUE, proximity = TRUE)

rf27 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_normal, ntree = 500, importance = TRUE, proximity = TRUE)

rf28 = randomForest(Hypertension ~ BMI + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy,data = scaled_data_heavy, ntree = 500, importance = TRUE, proximity = TRUE)
```


```{r fig.width= 9, fig.height= 11.5}
# Define your models and variable names
models <- list(rf21, rf25, rf22, rf26, rf23, rf27, rf24, rf28)
model_names <- c("O", "o", "L", "l", "N", "n", "H", "h")
variables <- c("CarbonhydrateEnergy", "FiberEnergy")

# Create a color palette for the classes
pal <- c("#0496ff", "#d00000")

# Extract the legend function
get_legend <- function(plot) {
  g <- ggplotGrob(plot)
  legend <- g$grobs[[which(sapply(g$grobs, function(x) x$name) == "guide-box")]]
  return(legend)
}


# Assuming you have stored the scaling parameters
mean_carb_energy <- mean(selected_data$CarbonhydrateEnergy)  # Original mean of the variable
sd_carb_energy <- sd(selected_data$CarbonhydrateEnergy)      # Original standard deviation

mean_fiber_energy <- mean(selected_data$FiberEnergy)         # Original mean of the variable
sd_fiber_energy <- sd(selected_data$FiberEnergy)             # Original standard deviation

# Adjust the x-axis in PDP to reflect original values
rescale_pdp <- function(pdp_data, variable_name, original_mean, original_sd) {
  pdp_data[[variable_name]] <- pdp_data[[variable_name]] * original_sd + original_mean
  return(pdp_data)
}

# Inside your loop for plotting
for (variable in variables) {
  plots <- list()  # Initialize an empty list to store plots
  
  # Loop through each model to generate PDPs for the current variable
  for (i in seq_along(models)) {
    model <- models[[i]]
    model_name <- model_names[i]
    
    # Create partial dependence for both classes
    pdp_class0 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 1, rug = TRUE)
    pdp_class1 <- pdp::partial(model, pred.var = variable, prob = TRUE, 
                               plot = FALSE, type = "classification", 
                               which.class = 2, rug = TRUE)
    
    # Rescale based on the variable being plotted
    if (variable == "CarbonhydrateEnergy") {
      pdp_class0 <- rescale_pdp(pdp_class0, "CarbonhydrateEnergy", mean_carb_energy, sd_carb_energy)
      pdp_class1 <- rescale_pdp(pdp_class1, "CarbonhydrateEnergy", mean_carb_energy, sd_carb_energy)
    } else if (variable == "FiberEnergy") {
      pdp_class0 <- rescale_pdp(pdp_class0, "FiberEnergy", mean_fiber_energy, sd_fiber_energy)
      pdp_class1 <- rescale_pdp(pdp_class1, "FiberEnergy", mean_fiber_energy, sd_fiber_energy)
    }
    
    # Create the plot
    p <- ggplot() +
      geom_line(data = pdp_class0, aes_string(x = variable, y = "yhat", color = "'Non-Risk'"), size = 1) +
      geom_line(data = pdp_class1, aes_string(x = variable, y = "yhat", color = "'Risk'"), size = 1) +
      labs(title = paste("Model:", model_name), y = "Partial Dependence", x = variable) +
      theme_minimal() +
      scale_y_continuous(limits = c(0, 1)) +  # Set y-axis limits between 0 and 1
      scale_color_manual(values = pal) +  # Use the custom color palette
      theme(
        plot.title = element_text(hjust = 0, size = 14),  # Larger plot title
        axis.title = element_text(size = 12),  # Larger axis titles
        axis.text = element_text(size = 10),   # Larger axis text
        legend.title = element_text(size = 12),  # Larger legend title
        legend.text = element_text(size = 10)    # Larger legend text
      )
    
    # Store the plot in the list
    plots[[i]] <- p
  }
  
  # Combine plots and display
  combined_plot <- wrap_plots(plots, ncol = 2, byrow = TRUE)  # Arrange in 2 columns
  final_plot_2 <- combined_plot + plot_layout(guides = "collect") & theme(legend.position = "bottom")
  print(final_plot_2)
}
```





```{r}
library(pdp)
library(lattice)  # for 3D plots

# Define the models and the variables of interest
models <- list(rf21, rf25, rf22, rf26, rf23, rf27, rf24, rf28)
model_names <- c("O", "o", "L", "l", "N", "n", "H", "h")

# Create an empty list to store all the plots
plots <- list()

# Loop through each model to generate 3D PDP plots
for (i in seq_along(models)) {
  model <- models[[i]]
  model_name <- model_names[i]
  
  # Create 3D partial dependence for the two variables
  pdp_3d <- pdp::partial(model, pred.var = c("CarbonhydrateEnergy", "FiberEnergy"), prob = TRUE, 
                         plot = FALSE, type = "classification", rug = TRUE)
  
  # Create the 3D plot and store it in the list
  p <- plotPartial(pdp_3d, levelplot = FALSE, zlab = "yhat", drape = TRUE,
                   colorkey = TRUE, screen = list(z = 220, x = -40),
                   main = paste("3D PDP for", model_name))
  
  # Save the plot to the list
  plots[[i]] <- p
}

# Combine the plots in a 2x4 layout
print(plots[[1]], split = c(1, 1, 4, 2), more = TRUE)  # First row, first column (O)
print(plots[[2]], split = c(1, 2, 4, 2), more = TRUE)  # First row, second column (o)
print(plots[[3]], split = c(2, 1, 4, 2), more = TRUE)  # First row, third column (L)
print(plots[[4]], split = c(2, 2, 4, 2), more = TRUE)  # First row, fourth column (l)
print(plots[[5]], split = c(3, 1, 4, 2), more = TRUE)  # Second row, first column (N)
print(plots[[6]], split = c(3, 2, 4, 2), more = TRUE)  # Second row, second column (n)
print(plots[[7]], split = c(4, 1, 4, 2), more = TRUE)  # Second row, third column (H)
print(plots[[8]], split = c(4, 2, 4, 2))               # Second row, fourth column (h)
```









































