---
title: "Supplementary Material"
author: "Project 1 Group 2"
date: "2024-10-26"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
# Library
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)      
library(readxl)    
library(janitor)   
library(stringr)   
library(tidyr)     
library(dplyr)
library(knitr)
library(ggplot2)
library(factoextra)
library(gt)
library(naniar)
library(ggcorrplot)
library(corrplot)
library(GGally)
library(randomForest)
library(e1071)  
library(caret)
library(plotly)
library(stargazer)
library(broom)
library(kableExtra)
library(gridExtra)
library(patchwork)
library(minerva)
library(mgcv)
library(MASS)
library(glmnet)
library(purrr)
library(insight)
library(grid)
```

# Data Import

```{r data import}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
raw_bsp <- read.csv(here("ZIPallFiles","inp13bsp.csv"), header=TRUE)
raw_bcn <- read.csv(here("ZIPallFiles","inp13bcn.csv"), header=TRUE)
raw_bhh <- read.csv(here("ZIPallFiles","inp13bhh.csv"), header=TRUE)
```

# Rename and Created

```{r rename+Identity}
bp_bb <- inner_join(raw_bp, raw_bb, by = "ABSPID") %>%
  rename(HypertensionStatus = HYPBC,
         BodyMass = SABDYMS,
         SocialStatus = SF2SA1QN,
         PersonID = ABSPID,
         Age = AGEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FibreEnergy_Day1 = FIBRPER1,
         FibreEnergy_Day2 = FIBRPER2,
         CarbohydrateEnergy_Day1 = CHOPER1,
         CarbohydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2) %>%
   mutate(Identity = factor(0))

bsp_bhh<- left_join(raw_bsp, raw_bhh,by = c("ABSHID")) %>%
  rename(BodyMass = SABDYMS,
         SocialStatus = SF2SA1DB,
         HouseID = ABSHID,
         Age = AGEEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FibreEnergy_Day1 = FIBRPER1,
         FibreEnergy_Day2 = FIBRPER2,
         CarbohydrateEnergy_Day1 = CHOPER1,
         CarbohydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2)%>%
   mutate(Identity = factor(1))

bcn <- raw_bcn %>%
  rename(HouseID = ABSHID,
         MedicalCondition = ICD10ME)
```

# First Filter

```{r filter 1}
bp_bb <- bp_bb %>%
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    HypertensionStatus == 5,
    BodyMass != 4,
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))

bsp_bhh <- bsp_bhh %>% 
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    BodyMass != 4,
    !HouseID %in% bcn$HouseID[bcn$MedicalCondition %in% c(7, 20)],
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))
```

# Merge

```{r merge}
raw_data <- bind_rows(bp_bb,bsp_bhh)
```

# Variable Selection

```{r variable selection}
selected_data <- raw_data %>%
  dplyr::select(PersonID, HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, Identity,
         AlcoholPercentage_Day1, AlcoholPercentage_Day2,
         Potassium_Day1, Potassium_Day2, Sodium_Day1, Sodium_Day2,
         FibreEnergy_Day1, FibreEnergy_Day2, 
         CarbohydrateEnergy_Day1, CarbohydrateEnergy_Day2)
```

# Type Check 1

```{r variable check before(Table1)}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Variable types table"))
```

# Na values + New Variables + Type Conversion

```{r navalue_new_typeconversion}
selected_data <- selected_data %>%
  mutate(
    across(c(Gender, SocialStatus, SmokeStatus), as.factor),
    BMI = na_if(BMI, 98) %>% na_if(99),
    Systolic = na_if(Systolic, 0) %>% na_if(998) %>% na_if(999),
    Diastolic = na_if(Diastolic, 0) %>% na_if(998) %>% na_if(999),
    Potassium = (Potassium_Day1 + Potassium_Day2) / 2, 
    Sodium = (Sodium_Day1 + Sodium_Day2) / 2,
    FibreEnergy = (FibreEnergy_Day1 + FibreEnergy_Day2)/2,
    CarbohydrateEnergy = (CarbohydrateEnergy_Day1 + CarbohydrateEnergy_Day2) / 2,
    AlcoholPercentage = (AlcoholPercentage_Day1 + AlcoholPercentage_Day2) /2,
    SodiumPotassiumRatio = Sodium / Potassium
    ) %>% 
  dplyr::select(PersonID,HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, AlcoholPercentage, SodiumPotassiumRatio, FibreEnergy, CarbohydrateEnergy, Identity)
```

# Type Check 2

```{r variable check after(Table2)}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Variable types table after correction"))
```

# Duplicate Values

```{r duplicate values(Table3)}
selected_data %>% 
  dplyr::select(-PersonID) %>% 
  distinct() %>% 
  dim() %>% 
  tibble::tibble(Dimension = c("Rows", "Columns"), Count = .) %>% 
  gt() %>%
  tab_header(title = "Dimensions of Unique Tibble") %>%
  tab_caption(caption = md("Dimensions of Unique Data Table"))
```

# Categrocial Distribution

```{r categrocial distribution bar plot(Figure1), warning=FALSE}
plot_categorical_distribution <- function(data) {
  categorical_vars <- data %>%
    dplyr::select(where(is.factor) | where(is.character)) %>%
    dplyr::select(-PersonID, -HouseID)
  figure_counter <- 1
  for (var in names(categorical_vars)) {
    non_na_data <- data %>%
      filter(!is.na(.data[[var]]))
    if (nrow(non_na_data) == 0) next
    
    figure_title <- paste("Proportion of Categories in ", var, sep = "")
    
    p <- ggplot(non_na_data, aes(x = .data[[var]])) +       
      geom_bar(aes(y = (..count..) / sum(..count..), fill = .data[[var]]), color = "black") +       
      scale_fill_brewer(palette = "Set3") +        
      labs(title = figure_title, x = "Category", y = "Proportion") +       
      theme_minimal() +       
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    print(p)
    figure_counter <- figure_counter + 1
    }
  }

plot_categorical_distribution(selected_data) 
```

# Merge Variable

```{r merge variable}
selected_data <- selected_data %>% 
  mutate(SmokeStatus = recode_factor(SmokeStatus, "2" = "1", "3" = "1","4" = "2","5" = "3"))
```

# Outliers and New Variables

```{r variable creation and outliers}
normal_ranges <- list(Systolic = c(90, 180), Diastolic = c(60, 120), BMI = c(16, 47.5), 
                      SodiumPotassiumRatio = c(0, 4), AlcoholPercentage = c(0, 85), 
                      CarbohydrateEnergy = c(0, 100), FibreEnergy = c(0, 100))

selected_data <- selected_data %>%
  filter(
    (is.na(Systolic) | (Systolic >= normal_ranges$Systolic[1] & Systolic <= normal_ranges$Systolic[2])),
    (is.na(Diastolic) | (Diastolic >= normal_ranges$Diastolic[1] & Diastolic <= normal_ranges$Diastolic[2])),
    (is.na(BMI) | (BMI >= normal_ranges$BMI[1] & BMI <= normal_ranges$BMI[2])),
    (is.na(AlcoholPercentage) | (AlcoholPercentage >= normal_ranges$AlcoholPercentage[1] & AlcoholPercentage <= normal_ranges$AlcoholPercentage[2])),
    (is.na(CarbohydrateEnergy) | (CarbohydrateEnergy >= normal_ranges$CarbohydrateEnergy[1] & CarbohydrateEnergy <= normal_ranges$CarbohydrateEnergy[2])),
    (is.na(FibreEnergy) | (FibreEnergy >= normal_ranges$FibreEnergy[1] & FibreEnergy <= normal_ranges$FibreEnergy[2])),
    (is.na(SodiumPotassiumRatio) | (SodiumPotassiumRatio >= normal_ranges$SodiumPotassiumRatio[1] & SodiumPotassiumRatio <= normal_ranges$SodiumPotassiumRatio[2]))
  ) %>%
  mutate(Hypertension = as.factor(case_when(Systolic >= 120 | Diastolic >= 80 ~ 1,TRUE ~ 0)))
```

# Variable check
```{r variable descriptions(Table4)}
variable_data <- data.frame(
  Variable = c("PersonID", "HouseID", "SocialStatus", "Age", "Gender", "BMI", 
               "Systolic", "Diastolic", "SmokeStatus", "AlcoholPercentage", 
               "SodiumPotassiumRatio", "FibreEnergy", "CarbohydrateEnergy", 
               "Identity"),
  Description = c("Selected person identifier", "Household identifier",
                  "ICD10 Conditions - Mini classification",
                  "Age of person", "Sex of person",
                  "Body Mass Index (BMI) - score measured",
                  "Systolic Blood Pressure (mmHg)",
                  "Diastolic Blood Pressure (mmHg)", "Smoking Status",
                  "Average of Percentage of energy from alcohol (Day1) and Percentage of energy from alcohol (Day2)",
                  "The ratio of the average of Sodium (supplement) Day 1 mg and Sodium (supplement) Day 2 mg over the average of Potassium (total) Day 1 mg and Potassium (total) Day 2 mg",
                  "Average of Percentage of energy from fibre (Day1) and Percentage of energy from fibre (Day2)",
                  "Average of Percentage of energy from carbohydrate (Day1) and Percentage of energy from carbohydrate (Day2)",
                  "Indicates whether the individual is part of the indigenous population"))

variable_table <- variable_data %>%
  gt() %>%
  tab_header(title = "All related variables") %>%
  cols_label(Variable = "Variable",Description = "Description") %>%
  tab_caption(caption = md("Variable Descriptions"))%>%
  fmt_markdown(columns = everything()) %>%
  tab_options(table.width = pct(100))

variable_table
```
This table may require table number as it may be placed in Methodology. 
```{r predictor(Table5)}
selected_predictors <- selected_data %>%
  dplyr::select(FibreEnergy, CarbohydrateEnergy)

predictor_table <- data.frame(Variable = names(selected_predictors), Type = sapply(selected_predictors, class)) %>%
  gt() %>%
  tab_header(title = "Variable Names and Their Types") %>%
  tab_caption(caption = md("Predictor variables and their data types"))

predictor_table
```

```{r confounding(Table6)}
confounding_vars <- selected_data %>%
  dplyr::select(-FibreEnergy, -CarbohydrateEnergy,-PersonID, -HouseID)

confounding_table <- data.frame(Variable = names(confounding_vars), Type = sapply(confounding_vars, class)) %>%
  gt() %>%
  tab_header(title = "Confounding Variables and Their Types") %>%
  tab_caption(caption = md("Confounding variables and their data types"))

confounding_table
```

```{r demographic and health characteristics by hypertension status(Table 7), warning=FALSE}
format_counts <- function(count, total) {
  paste(count, "(", round(count / total * 100, 1), "%)", sep = "")
}

format_counts_list <- function(values, data, column) {
  sapply(values, function(x) format_counts(sum(data[[column]] == x), nrow(data)))
}

format_mean_sd <- function(mean_val, sd_val) {
  paste(round(mean_val, 2), "(", round(sd_val, 2), ")", sep = "")
}

format_mean_sd_list <- function(data, columns) {
  sapply(columns, function(col) format_mean_sd(mean(data[[col]], na.rm = TRUE), sd(data[[col]], na.rm = TRUE)))
}

get_social_status <- function(data) {
  format_counts_list(1:5, data, "SocialStatus")
}

get_gender <- function(data) {
  format_counts_list(1:2, data, "Gender")
}

get_smoke_status <- function(data) {
  format_counts_list(1:3, data, "SmokeStatus")
}

get_identity <- function(data) {
  format_counts_list(0:1, data, "Identity")
}

get_age <- function(data) {
  c(
    format_counts(sum(data$Age >= 18 & data$Age <= 49), nrow(data)),
    format_counts(sum(data$Age > 50), nrow(data))
  )
}

get_health_metrics <- function(data) {
  format_mean_sd_list(data, c("BMI", "Systolic", "Diastolic", "AlcoholPercentage", "SodiumPotassiumRatio", "FibreEnergy", "CarbohydrateEnergy"))
}

riskdata <- selected_data %>% filter(Hypertension == 1)
noriskdata <- selected_data %>% filter(Hypertension == 0)

df <- data.frame(
  Variable = c(
    "Lowest 20%, N(%)", "Second quintile, N(%)", "Third quintile, N(%)", "Fourth quintile, N(%)", "Highest 20%*, N(%)",
    "Male, N(%)", "Female, N(%)",
    "Currently Smoker, N(%)", "Ex-Smoker, N(%)", "Never smoked, N(%)",
    "Non Indigenous, N(%)", "Indigenous, N(%)",
    "18-49, N(%)", ">50, N(%)",
    "BMI, kg/m² , mean(SD)", 
    "Systolic, mmHg, mean(SD)", "Diastolic, mmHg, mean(SD)", 
    "Alcohol Percentage, mean(SD)", 
    "Sodium Potassium Ratio, mean(SD)", 
    "Fibre Energy, mean(SD)", 
    "Carbohydrate Energy, mean(SD)"
  ),
  Group = c(
    "SocialStatus", "SocialStatus", "SocialStatus", "SocialStatus", "SocialStatus",
    "Gender", "Gender",
    "SmokeStatus", "SmokeStatus", "SmokeStatus",
    "Identity", "Identity",
    "Age", "Age",
    NA, NA, NA, NA, NA, NA, NA
  ),
  Total = c(
    get_social_status(selected_data),
    get_gender(selected_data),
    get_smoke_status(selected_data),
    get_identity(selected_data),
    get_age(selected_data),
    get_health_metrics(selected_data)
  ),
  HypertensionRisk = c(
    get_social_status(riskdata),
    get_gender(riskdata),
    get_smoke_status(riskdata),
    get_identity(riskdata),
    get_age(riskdata),
    get_health_metrics(riskdata)
  ),
  NoHypertensionRisk = c(
    get_social_status(noriskdata),
    get_gender(noriskdata),
    get_smoke_status(noriskdata),
    get_identity(noriskdata),
    get_age(noriskdata),
    get_health_metrics(noriskdata)
  )
)

total_count <- nrow(selected_data)
risk_count <- nrow(riskdata)
no_risk_count <- nrow(noriskdata)

df %>%
  format_table(ci_brackets = c("(", ")")) %>%
  export_table(format = "html", title = "Table 1: Demographic and Health Characteristics by Hypertension Status")
```

# Missing Values

```{r missing value(Table7), warning = FALSE}
selected_data %>% 
  dplyr::select(-PersonID, -HouseID) %>% 
  miss_var_summary() %>% 
  gt() %>%
  tab_header(title = "Missing Value") %>%
  tab_caption(caption = md("Missing Values in Selected Data")) %>%
  cols_label(variable = "Variable Name", n_miss = "Missing Count", pct_miss = "Percentage Missing") %>%
  fmt_number(columns = everything(), decimals = 2)
```

```{r heat map(Figure2)}
corr_matrix <- selected_data%>%
  dplyr::select(where(is.numeric)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  cor(use = "complete.obs")

ggcorrplot(corr_matrix, lab = TRUE, 
           title = "Correlation Heatmap for ALL", 
           colors = c("blue", "white", "red"),
           tl.cex = 9, lab_size = 4) + theme(legend.position = "none")
```

# Regression Imputation

```{r regression imputation(Figure3), fig.height = 5, fig.width= 11}
selected_data$BMI[is.na(selected_data$BMI)] <- lm(BMI ~ Gender + SocialStatus + Age + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$BMI), ])

selected_data$Systolic[is.na(selected_data$Systolic)] <- lm(Systolic ~ Gender + SocialStatus + Age + BMI + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$Systolic), ])

selected_data$Diastolic[is.na(selected_data$Diastolic)] <- lm(Diastolic ~ Gender + SocialStatus + Age + BMI + Systolic + SodiumPotassiumRatio + FibreEnergy + CarbohydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% predict(newdata = selected_data[is.na(selected_data$Diastolic), ])

vis_miss(selected_data %>% dplyr::select(-PersonID, -HouseID)) +
  ggtitle("Missing Data Visualization of Selected Variables")
```

# Normalization

```{r normalization}
numeric_data <- selected_data %>% dplyr::select_if(is.numeric)
preprocess_params <- preProcess(numeric_data, method = c("center", "scale"))
scaled_numeric_data <- predict(preprocess_params, numeric_data)
normalized_data <- bind_cols(scaled_numeric_data, selected_data %>% dplyr::select_if(Negate(is.numeric)))
```

# Multicollinearity

```{r multicollinearity matrix plot(Figure4), fig.height=10, fig.width=10}
selected_data %>%
  dplyr::select(where(is.numeric), -Systolic, -Diastolic) %>%
  ggscatmat() +
  ggtitle("Scatterplot Matrix of Selected Numeric Variables")
```


# Correlation

## Scatter plot
```{r warning=FALSE, message = FALSE, fig.width = 12, fig.height= 8}
Fibre_syst_plot <- ggplot(selected_data, aes(x = FibreEnergy, y = Systolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +
  labs(x = "Percentage of Energy from Fibre", y = "Systolic (mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = "A", hjust = 1.5, vjust = 1.5, size = 6, fontface = "bold")

Fibre_dias_plot <- ggplot(selected_data, aes(x = FibreEnergy, y = Diastolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +
  labs(x = "Percentage of Energy from Fibre", y = "Diastolic (mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = "B", hjust = 1.5, vjust = 1.5, size = 6, fontface = "bold")

carbo_syst_plot <- ggplot(selected_data, aes(x = CarbohydrateEnergy, y = Systolic)) +
  geom_point(color = "#2CA02C", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +
  labs(x = "Percentage of Energy from Carbohydrate", y = "Systolic (mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = "C", hjust = 1.5, vjust = 1.5, size = 6, fontface = "bold")

carbo_dias_plot <- ggplot(selected_data, aes(x = CarbohydrateEnergy, y = Diastolic)) +
  geom_point(color = "#2CA02C", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +
  labs(x = "Percentage of Energy from Carbohydrate", y = "Diastolic (mmHg)") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = "D", hjust = 1.5, vjust = 1.5, size = 6, fontface = "bold")

explanation <- textGrob(
  label = "A. Scatter plot of Fibre Energy vs Systolic                 B. Scatter plot of Fibre Energy vs Diastolic\nC. Scatter plot of Carbohydrate Energy vs Systolic   D. Scatter plot of Carbohydrate Energy vs Diastolic",
  x = unit(0, "npc"), y = unit(0.1, "npc"), just = c("left", "bottom"),
  gp = gpar(fontsize = 12, fontface = "italic")
)

grid.arrange(
  arrangeGrob(
    Fibre_syst_plot, Fibre_dias_plot, carbo_syst_plot, carbo_dias_plot,
    nrow = 2,
    top = textGrob("Figure 1: Relationship Between Energy Intake from Fibre and Carbohydrate and Blood Pressure", 
                   gp = gpar(fontsize = 16))
  ),
  explanation,
  heights = unit.c(unit(1, "npc") - unit(3, "lines"), unit(3, "lines"))
)
```



## Linear correlation table
```{r linar correlation table(Table8)}
Fibre_syst_pearson <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "pearson")
Fibre_syst_spearman <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "spearman")
Fibre_dias_pearson <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "pearson")
Fibre_dias_spearman <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "spearman")

carbo_syst_pearson <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "pearson")
carbo_syst_spearman <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "spearman")
carbo_dias_pearson <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "pearson")
carbo_dias_spearman <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "spearman")

correlation_table <- data.frame(
  Measure = c("FibreEnergy vs Systolic", "FibreEnergy vs Diastolic",
              "CarbohydrateEnergy vs Systolic", "CarbohydrateEnergy vs Diastolic"),
  Pearson_Correlation = c(Fibre_syst_pearson, Fibre_dias_pearson, carbo_syst_pearson, carbo_dias_pearson),
  Spearman_Correlation = c(Fibre_syst_spearman, Fibre_dias_spearman, carbo_syst_spearman, carbo_dias_spearman))

correlation_table %>%
  gt() %>%
  tab_header(
    title = "Table 2: Correlation Between Nutrient Energy Intake and Blood Pressure",
    subtitle = "Correlation values for both Systolic and Diastolic Blood Pressure") %>%
  fmt_number(columns = c(Pearson_Correlation, Spearman_Correlation), decimals = 2)
```

## Non-linear correlation table

```{r non-linear correlation table(Table9)}
Fibre_syst_kendall <- cor(selected_data$FibreEnergy, selected_data$Systolic, method = "kendall")
Fibre_syst_mic <- mine(selected_data$FibreEnergy, selected_data$Systolic)$MIC
Fibre_dias_kendall <- cor(selected_data$FibreEnergy, selected_data$Diastolic, method = "kendall")
Fibre_dias_mic <- mine(selected_data$FibreEnergy, selected_data$Diastolic)$MIC
carbo_syst_kendall <- cor(selected_data$CarbohydrateEnergy, selected_data$Systolic, method = "kendall")
carbo_syst_mic <- mine(selected_data$CarbohydrateEnergy, selected_data$Systolic)$MIC
carbo_dias_kendall <- cor(selected_data$CarbohydrateEnergy, selected_data$Diastolic, method = "kendall")
carbo_dias_mic <- mine(selected_data$CarbohydrateEnergy, selected_data$Diastolic)$MIC

correlation_table_mic_kendall <- data.frame(
  Measure = c("FibreEnergy vs Systolic", "FibreEnergy vs Diastolic",
              "CarbohydrateEnergy vs Systolic", "CarbohydrateEnergy vs Diastolic"),
  Kendall_Correlation = c(Fibre_syst_kendall, Fibre_dias_kendall, carbo_syst_kendall, carbo_dias_kendall),
  MIC_Correlation = c(Fibre_syst_mic, Fibre_dias_mic, carbo_syst_mic, carbo_dias_mic))

correlation_table_mic_kendall %>%
  gt() %>%
  tab_header(
    title = "Table 3: Kendall's Tau and MIC Correlation between Nutrient Energy Intake and Blood Pressure") %>%
  fmt_number(columns = c(Kendall_Correlation, MIC_Correlation), decimals = 3)
```

## Hypothesis testing

```{r}
median_Fibre <- median(selected_data$FibreEnergy)
selected_data$FibreGroup <- ifelse(selected_data$FibreEnergy > median_Fibre, "High Fibre", "Low Fibre")

median_carb <- median(selected_data$CarbohydrateEnergy)
selected_data$CarboGroup <- ifelse(selected_data$CarbohydrateEnergy > median_carb, "High Carbo", "Low Carbo")
```

```{r Anova table(Table10)}
Fibre_systolic_anova <- aov(Systolic ~ FibreGroup, data = selected_data)
Fibre_diastolic_anova <- aov(Diastolic ~ FibreGroup, data = selected_data)
carbo_systolic_anova <- aov(Systolic ~ CarboGroup, data = selected_data)
carbo_diastolic_anova <- aov(Diastolic ~ CarboGroup, data = selected_data)

anova_results <- data.frame(
  Measure = c("Fibre vs Systolic", "Fibre vs Diastolic", 
              "Carbohydrate vs Systolic", "Carbohydrate vs Diastolic"),
  F_Statistic = c(summary(Fibre_systolic_anova)[[1]][["F value"]][1],
                  summary(Fibre_diastolic_anova)[[1]][["F value"]][1],
                  summary(carbo_systolic_anova)[[1]][["F value"]][1],
                  summary(carbo_diastolic_anova)[[1]][["F value"]][1]),
  P_Value = c(summary(Fibre_systolic_anova)[[1]][["Pr(>F)"]][1],
              summary(Fibre_diastolic_anova)[[1]][["Pr(>F)"]][1],
              summary(carbo_systolic_anova)[[1]][["Pr(>F)"]][1],
              summary(carbo_diastolic_anova)[[1]][["Pr(>F)"]][1])
)

anova_results %>%
  gt() %>%
  tab_header(
    title = "Table 4: ANOVA Test Results: Nutrient Energy Intake vs Blood Pressure"
  ) %>%
  fmt_number(columns = c(F_Statistic, P_Value), decimals = 3) %>%
  cols_label(
    Measure = "Nutrient vs Blood Pressure",
    F_Statistic = "F-Statistic",
    P_Value = "P-Value"
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.align = "center"
  )
```

# Model comparison

## Systolic
```{r model comparision metric[Systolic]}
set.seed(3888)
k <- 5
repeats <- 3

gam_rmse_systolic <- numeric()
gam_mae_systolic <- numeric()
gam_r2_systolic <- numeric()

mlr_rmse_systolic <- numeric()
mlr_mae_systolic <- numeric()
mlr_r2_systolic <- numeric()

robust_rmse_systolic <- numeric()
robust_mae_systolic <- numeric()
robust_r2_systolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Systolic
    #GAM
    gam_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))
    gam_mae <- mean(abs(gam_predictions - actual))
    gam_r2 <- cor(gam_predictions, actual)^2
    
    gam_rmse_systolic <- c(gam_rmse_systolic, gam_rmse)
    gam_mae_systolic <- c(gam_mae_systolic, gam_mae)
    gam_r2_systolic <- c(gam_r2_systolic, gam_r2)
    #Multiple Linear Regression
    mlr_model <- lm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                      Identity + SodiumPotassiumRatio, data = data_train)
    
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))
    mlr_mae <- mean(abs(mlr_predictions - actual))
    mlr_r2 <- cor(mlr_predictions, actual)^2
    
    mlr_rmse_systolic <- c(mlr_rmse_systolic, mlr_rmse)
    mlr_mae_systolic <- c(mlr_mae_systolic, mlr_mae)
    mlr_r2_systolic <- c(mlr_r2_systolic, mlr_r2)
    #Robust Regression
    robust_model <- rlm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                          AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                          Identity + SodiumPotassiumRatio, data = data_train)
    
    robust_predictions <- predict(robust_model, newdata = data_test)
    
    robust_rmse <- sqrt(mean((robust_predictions - actual)^2))
    robust_mae <- mean(abs(robust_predictions - actual))
    robust_r2 <- cor(robust_predictions, actual)^2
    
    robust_rmse_systolic <- c(robust_rmse_systolic, robust_rmse)
    robust_mae_systolic <- c(robust_mae_systolic, robust_mae)
    robust_r2_systolic <- c(robust_r2_systolic, robust_r2)
  }
}

mean_systolic_gam_rmse <- mean(gam_rmse_systolic)
mean_systolic_gam_mae <- mean(gam_mae_systolic)
mean_systolic_gam_r2 <- mean(gam_r2_systolic)

mean_systolic_mlr_rmse <- mean(mlr_rmse_systolic)
mean_systolic_mlr_mae <- mean(mlr_mae_systolic)
mean_systolic_mlr_r2 <- mean(mlr_r2_systolic)

mean_systolic_robust_rmse <- mean(robust_rmse_systolic)
mean_systolic_robust_mae <- mean(robust_mae_systolic)
mean_systolic_robust_r2 <- mean(robust_r2_systolic)
```

```{r model performance[Systolic](Figure6), fig.width=10}
Systolic_df <- data.frame(
    Model = c("GAM", "Multiple Linear", "Robust"),
    RMSE = c(mean_systolic_gam_rmse, mean_systolic_mlr_rmse, mean_systolic_robust_rmse),
    MAE = c(mean_systolic_gam_mae, mean_systolic_mlr_mae, mean_systolic_robust_mae),
    R2 = c(mean_systolic_gam_r2, mean_systolic_mlr_r2, mean_systolic_robust_r2)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 2: Model Comparison (Systolic)",
      font = list(size = 16),x = 0.5, y = 1.15),
    annotations = list(
      list(text = "A",x = 0, y = 1.05, xref = "paper", yref = "paper", showarrow = FALSE, font = list(size = 14, color = "black"), xanchor = "center"),
      list(text = "B", x = 0.35, y = 1.05, xref = "paper", yref = "paper", showarrow = FALSE, font = list(size = 14, color = "black"), xanchor = "center"),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"),
      list(
        text = "A. RMSE comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.15,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.2,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.25,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )

SystolicFig
```
## Diastolic
```{r model comparision metric[Diastolic]}
set.seed(3888)
gam_rmse_diastolic <- numeric()
gam_mae_diastolic <- numeric()
gam_r2_diastolic <- numeric()

mlr_rmse_diastolic <- numeric()
mlr_mae_diastolic <- numeric()
mlr_r2_diastolic <- numeric()

robust_rmse_diastolic <- numeric()
robust_mae_diastolic <- numeric()
robust_r2_diastolic <- numeric()

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Diastolic
    #GAM
    gam_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))
    gam_mae <- mean(abs(gam_predictions - actual))
    gam_r2 <- cor(gam_predictions, actual)^2
    
    gam_rmse_diastolic <- c(gam_rmse_diastolic, gam_rmse)
    gam_mae_diastolic <- c(gam_mae_diastolic, gam_mae)
    gam_r2_diastolic <- c(gam_r2_diastolic, gam_r2)
    #Multiple Linear Regression
    mlr_model <- lm(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                      Identity + SodiumPotassiumRatio, data = data_train)
    
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))
    mlr_mae <- mean(abs(mlr_predictions - actual))
    mlr_r2 <- cor(mlr_predictions, actual)^2
    
    mlr_rmse_diastolic <- c(mlr_rmse_diastolic, mlr_rmse)
    mlr_mae_diastolic <- c(mlr_mae_diastolic, mlr_mae)
    mlr_r2_diastolic <- c(mlr_r2_diastolic, mlr_r2)
    #Robust Regression
    robust_model <- rlm(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                          AlcoholPercentage + FibreEnergy + CarbohydrateEnergy + 
                          Identity + SodiumPotassiumRatio, data = data_train)
    
    robust_predictions <- predict(robust_model, newdata = data_test)
    
    robust_rmse <- sqrt(mean((robust_predictions - actual)^2))
    robust_mae <- mean(abs(robust_predictions - actual))
    robust_r2 <- cor(robust_predictions, actual)^2
    
    robust_rmse_diastolic <- c(robust_rmse_diastolic, robust_rmse)
    robust_mae_diastolic <- c(robust_mae_diastolic, robust_mae)
    robust_r2_diastolic <- c(robust_r2_diastolic, robust_r2)
  }
}

mean_diastolic_gam_rmse <- mean(gam_rmse_diastolic)
mean_diastolic_gam_mae <- mean(gam_mae_diastolic)
mean_diastolic_gam_r2 <- mean(gam_r2_diastolic)

mean_diastolic_mlr_rmse <- mean(mlr_rmse_diastolic)
mean_diastolic_mlr_mae <- mean(mlr_mae_diastolic)
mean_diastolic_mlr_r2 <- mean(mlr_r2_diastolic)

mean_diastolic_robust_rmse <- mean(robust_rmse_diastolic)
mean_diastolic_robust_mae <- mean(robust_mae_diastolic)
mean_diastolic_robust_r2 <- mean(robust_r2_diastolic)
```

```{r model performance[Diastolic](Figure7),fig.width=10}
Diastolic_df <- data.frame(
    Model = c("GAM", "Multiple Linear", "Robust"),
    RMSE = c(mean_diastolic_gam_rmse, mean_diastolic_mlr_rmse, mean_diastolic_robust_rmse),
    MAE = c(mean_diastolic_gam_mae, mean_diastolic_mlr_mae, mean_diastolic_robust_mae),
    R2 = c(mean_diastolic_gam_r2, mean_diastolic_mlr_r2, mean_diastolic_robust_r2)
)

colors <- c('#FF5179', '#BF59D9', '#E76BF3')

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(
  Diastolic_rmse, 
  Diastolic_mae, 
  Diastolic_r2, 
  nrows = 1, 
  shareY = FALSE
) %>%
  layout(
    title = list(
      text = "Figure 3: Model Comparison (Diastolic)",
      font = list(size = 16),x = 0.5, y = 1.15),
    annotations = list(
      list(text = "A",x = 0, y = 1.05, xref = "paper", yref = "paper", showarrow = FALSE, font = list(size = 14, color = "black"), xanchor = "center"),
      list(text = "B", x = 0.35, y = 1.05, xref = "paper", yref = "paper", showarrow = FALSE, font = list(size = 14, color = "black"), xanchor = "center"),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"),
      list(
        text = "A. RMSE comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.15,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.2,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model, Multiple Linear Regression and Robust Regression",
        x = 0,
        y = -0.25,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )

DiastolicFig
```

# Full model vs Reduced model
```{r full model reduced model cv[Systolic]}
set.seed(3888)
k <- 5
repeats <- 3

gam_full_rmse_systolic <- numeric()
gam_full_mae_systolic <- numeric()
gam_full_r2_systolic <- numeric()

gam_reduced_rmse_systolic <- numeric()
gam_reduced_mae_systolic <- numeric()
gam_reduced_r2_systolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Systolic
    #GAM full
    gam_full_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_full_predictions <- predict(gam_full_model, newdata = data_test)
    
    gam_full_rmse <- sqrt(mean((gam_full_predictions - actual)^2))
    gam_full_mae <- mean(abs(gam_full_predictions - actual))
    gam_full_r2 <- cor(gam_full_predictions, actual)^2
    
    gam_full_rmse_systolic <- c(gam_full_rmse_systolic, gam_full_rmse)
    gam_full_mae_systolic <- c(gam_full_mae_systolic, gam_full_mae)
    gam_full_r2_systolic <- c(gam_full_r2_systolic, gam_full_r2)
    
    #GAM reduced
    gam_reduced_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + Identity + s(SodiumPotassiumRatio), data = data_train)
    gam_reduced_predictions <- predict(gam_reduced_model, newdata = data_test)
    
    gam_reduced_rmse <- sqrt(mean((gam_reduced_predictions - actual)^2))
    gam_reduced_mae <- mean(abs(gam_reduced_predictions - actual))
    gam_reduced_r2 <- cor(gam_reduced_predictions, actual)^2
    
    gam_reduced_rmse_systolic <- c(gam_reduced_rmse_systolic, gam_reduced_rmse)
    gam_reduced_mae_systolic <- c(gam_reduced_mae_systolic, gam_reduced_mae)
    gam_reduced_r2_systolic <- c(gam_reduced_r2_systolic, gam_reduced_r2)
  }
}

mean_systolic_gam_full_rmse <- mean(gam_full_rmse_systolic)
mean_systolic_gam_full_mae <- mean(gam_full_mae_systolic)
mean_systolic_gam_full_r2 <- mean(gam_full_r2_systolic)

mean_systolic_gam_reduced_rmse <- mean(gam_reduced_rmse_systolic)
mean_systolic_gam_reduced_mae <- mean(gam_reduced_mae_systolic)
mean_systolic_gam_reduced_r2 <- mean(gam_reduced_r2_systolic)
```

```{r plot[Systolic]}
Systolic_df <- data.frame(
    Model = c("GAM Full", "GAM Reduced"),
    RMSE = c(mean_systolic_gam_full_rmse, mean_systolic_gam_reduced_rmse),
    MAE = c(mean_systolic_gam_full_mae, mean_systolic_gam_reduced_mae),
    R2 = c(mean_systolic_gam_full_r2, mean_systolic_gam_reduced_r2)
)

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 4: GAM Full Model* and Reduced Model** Comparison (Systolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Systolic)",
        x = 0,
        y = -0.15,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Systolic)",
        x = 0,
        y = -0.2, 
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Systolic)",
        x = 0,
        y = -0.25,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ), list(
        text = "*Full Model: Model with all risk factors and carb and fiber",
        x = 0,
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 10),
        xanchor = "left"
      ), list(
        text = "**Reduced Model: Model with only risk factors",
        x = 0,
        y = -0.35,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 10),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )

SystolicFig
```

```{r}
set.seed(3888)
k <- 5
repeats <- 3

gam_full_rmse_diastolic <- numeric()
gam_full_mae_diastolic <- numeric()
gam_full_r2_diastolic <- numeric()

gam_reduced_rmse_diastolic <- numeric()
gam_reduced_mae_diastolic <- numeric()
gam_reduced_r2_diastolic <- numeric()

n <- nrow(normalized_data)

for (r in 1:repeats) {
  folds <- sample(rep(1:k, length.out = n))
  
  for (i in 1:k) {
    train_idx <- which(folds != i)
    test_idx <- which(folds == i)
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Diastolic
    #GAM full
    gam_full_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = data_train)
    
    gam_full_predictions <- predict(gam_full_model, newdata = data_test)
    
    gam_full_rmse <- sqrt(mean((gam_full_predictions - actual)^2))
    gam_full_mae <- mean(abs(gam_full_predictions - actual))
    gam_full_r2 <- cor(gam_full_predictions, actual)^2
    
    gam_full_rmse_diastolic <- c(gam_full_rmse_diastolic, gam_full_rmse)
    gam_full_mae_diastolic <- c(gam_full_mae_diastolic, gam_full_mae)
    gam_full_r2_diastolic <- c(gam_full_r2_diastolic, gam_full_r2)
    
    #GAM reduced
    gam_reduced_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + Identity + s(SodiumPotassiumRatio), data = data_train)
    gam_reduced_predictions <- predict(gam_reduced_model, newdata = data_test)
    
    gam_reduced_rmse <- sqrt(mean((gam_reduced_predictions - actual)^2))
    gam_reduced_mae <- mean(abs(gam_reduced_predictions - actual))
    gam_reduced_r2 <- cor(gam_reduced_predictions, actual)^2
    
    gam_reduced_rmse_diastolic <- c(gam_reduced_rmse_diastolic, gam_reduced_rmse)
    gam_reduced_mae_diastolic <- c(gam_reduced_mae_diastolic, gam_reduced_mae)
    gam_reduced_r2_diastolic <- c(gam_reduced_r2_diastolic, gam_reduced_r2)
  }
}

mean_diastolic_gam_full_rmse <- mean(gam_full_rmse_diastolic)
mean_diastolic_gam_full_mae <- mean(gam_full_mae_diastolic)
mean_diastolic_gam_full_r2 <- mean(gam_full_r2_diastolic)

mean_diastolic_gam_reduced_rmse <- mean(gam_reduced_rmse_diastolic)
mean_diastolic_gam_reduced_mae <- mean(gam_reduced_mae_diastolic)
mean_diastolic_gam_reduced_r2 <- mean(gam_reduced_r2_diastolic)
```

```{r}
Diastolic_df <- data.frame(
    Model = c("GAM Full", "GAM Reduced"),
    RMSE = c(mean_diastolic_gam_full_rmse, mean_diastolic_gam_reduced_rmse),
    MAE = c(mean_diastolic_gam_full_mae, mean_diastolic_gam_reduced_mae),
    R2 = c(mean_diastolic_gam_full_r2, mean_diastolic_gam_reduced_r2)
)

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 5: GAM Full Model* and Reduced Model** Comparison (Diastolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70, 
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Diastolic)",
        x = 0,
        y = -0.4,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Diastolic)",
        x = 0,
        y = -0.45,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model Full Model* and Reduced Model** Comparison (Diastolic)",
        x = 0,
        y = -0.5,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ), list(
        text = "*Full Model: Model with all risk factors and carb and fiber",
        x = 0,
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 10),
        xanchor = "left"
      ), list(
        text = "**Reduced Model: Model with only risk factors",
        x = 0,
        y = -0.35,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 10),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )

DiastolicFig
```


```{r}
gam_full_model_S <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FibreEnergy) + s(CarbohydrateEnergy) + 
                       Identity + s(SodiumPotassiumRatio),
                     data = normalized_data)
gam_reduced_model_S <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + Identity + s(SodiumPotassiumRatio), data = normalized_data)

full_model_summary <- summary(gam_full_model_S)
reduced_model_summary <- summary(gam_reduced_model_S)

full_model_rsq <- full_model_summary$r.sq
reduced_model_rsq <- reduced_model_summary$r.sq

full_model_dev_expl <- full_model_summary$dev.expl
reduced_model_dev_expl <- reduced_model_summary$dev.expl

full_model_gcv <- full_model_summary$sp.criterion[1]
reduced_model_gcv <- reduced_model_summary$sp.criterion[1]

full_model_params <- full_model_summary$p.table
reduced_model_params <- reduced_model_summary$p.table

full_model_smooth <- full_model_summary$s.table
reduced_model_smooth <- reduced_model_summary$s.table

parametric_full <- data.frame(
  Term = rownames(full_model_params),
  Estimate = full_model_params[, 1],
  Std_Error = full_model_params[, 2],
  T_Value = full_model_params[, 3],
  P_Value = full_model_params[, 4],
  Model = "Full Model"
)

parametric_reduced <- data.frame(
  Term = rownames(reduced_model_params),
  Estimate = reduced_model_params[, 1],
  Std_Error = reduced_model_params[, 2],
  T_Value = reduced_model_params[, 3],
  P_Value = reduced_model_params[, 4],
  Model = "Reduced Model"
)

smooth_full <- data.frame(
  Term = rownames(full_model_smooth),
  EDF = full_model_smooth[, 1],
  Ref_DF = full_model_smooth[, 2],
  F_Statistic = full_model_smooth[, 3],
  P_Value = full_model_smooth[, 4],
  Model = "Full Model"
)

smooth_reduced <- data.frame(
  Term = rownames(reduced_model_smooth),
  EDF = reduced_model_smooth[, 1],
  Ref_DF = reduced_model_smooth[, 2],
  F_Statistic = reduced_model_smooth[, 3],
  P_Value = reduced_model_smooth[, 4],
  Model = "Reduced Model"
)

parametric_full$EDF <- NA
parametric_full$Ref_DF <- NA
parametric_full$F_Statistic <- NA

parametric_reduced$EDF <- NA
parametric_reduced$Ref_DF <- NA
parametric_reduced$F_Statistic <- NA

smooth_full$Estimate <- NA
smooth_full$Std_Error <- NA
smooth_full$T_Value <- NA

smooth_reduced$Estimate <- NA
smooth_reduced$Std_Error <- NA
smooth_reduced$T_Value <- NA

parametric_full <- parametric_full[, c("Term", "Estimate", "Std_Error", "T_Value", "P_Value", "EDF", "Ref_DF", "F_Statistic", "Model")]
parametric_reduced <- parametric_reduced[, c("Term", "Estimate", "Std_Error", "T_Value", "P_Value", "EDF", "Ref_DF", "F_Statistic", "Model")]
smooth_full <- smooth_full[, c("Term", "Estimate", "Std_Error", "T_Value", "P_Value", "EDF", "Ref_DF", "F_Statistic", "Model")]
smooth_reduced <- smooth_reduced[, c("Term", "Estimate", "Std_Error", "T_Value", "P_Value", "EDF", "Ref_DF", "F_Statistic", "Model")]

model_results <- rbind(parametric_full, parametric_reduced, smooth_full, smooth_reduced)

full_model_results <- model_results[model_results$Model == "Full Model", ]
reduced_model_results <- model_results[model_results$Model == "Reduced Model", ]

full_model_results %>%
  gt() %>%
  tab_header(
    title = "GAM Model Summary: Full Model"
  ) %>%
  fmt_number(columns = c(Estimate, Std_Error, T_Value, P_Value, EDF, Ref_DF, F_Statistic), decimals = 3) %>%
  cols_label(
    Term = "Variable",
    Estimate = "Estimate",
    Std_Error = "Std Error",
    T_Value = "T-Value",
    P_Value = "P-Value",
    EDF = "EDF",
    Ref_DF = "Reference DF",
    F_Statistic = "F-Statistic"
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.align = "center"
  )

reduced_model_results %>%
  gt() %>%
  tab_header(
    title = "GAM Model Summary: Reduced Model"
  ) %>%
  fmt_number(columns = c(Estimate, Std_Error, T_Value, P_Value, EDF, Ref_DF, F_Statistic), decimals = 3) %>%
  cols_label(
    Term = "Variable",
    Estimate = "Estimate",
    Std_Error = "Std Error",
    T_Value = "T-Value",
    P_Value = "P-Value",
    EDF = "EDF",
    Ref_DF = "Reference DF",
    F_Statistic = "F-Statistic"
  ) %>%
  tab_options(
    table.font.size = "medium",
    heading.align = "center"
  )
```


# Stratify

## Function for CV

```{r function for systiolic cv}
k <- 5
repeats <- 3

sys_cv <- function(data) {
  n <- nrow(data)

  folds <- map(1:repeats, ~ sample(rep(1:k, length.out = n)))
  
  results <- map_dfr(folds, function(fold) {
    map_dfr(1:k, function(i) {
      train_idx <- which(fold != i)
      test_idx <- which(fold == i)
      
      data_train <- data[train_idx, ]
      data_test <- data[test_idx, ]
      
      actual <- data_test$Systolic
      
      gam_model <- gam(Systolic ~ s(FibreEnergy) + s(CarbohydrateEnergy), data = data_train)
      
      gam_predictions <- predict(gam_model, newdata = data_test)
      
      tibble(
        rmse = sqrt(mean((gam_predictions - actual)^2)),
        mae = mean(abs(gam_predictions - actual)),
        r2 = cor(gam_predictions, actual)^2
      )
    })
  })

  summary_results <- results %>%
    summarize(
      mean_rmse = mean(rmse),
      mean_mae = mean(mae),
      mean_r2 = mean(r2)
    )
  
  return(summary_results)
}
```

```{r function for diastolic cv}
dia_cv <- function(data) {
  n <- nrow(data)

  folds <- map(1:repeats, ~ sample(rep(1:k, length.out = n)))
  
  results <- map_dfr(folds, function(fold) {
    map_dfr(1:k, function(i) {
      train_idx <- which(fold != i)
      test_idx <- which(fold == i)
      
      data_train <- data[train_idx, ]
      data_test <- data[test_idx, ]
      
      actual <- data_test$Diastolic
      
      gam_model <- gam(Diastolic ~ s(FibreEnergy) + s(CarbohydrateEnergy), data = data_train)
      
      gam_predictions <- predict(gam_model, newdata = data_test)
      
      tibble(
        rmse = sqrt(mean((gam_predictions - actual)^2)),
        mae = mean(abs(gam_predictions - actual)),
        r2 = cor(gam_predictions, actual)^2
      )
    })
  })

  summary_results <- results %>%
    summarize(
      mean_rmse = mean(rmse),
      mean_mae = mean(mae),
      mean_r2 = mean(r2)
    )
  
  return(summary_results)
}
```

## By Age
### Group
```{r age group, warning = FALSE, message=FALSE}
age1 <- selected_data %>% filter(Age < 35) %>% select_if(is.numeric)

age2 <- selected_data %>%
  filter(Age >= 35 & Age < 65) %>% select_if(is.numeric)

age3 <- selected_data %>% filter(Age >= 65) %>% select_if(is.numeric)
```

### Systolic
```{r age systolic result}
set.seed(3888)
results_age1 <- sys_cv(age1)
results_age2 <- sys_cv(age2)
results_age3 <- sys_cv(age3)

gam_rmse_diastolic_ag1 <- results_age1$mean_rmse
gam_mae_diastolic_ag1 <- results_age1$mean_mae
gam_r2_diastolic_ag1 <- results_age1$mean_r2

gam_rmse_diastolic_ag2 <- results_age2$mean_rmse
gam_mae_diastolic_ag2 <- results_age2$mean_mae
gam_r2_diastolic_ag2 <- results_age2$mean_r2

gam_rmse_diastolic_ag3 <- results_age3$mean_rmse
gam_mae_diastolic_ag3 <- results_age3$mean_mae
gam_r2_diastolic_ag3 <- results_age3$mean_r2
```

```{r age systolic(Figure8), fig.width=10}
Systolic_df <- data.frame(
    Model = c("18-34", "35-64","65-85"),
    RMSE = c(gam_rmse_diastolic_ag1, gam_rmse_diastolic_ag2, gam_rmse_diastolic_ag3),
    MAE = c(gam_mae_diastolic_ag1, gam_mae_diastolic_ag2, gam_mae_diastolic_ag3),
    R2 = c(gam_r2_diastolic_ag1, gam_r2_diastolic_ag2, gam_r2_diastolic_ag3)
)

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

AgeSystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 6: Different Age Group GAM Comparison (Systolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for Generalized Additive Model",
        x = 0,
        y = -0.2,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model",
        x = 0,
        y = -0.25,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model",
        x = 0,
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),list(
        text = "Age Groups",
        y = -0.15,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "center"
      )
    ),
    margin = list(b = 140, t = 80)
  )

AgeSystolicFig
```

### Diastolic

```{r age diastolic result}
set.seed(3888)
results_age1 <- dia_cv(age1)
results_age2 <- dia_cv(age2)
results_age3 <- dia_cv(age3)

gam_rmse_diastolic_ag1 <- results_age1$mean_rmse
gam_mae_diastolic_ag1 <- results_age1$mean_mae
gam_r2_diastolic_ag1 <- results_age1$mean_r2

gam_rmse_diastolic_ag2 <- results_age2$mean_rmse
gam_mae_diastolic_ag2 <- results_age2$mean_mae
gam_r2_diastolic_ag2 <- results_age2$mean_r2

gam_rmse_diastolic_ag3 <- results_age3$mean_rmse
gam_mae_diastolic_ag3 <- results_age3$mean_mae
gam_r2_diastolic_ag3 <- results_age3$mean_r2
```


```{r age diastolic(Figure9), fig.width=10}
Diastolic_df <- data.frame(
    Model = c("18-34", "35-64","65-85"),
    RMSE = c(gam_rmse_diastolic_ag1, gam_rmse_diastolic_ag2, gam_rmse_diastolic_ag3),
    MAE = c(gam_mae_diastolic_ag1, gam_mae_diastolic_ag2, gam_mae_diastolic_ag3),
    R2 = c(gam_r2_diastolic_ag1, gam_r2_diastolic_ag2, gam_r2_diastolic_ag3)
)

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 7: Different Age Group GAM Comparison (Diastolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for Generalized Additive Model",
        x = 0,
        y = -0.2,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for Generalized Additive Model",
        x = 0,
        y = -0.25,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for Generalized Additive Model",
        x = 0,
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),list(
        text = "Age Groups",
        y = -0.15,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "center"
      )
    ),
    margin = list(b = 140, t = 80)
  )

DiastolicFig
```

### Partial effect plots - align with the model summary

```{r age model}
age_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age1)
age_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age2)
age_model_3 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age3)

age_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age1)
age_model_5 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age2)
age_model_6 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = age3)

```

#### FibreEnergy
```{r age Fibreenergy(Figure10), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2), oma = c(7, 0, 2, 0))

Fibre_color <- "blue"

plot_labels <- c("A", "B", "C", "D", "E", "F")

plot(age_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[1]))
plot(age_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[2]))
plot(age_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[3]))
plot(age_model_5, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[4]))
plot(age_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[5]))
plot(age_model_6, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[6]))

mtext("Figure 8: Relationship between FibreEnergy and Blood Pressure by Age Group", outer = TRUE, cex = 1.2)

mtext("A. 18-34 Age Group - Systolic - Fibreenergy\nB. 18-34 Age Group - Diastolic - Fibreenergy\nC. 35-64 Age Group - Systolic - Fibreenergy\nD. 35-64 Age Group - Diastolic - Fibreenergy\nE. 65-85 Age Group - Systolic - Fibreenergy\nF. 65-85 Age Group - Diastolic - Fibreenergy", side = 1, adj = 0.05, line = 5, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r age carbohydrateenergy(Figure11), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2), oma = c(7, 0, 2, 0))

carb_color <- "red"

plot(age_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "A")
plot(age_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "B")
plot(age_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "C")
plot(age_model_5, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "D")
plot(age_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "E")
plot(age_model_6, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "F")

mtext("Figure 9: Relationship between FibreEnergy and Blood Pressure by Age Group", outer = TRUE, cex = 1.2)

mtext("A. 18-34 Age Group - Systolic - CarbohydrateEnergy\nB. 18-24 Age Group - Diastolic - CarbohydrateEnergy\nC. 36-64 Age Group - Systolic - CarbohydrateEnergy\nD. 35-64 Age Group - Diastolic - CarbohydrateEnergy\nE. 65-85 Age Group - Systolic - CarbohydrateEnergy\nF. 65-85 Age Group - Diastolic - CarbohydrateEnergy", side = 1, adj = 0.05, line = 5, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

## By BMI
### Group
```{r bmi group}
light <- selected_data %>% filter(BMI < 24.9)
normal <- selected_data %>% filter(BMI >= 24.9 & BMI <= 29.9)
heavy <- selected_data %>% filter(BMI > 29.9)
```

### Systolic
```{r bmi result}
set.seed(3888)
results_light <- sys_cv(light)
results_normal <- sys_cv(normal)
results_heavy <- sys_cv(heavy)

gam_rmse_systolic_light <- results_light$mean_rmse
gam_mae_systolic_light <- results_light$mean_mae
gam_r2_systolic_light <- results_light$mean_r2

gam_rmse_systolic_normal <- results_normal$mean_rmse
gam_mae_systolic_normal <- results_normal$mean_mae
gam_r2_systolic_normal <- results_normal$mean_r2

gam_rmse_systolic_heavy <- results_heavy$mean_rmse
gam_mae_systolic_heavy <- results_heavy$mean_mae
gam_r2_systolic_heavy <- results_heavy$mean_r2
```

```{r bmi result(Figure12), fig.width=10}
Systolic_df <- data.frame(
    Group = c("Light BMI", "Normal BMI", "Heavy BMI"),
    RMSE = c(gam_rmse_systolic_light, gam_rmse_systolic_normal, gam_rmse_systolic_heavy),
    MAE = c(gam_mae_systolic_light, gam_mae_systolic_normal, gam_mae_systolic_heavy),
    R2 = c(gam_r2_systolic_light, gam_r2_systolic_normal, gam_r2_systolic_heavy)
)

rmse_range <- c(min(Systolic_df$RMSE) * 0.95, max(Systolic_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_df$MAE) * 0.95, max(Systolic_df$MAE) * 1.05)
r2_range <- c(min(Systolic_df$R2) * 0.95, max(Systolic_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 10: Differen BMI Group GAM Comparison (Systolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.10,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.10,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.10,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for models",
        x = 0,
        y = -0.30,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0,
        y = -0.35,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0,
        y = -0.40,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )

SystolicFig
```

### Diastolic
```{r bmi diastolic result}
set.seed(3888)
results_light_dias <- dia_cv(light)
results_normal_dias <- dia_cv(normal)
results_heavy_dias <- dia_cv(heavy)

gam_rmse_diastolic_light <- results_light_dias$mean_rmse
gam_mae_diastolic_light <- results_light_dias$mean_mae
gam_r2_diastolic_light <- results_light_dias$mean_r2

gam_rmse_diastolic_normal <- results_normal_dias$mean_rmse
gam_mae_diastolic_normal <- results_normal_dias$mean_mae
gam_r2_diastolic_normal <- results_normal_dias$mean_r2

gam_rmse_diastolic_heavy <- results_heavy_dias$mean_rmse
gam_mae_diastolic_heavy <- results_heavy_dias$mean_mae
gam_r2_diastolic_heavy <- results_heavy_dias$mean_r2
```

```{r bmi diastolic(Figure13), fig.width=10}
Diastolic_df <- data.frame(
    Group = c("Light BMI", "Normal BMI", "Heavy BMI"),
    RMSE = c(gam_rmse_diastolic_light, gam_rmse_diastolic_normal, gam_rmse_diastolic_heavy),
    MAE = c(gam_mae_diastolic_light, gam_mae_diastolic_normal, gam_mae_diastolic_heavy),
    R2 = c(gam_r2_diastolic_light, gam_r2_diastolic_normal, gam_r2_diastolic_heavy)
)

rmse_range <- c(min(Diastolic_df$RMSE) * 0.95, max(Diastolic_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_df$MAE) * 0.95, max(Diastolic_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_df$R2) * 0.95, max(Diastolic_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae <- plot_ly(Diastolic_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2 <- plot_ly(Diastolic_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 11: Model Comparison (Diastolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(
        text = "A",
        x = 0,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "B",
        x = 0.35,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for models",
        x = 0, # Align to the far left
        y = -0.30,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0, # Same x as the first caption
        y = -0.35, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0, # Same x as the first caption
        y = -0.40, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80) # Adjust margins as needed
  )
DiastolicFig
```

### Partial effect plots - align with the model summary

```{r bmi model}
bmi_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = light)
bmi_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = normal)
bmi_model_3 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = heavy)

bmi_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = light)
bmi_model_5 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = normal)
bmi_model_6 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = heavy)

```

#### FibreEnergy
```{r age Fibreenergy(Figure14), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2), oma = c(7, 0, 2, 0))

plot(bmi_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "A")
plot(bmi_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "B")
plot(bmi_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "C")
plot(bmi_model_5, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "D")
plot(bmi_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "E")
plot(bmi_model_6, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = "F")

mtext("Figure 12: Relationship between FibreEnergy and Blood Pressure by BMI Group", outer = TRUE, cex = 1.2)

mtext("A. 16-24.9 BMI Group - Systolic - FibreEnergy\nB. 16-24.9 BMI Group - Diastolic - FibreEnergy\nC. 24.9-29.9 BMI Group - Systolic - FibreEnergy\nD. 24.9-29.9 BMI Group - Diastolic - FibreEnergy\nE. 29.9-47.5 BMI Group - Systolic - FibreEnergy\nF. 29.9-47.5 BMI Group - Diastolic - FibreEnergy",side = 1, adj = 0.05, line = 5, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r age carbohydrateenergy(Figure15), fig.width= 12, fig.height=14}
par(mfrow = c(3, 2), oma = c(7, 0, 2, 0))

plot(bmi_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "A")
plot(bmi_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "B")
plot(bmi_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "C")
plot(bmi_model_5, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "D")
plot(bmi_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "E")
plot(bmi_model_6, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = "F")

mtext("Figure 13: Relationship between CarbohydrateEnergy and Blood Pressure by BMI Group", outer = TRUE, cex = 1.2)

mtext("A. 16-24.9 BMI Group - Systolic - CarbohydrateEnergy\nB. 16-24.9 BMI Group - Diastolic - CarbohydrateEnergy\nC. 24.9-29.9 BMI Group - Systolic - CarbohydrateEnergy\nD. 24.9-29.9 BMI Group - Diastolic - CarbohydrateEnergy\nE. 29.9-47.5 BMI Group - Systolic - CarbohydrateEnergy\nF. 29.9-47.5 BMI Group - Diastolic - CarbohydrateEnergy", side = 1, adj = 0.05, line = 5, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

## By AlcoholPercentage
### Group
```{r alcohol group}
nonalcohol <- selected_data %>% filter(AlcoholPercentage == 0)

alcohol <- selected_data %>% filter(AlcoholPercentage > 0)
```

### Systolic

```{r alcohol systolic result}
set.seed(3888)
results_nonalcohol <- sys_cv(nonalcohol)
results_alcohol <- sys_cv(alcohol)

gam_rmse_systolic_nonalcohol <- results_nonalcohol$mean_rmse
gam_mae_systolic_nonalcohol <- results_nonalcohol$mean_mae
gam_r2_systolic_nonalcohol <- results_nonalcohol$mean_r2

gam_rmse_systolic_alcohol <- results_alcohol$mean_rmse
gam_mae_systolic_alcohol <- results_alcohol$mean_mae
gam_r2_systolic_alcohol <- results_alcohol$mean_r2
```

```{r alcohol systolic result(Figure16), fig.width=10}
Systolic_alcohol_df <- data.frame(
    Group = c("Non-Alcohol", "Alcohol"),
    RMSE = c(gam_rmse_systolic_nonalcohol, gam_rmse_systolic_alcohol),
    MAE = c(gam_mae_systolic_nonalcohol, gam_mae_systolic_alcohol),
    R2 = c(gam_r2_systolic_nonalcohol, gam_r2_systolic_alcohol)
)

rmse_range_alcohol <- c(min(Systolic_alcohol_df$RMSE) * 0.95, max(Systolic_alcohol_df$RMSE) * 1.05)
mae_range_alcohol <- c(min(Systolic_alcohol_df$MAE) * 0.95, max(Systolic_alcohol_df$MAE) * 1.05)
r2_range_alcohol <- c(min(Systolic_alcohol_df$R2) * 0.95, max(Systolic_alcohol_df$R2) * 1.05)

Systolic_rmse_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2_alcohol <- plot_ly(Systolic_alcohol_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

SystolicFig_alcohol <- subplot(Systolic_rmse_alcohol, Systolic_mae_alcohol, Systolic_r2_alcohol, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 14: Alcohol Group Model Comparison (Systolic)",
      font = list(size = 16), # Slightly larger title font
      x = 0.5,
      y = 1.15  # Move the title further from the subplots
    ),
    annotations = list(
      # "A" label to the left of RMSE subplot, near y-axis
      list(
        text = "A",
        x = 0, # Position it to the left of the first subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "B" label to the left of MAE subplot, near y-axis
      list(
        text = "B",
        x = 0.35, # Position it to the left of the second subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "C" label to the left of R² subplot, near y-axis
      list(
        text = "C",
        x = 0.70, # Position it to the left of the third subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # Captions grouped together at the bottom-left
      list(
        text = "A. RMSE comparison for models",
        x = 0, # Align to the far left
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0, # Same x as the first caption
        y = -0.35, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0, # Same x as the first caption
        y = -0.4, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80) # Adjust margins as needed
  )

SystolicFig_alcohol
```

### Diastolic
```{r alcohol diastolic result}
set.seed(3888)
results_nonalcohol_dias <- dia_cv(nonalcohol)
results_alcohol_dias <- dia_cv(alcohol)

gam_rmse_diastolic_nonalcohol <- results_nonalcohol_dias$mean_rmse
gam_mae_diastolic_nonalcohol <- results_nonalcohol_dias$mean_mae
gam_r2_diastolic_nonalcohol <- results_nonalcohol_dias$mean_r2

gam_rmse_diastolic_alcohol <- results_alcohol_dias$mean_rmse
gam_mae_diastolic_alcohol <- results_alcohol_dias$mean_mae
gam_r2_diastolic_alcohol <- results_alcohol_dias$mean_r2
```

```{r alcohol diastolic result(Figure17), fig.width=10}
Diastolic_alcohol_df <- data.frame(
    Group = c("Non-Alcohol", "Alcohol"),
    RMSE = c(gam_rmse_diastolic_nonalcohol, gam_rmse_diastolic_alcohol),
    MAE = c(gam_mae_diastolic_nonalcohol, gam_mae_diastolic_alcohol),
    R2 = c(gam_r2_diastolic_nonalcohol, gam_r2_diastolic_alcohol)
)

rmse_range_alcohol <- c(min(Diastolic_alcohol_df$RMSE) * 0.95, max(Diastolic_alcohol_df$RMSE) * 1.05)
mae_range_alcohol <- c(min(Diastolic_alcohol_df$MAE) * 0.95, max(Diastolic_alcohol_df$MAE) * 1.05)
r2_range_alcohol <- c(min(Diastolic_alcohol_df$R2) * 0.95, max(Diastolic_alcohol_df$R2) * 1.05)

Diastolic_rmse_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(RMSE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_mae_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(MAE, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Diastolic_r2_alcohol <- plot_ly(Diastolic_alcohol_df, x = ~Group, y = ~round(R2, 3), type = 'bar', color = ~Group, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range_alcohol),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

DiastolicFig_alcohol <- subplot(Diastolic_rmse_alcohol, Diastolic_mae_alcohol, Diastolic_r2_alcohol, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 15: Model Comparison (Diastolic)",
      font = list(size = 16), # Slightly larger title font
      x = 0.5,
      y = 1.15  # Move the title further from the subplots
    ),
    annotations = list(
      # "A" label to the left of RMSE subplot, near y-axis
      list(
        text = "A",
        x = 0, # Position it to the left of the first subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "B" label to the left of MAE subplot, near y-axis
      list(
        text = "B",
        x = 0.35, # Position it to the left of the second subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "C" label to the left of R² subplot, near y-axis
      list(
        text = "C",
        x = 0.70, # Position it to the left of the third subplot
        y = 1.05,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # Captions grouped together at the bottom-left
      list(
        text = "A. RMSE comparison for models",
        x = 0, # Align to the far left
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0, # Same x as the first caption
        y = -0.35, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0, # Same x as the first caption
        y = -0.40, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80) # Adjust margins as needed
  )

DiastolicFig_alcohol
```

### Partial effect plots - align with the model summary

```{r alcohol model}
alc_model_1 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = nonalcohol)
alc_model_2 <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = alcohol)
alc_model_3 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = nonalcohol)
alc_model_4 <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = alcohol)
```

#### FibreEnergy
```{r alcohol Fibreenergy, fig.width= 12, fig.height=10}
par(mfrow = c(2, 2), oma = c(4, 0, 2, 0))


plot_labels <- c("A", "B", "C", "D")

plot(alc_model_1, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[1]))
plot(alc_model_3, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[2]))
plot(alc_model_2, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[3]))
plot(alc_model_4, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[4]))

mtext("Figure 16: Relationship between FibreEnergy and Blood Pressure by Alcohol Group", outer = TRUE, cex = 1.2)

mtext("A. Non-alcohol Group - Systolic - FibreEnergy\nB. Non-alcohol Group - Diastolic - FibreEnergy\nC. Alcohol Group - Systolic - FibreEnergy\nD. Alcohol Group - Diastolic - FibreEnergy", side = 1, adj = 0.05, line = 2, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r alcohol carbohydrateenergy, fig.width= 12, fig.height=10}
par(mfrow = c(2, 2), oma = c(4, 0, 2, 0))


plot_labels <- c("A", "B", "C", "D")

plot(alc_model_1, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[1]))
plot(alc_model_3, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[2]))
plot(alc_model_2, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[3]))
plot(alc_model_4, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[4]))

mtext("Figure 17: Relationship between CarbohydrateEnergy and Blood Pressure by Alcohol Group", outer = TRUE, cex = 1.2)

mtext("A. Non-alcohol Group - Systolic - CarbohydrateEnergy\nB. Non-alcohol Group - Diastolic - CarbohydrateEnergy\nC. Alcohol Group - Systolic - CarbohydrateEnergy\nD. Alcohol Group - Diastolic - CarbohydrateEnergy", side = 1, adj = 0.05, line = 2, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

## By Gender
### Group
```{r gender group, warning = FALSE, message=FALSE}
male <- selected_data %>% filter(Gender == 1) %>% select_if(is.numeric)

female <- selected_data %>% filter(Gender == 2) %>% select_if(is.numeric)
```


### Systolic
```{r gender systolic}
set.seed(3888)
results_male <- sys_cv(male)
results_female <- sys_cv(female)

gam_rmse_diastolic_male <- results_male$mean_rmse
gam_mae_diastolic_male <- results_male$mean_mae
gam_r2_diastolic_male <- results_male$mean_r2

gam_rmse_diastolic_female <- results_female$mean_rmse
gam_mae_diastolic_female <- results_female$mean_mae
gam_r2_diastolic_female <- results_female$mean_r2
```

```{r gender systolic result(Figure20), fig.width=10}
Systolic_gender_df <- data.frame(
    Model = c("Male", "Female"),
    RMSE = c(gam_rmse_diastolic_male, gam_rmse_diastolic_female),
    MAE = c(gam_mae_diastolic_male, gam_mae_diastolic_female),
    R2 = c(gam_r2_diastolic_male, gam_r2_diastolic_female)
)

rmse_range <- c(min(Systolic_gender_df$RMSE) * 0.95, max(Systolic_gender_df$RMSE) * 1.05)
mae_range <- c(min(Systolic_gender_df$MAE) * 0.95, max(Systolic_gender_df$MAE) * 1.05)
r2_range <- c(min(Systolic_gender_df$R2) * 0.95, max(Systolic_gender_df$R2) * 1.05)

Systolic_rmse <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_mae <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

Systolic_r2 <- plot_ly(Systolic_gender_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)')

GenderSystolicFig <- subplot(Systolic_rmse, Systolic_mae, Systolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 18: Gender Group Model Comparison (Systolic)",
      font = list(size = 16), # Slightly larger title font
      x = 0.5,
      y = 1.15  # Move the title further from the subplots
    ),
    annotations = list(
      # "A" label to the left of RMSE subplot, near y-axis
      list(
        text = "A",
        x = 0, # Position it to the left of the first subplot
        y = 1.125,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "B" label to the left of MAE subplot, near y-axis
      list(
        text = "B",
        x = 0.35, # Position it to the left of the second subplot
        y = 1.125,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # "C" label to the left of R² subplot, near y-axis
      list(
        text = "C",
        x = 0.70, # Position it to the left of the third subplot
        y = 1.125,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      # Captions grouped together at the bottom-left
      list(
        text = "A. RMSE comparison for models",
        x = 0, # Align to the far left
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0, # Same x as the first caption
        y = -0.35, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0, # Same x as the first caption
        y = -0.40, # Slightly lower for spacing
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80) # Adjust margins as needed
  )

GenderSystolicFig
```

### Diastolic
```{r gender diastolic result}
set.seed(3888)

results_male <- dia_cv(male)
results_female <- dia_cv(female)

gam_rmse_diastolic_male <- results_male$mean_rmse
gam_mae_diastolic_male <- results_male$mean_mae
gam_r2_diastolic_male <- results_male$mean_r2

gam_rmse_diastolic_female <- results_female$mean_rmse
gam_mae_diastolic_female <- results_female$mean_mae
gam_r2_diastolic_female <- results_female$mean_r2
```


```{r gender diastolic result(Figure21), fig.width=10}
Diastolic_gender_df <- data.frame(
    Model = c("Male", "Female"),
    RMSE = c(gam_rmse_diastolic_male, gam_rmse_diastolic_female),
    MAE = c(gam_mae_diastolic_male, gam_mae_diastolic_female),
    R2 = c(gam_r2_diastolic_male, gam_r2_diastolic_female)
)

rmse_range <- c(min(Diastolic_gender_df$RMSE) * 0.95, max(Diastolic_gender_df$RMSE) * 1.05)
mae_range <- c(min(Diastolic_gender_df$MAE) * 0.95, max(Diastolic_gender_df$MAE) * 1.05)
r2_range <- c(min(Diastolic_gender_df$R2) * 0.95, max(Diastolic_gender_df$R2) * 1.05)

Diastolic_rmse <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'RMSE', range = rmse_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_mae <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'MAE', range = mae_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

Diastolic_r2 <- plot_ly(Diastolic_gender_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    yaxis = list(title = 'R²', range = r2_range),
    showlegend = FALSE,
    paper_bgcolor = 'rgba(245, 246, 249, 1)',
    plot_bgcolor = 'rgba(0, 0, 0, 0)'
  )

DiastolicFig <- subplot(Diastolic_rmse, Diastolic_mae, Diastolic_r2, nrows = 1, shareY = FALSE) %>%
  layout(
    title = list(
      text = "Figure 19: Gender Group Model Comparison (Diastolic)",
      font = list(size = 16),
      x = 0.5,
      y = 1.15
    ),
    annotations = list(
      list(text = "A", x = 0, y = 1.1, xref = "paper", yref = "paper", showarrow = FALSE, font = list(size = 14, color = "black"), xanchor = "center"),
      list(
        text = "B",
        x = 0.35,
        y = 1.1,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "C",
        x = 0.70,
        y = 1.1,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 14, color = "black"),
        xanchor = "center"
      ),
      list(
        text = "A. RMSE comparison for models",
        x = 0,
        y = -0.3,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "B. MAE comparison for models",
        x = 0,
        y = -0.35,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      ),
      list(
        text = "C. R² comparison for models",
        x = 0,
        y = -0.40,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 12),
        xanchor = "left"
      )
    ),
    margin = list(b = 140, t = 80)
  )
DiastolicFig
```

### Partial effect plots - align with the model summary

```{r gender model}
systolic_male <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = male)
systolic_female <- gam(Systolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = female)

diastolic_male <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = male)
diastolic_female <- gam(Diastolic ~ s(CarbohydrateEnergy) + s(FibreEnergy), data = female)
```

#### FibreEnergy
```{r gender Fibreenergy(Figure22), fig.width= 12, fig.height=10}
par(mfrow = c(2, 2), oma = c(4, 0, 2, 0))


plot_labels <- c("A", "B", "C", "D")

plot(systolic_male, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[1]))
plot(diastolic_male, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[2]))
plot(systolic_female, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[3]))
plot(diastolic_female, select = 2, seWithMean = TRUE, shade = TRUE, col = Fibre_color, main = paste(plot_labels[4]))

mtext("Figure 20: Relationship between FibreEnergy and Blood pressure by Gender Group", outer = TRUE, cex = 1.2)

mtext("A. Male Group - Systolic - FibreEnergy\nB. Male Group - Diastolic - FibreEnergy\nC. Female Group - Systolic - FibreEnergy\nD. Female Group - Diastolic - FibreEnergy", side = 1, adj = 0.05, line = 2, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))
```

#### CarbohydrateEnergy
```{r gender carbohydrateenergy(Figure23), fig.width= 12, fig.height=10}
par(mfrow = c(2, 2), oma = c(4, 0, 2, 0))


plot_labels <- c("A", "B", "C", "D")

plot(systolic_male, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[1]))
plot(diastolic_male, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[2]))
plot(systolic_female, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[3]))
plot(diastolic_female, select = 1, seWithMean = TRUE, shade = TRUE, col = carb_color, main = paste(plot_labels[4]))

mtext("Figure 21: Relationship between CarbohydrateEnergy and Blood pressure by Gender Group", outer = TRUE, cex = 1.2)

mtext("A. Male Group - Systolic - CarbohydrateEnergy\nB. Male Group - Diastolic - CarbohydrateEnergy\nC. Female Group - Systolic - CarbohydrateEnergy\nD. Female Group - Diastolic - CarbohydrateEnergy", side = 1, adj = 0.05, line = 2, outer = TRUE, cex = 0.7)

par(mfrow = c(1, 1))

```