---
title: "Final_data_clean"
author: "Group2"
date: "2024-08-21"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
table-of-contents: true 
number-sections: true 
embed-resources: true 
editor_options: 
  markdown: 
    wrap: 72
---


# Library
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)      
library(readxl)    
library(janitor)   
library(stringr)   
library(tidyr)     
library(dplyr)
library(knitr)
library(ggplot2)
library(factoextra)
library(gt)
library(naniar)
library(ggcorrplot)
library(corrplot)
library(GGally)
library(randomForest)
library(pdp)
library(e1071)  
library(dplyr) 
library(caret)
library(plotly)
library(stargazer)
library(broom)
library(knitr)
library(kableExtra)
library(gridExtra)
library(patchwork)
```
# Data Import
```{r data import}
raw_bp <- read.csv(here("ZIPallFiles","AHSnpa11bp.csv"), header=TRUE)
raw_bb <- read.csv(here("ZIPallFiles","AHSnpa11bb.csv"), header=TRUE)
raw_bsp <- read.csv(here("ZIPallFiles","inp13bsp.csv"), header=TRUE)
raw_bcn <- read.csv(here("ZIPallFiles","inp13bcn.csv"), header=TRUE)
raw_bhh <- read.csv(here("ZIPallFiles","inp13bhh.csv"), header=TRUE)
```
# Rename and Created
```{r rename+Identity}
bp_bb <- inner_join(raw_bp, raw_bb, by = "ABSPID") %>%
  rename(HypertensionStatus = HYPBC,
         BodyMass = SABDYMS,
         SocialStatus = SF2SA1QN,
         PersonID = ABSPID,
         Age = AGEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2) %>%
   mutate(Identity = factor(0))

bsp_bhh<- left_join(raw_bsp, raw_bhh,by = c("ABSHID")) %>%
  rename(BodyMass = SABDYMS,
         SocialStatus = SF2SA1DB,
         HouseID = ABSHID,
         Age = AGEEC,
         Gender = SEX,
         BMI = BMISC,
         Systolic = SYSTOL,
         Diastolic = DIASTOL,
         SmokeStatus = SMKSTAT,
         AlcoholPercentage_Day1 = ALCPER1,
         AlcoholPercentage_Day2 = ALCPER2,
         Potassium_Day1 = POTAST1,
         Potassium_Day2 = POTAST2,
         Sodium_Day1 = SODIUMT1,
         Sodium_Day2 = SODIUMT2,
         FiberEnergy_Day1 = FIBRPER1,
         FiberEnergy_Day2 = FIBRPER2,
         CarbonhydrateEnergy_Day1 = CHOPER1,
         CarbonhydrateEnergy_Day2 = CHOPER2,
         EnergyBMR_Day1 = EIBMR1,
         EnergyBMR_Day2 = EIBMR2)%>%
   mutate(Identity = factor(1))

bcn <- raw_bcn %>%
  rename(HouseID = ABSHID,
         MedicalCondition = ICD10ME)
```
# First Filter
```{r filter 1}
bp_bb <- bp_bb %>%
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    HypertensionStatus == 5,
    BodyMass != 4,
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))

bsp_bhh <- bsp_bhh %>% 
  mutate(EnergyBMR_Day1 = na_if(EnergyBMR_Day1, 997) %>% na_if(998),
         EnergyBMR_Day2 = na_if(EnergyBMR_Day2, 997) %>% na_if(998),
         EnergyBMR = (EnergyBMR_Day1 + EnergyBMR_Day2)/2) %>%
  filter(
    Age >= 18, 
    BodyMass != 4,
    !HouseID %in% bcn$HouseID[bcn$MedicalCondition %in% c(7, 20)],
    (is.na(EnergyBMR) | EnergyBMR >= 0.9))
```
# Merge
```{r merge}
raw_data <- bind_rows(bp_bb,bsp_bhh)
```
# Variable Selection
```{r variable selection}
selected_data <- raw_data %>%
select(
    PersonID, HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, Identity,
    AlcoholPercentage_Day1, AlcoholPercentage_Day2,
    Potassium_Day1, Potassium_Day2, 
    Sodium_Day1, Sodium_Day2,
    FiberEnergy_Day1, FiberEnergy_Day2, 
    CarbonhydrateEnergy_Day1, CarbonhydrateEnergy_Day2
  )
```



# Type Check 1
```{r variable check before}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 1: Variable types table"))
```
# Na values + New Variables + Type Conversion
```{r navalue_new_typeconversion}
selected_data <- selected_data %>%
  mutate(
    across(c(Gender, SocialStatus, SmokeStatus), as.factor),
    BMI = na_if(BMI, 98) %>% na_if(99),
    Systolic = na_if(Systolic, 0) %>% na_if(998) %>% na_if(999),
    Diastolic = na_if(Diastolic, 0) %>% na_if(998) %>% na_if(999),
    Potassium = (Potassium_Day1 + Potassium_Day2) / 2, 
    Sodium = (Sodium_Day1 + Sodium_Day2) / 2,
    FiberEnergy = (FiberEnergy_Day1 + FiberEnergy_Day2)/2,
    CarbonhydrateEnergy = (CarbonhydrateEnergy_Day1 + CarbonhydrateEnergy_Day2) / 2,
    AlcoholPercentage = (AlcoholPercentage_Day1 + AlcoholPercentage_Day2) /2,
    PotassiumSodiumRatio = Sodium / Potassium
    ) %>% 
  select(PersonID,HouseID, SocialStatus, Age, Gender, BMI, Systolic, Diastolic, SmokeStatus, AlcoholPercentage, PotassiumSodiumRatio, FiberEnergy, CarbonhydrateEnergy, Identity)
```
# Type Check 2
```{r variable check after}
data.frame(Variable = names(selected_data), Type = sapply(selected_data, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
 tab_caption(caption = md("Table 2: Variable types table after correction"))
```
# Duplicate Values
```{r duplicate values}
selected_data %>% 
  select(-PersonID) %>% 
  distinct() %>% 
  dim() %>% 
  tibble::tibble(Dimension = c("Rows", "Columns"), Count = .) %>% 
  gt() %>%
  tab_header(title = "Dimensions of Unique Tibble") %>%
  tab_caption(caption = md("Table 3: Dimensions of Unique Data Table"))
```

# Categrocial Distribution
```{r categrocial distribution bar plot}
plot_categorical_distribution <- function(data) {
  categorical_vars <- data %>%
    select(where(is.factor) | where(is.character)) %>%
    select(-PersonID, -HouseID)
  
  for (var in names(categorical_vars)) {
    non_na_data <- data %>%
      filter(!is.na(.data[[var]]))
    
    if (nrow(non_na_data) == 0) next
    
    p <- ggplot(non_na_data, aes(x = .data[[var]])) +
      geom_bar(aes(y = (..count..) / sum(..count..), fill = .data[[var]]), color = "black") +
      scale_fill_brewer(palette = "Set3") + 
      labs(title = paste("Proportion of Categories in", var), 
           x = "Category", 
           y = "Proportion") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  
    
    print(p)
  }
}

plot_categorical_distribution(selected_data) 
```
# Merge Variable
```{r merge variable}
selected_data <- selected_data %>% 
  mutate(SmokeStatus = recode_factor(SmokeStatus, "2" = "1", "3" = "1","4" = "2","5" = "3"))
```

# Outliers and New Variables
```{r variable creation and outliers}
normal_ranges <- list(Systolic = c(90, 180),
                      Diastolic = c(60, 120),
                      BMI = c(16, 47.5), 
                      PotassiumSodiumRatio = c(0, 4), 
                      AlcoholPercentage = c(0, 85), 
                      CarbonhydrateEnergy = c(0, 100), 
                      FiberEnergy = c(0, 100))

selected_data <- selected_data %>%
  filter(
    (is.na(Systolic) | (Systolic >= normal_ranges$Systolic[1] & Systolic <= normal_ranges$Systolic[2])),
    (is.na(Diastolic) | (Diastolic >= normal_ranges$Diastolic[1] & Diastolic <= normal_ranges$Diastolic[2])),
    (is.na(BMI) | (BMI >= normal_ranges$BMI[1] & BMI <= normal_ranges$BMI[2])),
    (is.na(AlcoholPercentage) | (AlcoholPercentage >= normal_ranges$AlcoholPercentage[1] & AlcoholPercentage <= normal_ranges$AlcoholPercentage[2])),
    (is.na(CarbonhydrateEnergy) | (CarbonhydrateEnergy >= normal_ranges$CarbonhydrateEnergy[1] & CarbonhydrateEnergy <= normal_ranges$CarbonhydrateEnergy[2])),
    (is.na(FiberEnergy) | (FiberEnergy >= normal_ranges$FiberEnergy[1] & FiberEnergy <= normal_ranges$FiberEnergy[2])),
    (is.na(PotassiumSodiumRatio) | (PotassiumSodiumRatio >= normal_ranges$PotassiumSodiumRatio[1] & PotassiumSodiumRatio <= normal_ranges$PotassiumSodiumRatio[2]))
  ) %>%
  mutate(
    Hypertension = as.factor(case_when(
      Systolic >= 120 | Diastolic >= 80 ~ 1,
      TRUE ~ 0
    ))
  )
 # select(-Sodium, -Potassium)
```

# Missing Values
```{r missing value, warning = FALSE}
selected_data %>% 
  select(-PersonID, -HouseID) %>% 
  miss_var_summary() %>% 
  gt() %>%
  tab_header(title = "Missing Value") %>%
  tab_caption(caption = md("Table 4: Missing Values in Selected Data")) %>%
  cols_label(variable = "Variable Name", n_miss = "Missing Count", pct_miss = "Percentage Missing") %>%
  fmt_number(columns = everything(), decimals = 2)
```

```{r heat map}
corr_matrix <- selected_data%>%
  select(where(is.numeric)) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  cor(use = "complete.obs")

ggcorrplot(corr_matrix, lab = TRUE, 
           title = "Correlation Heatmap for ALL", 
           colors = c("blue", "white", "red"),
           tl.cex = 9, lab_size = 4) + theme(legend.position = "none")
```

# Regression Imputation
```{r regression imputation, fig.height = 5, fig.width= 11}
selected_data$BMI[is.na(selected_data$BMI)] <- lm(BMI ~ Gender + SocialStatus + Age + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$BMI), ])

selected_data$Systolic[is.na(selected_data$Systolic)] <- lm(Systolic ~ Gender + SocialStatus + Age + BMI + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>% 
  predict(newdata = selected_data[is.na(selected_data$Systolic), ])

selected_data$Diastolic[is.na(selected_data$Diastolic)] <- lm(Diastolic ~ Gender + SocialStatus + Age + BMI + Systolic + PotassiumSodiumRatio + FiberEnergy + CarbonhydrateEnergy + AlcoholPercentage + SmokeStatus + Identity, data = selected_data, na.action = na.exclude) %>%
  predict(newdata = selected_data[is.na(selected_data$Diastolic), ])

vis_miss(selected_data %>% 
           select(-PersonID, -HouseID))
```
# Normalization
```{r normalization}
#numeric_cols <- sapply(selected_data, is.numeric)  
#normalized_data <- selected_data
#normalized_data[numeric_cols] <- scale(selected_data[numeric_cols])



numeric_data <- selected_data %>% select_if(is.numeric)

preprocess_params <- preProcess(numeric_data, method = c("center", "scale"))

scaled_numeric_data <- predict(preprocess_params, numeric_data)

normalized_data <- bind_cols(scaled_numeric_data, selected_data %>% select_if(Negate(is.numeric)))
```
# Multicollinearity
```{r multicollinearity, fig.height=10, fig.width=10}
selected_data %>% 
  select(where(is.numeric), -Systolic, -Diastolic) %>% 
  ggscatmat()
```

# Final check
```{r}
selected_predictors <- selected_data %>%
  select(FiberEnergy, CarbonhydrateEnergy)

# Create the table using gt()
table <- data.frame(Variable = names(selected_predictors), 
                    Type = sapply(selected_predictors, class)) %>%
  gt() %>%
  tab_header(
    title = "Variable Names and Their Types"
  ) %>%
  tab_caption(
    caption = md("**Table 2**: Predictor variables and their data types")
  )

# Print the table
print(table)

# Select all variables except 'FiberEnergy' and 'CarbonhydrateEnergy'
confounding_vars <- selected_data %>%
  select(-FiberEnergy, -CarbonhydrateEnergy,-PersonID, -HouseID)

# Create the table using gt()
confounding_table <- data.frame(Variable = names(confounding_vars), 
                                Type = sapply(confounding_vars, class)) %>%
  gt() %>%
  tab_header(
    title = "Confounding Variables and Their Types"
  ) %>%
  tab_caption(
    caption = md("**Table 3**: Confounding variables and their data types")
  )

# Print the table
print(confounding_table)
```


```{r}
variable_data <- data.frame(
  Variable = c("PersonID", "HouseID", "SocialStatus", "Age", "Gender", "BMI", 
               "Systolic", "Diastolic", "SmokeStatus", "AlcoholPercentage", 
               "PotassiumSodiumRatio", "FiberEnergy", "CarbohydrateEnergy", 
               "Identity"),
  Description = c("Selected person identifier",
                  "Household identifier",
                  "ICD10 Conditions - Mini classification",
                  "Age of person",
                  "Sex of person",
                  "Body Mass Index (BMI) - score measured",
                  "Systolic Blood Pressure (mmHg)",
                  "Diastolic Blood Pressure (mmHg)",
                  "Smoker Status",
                  "Average of Percentage of energy from alcohol (Day1) and Percentage of energy from alcohol (Day2)",
                  "The ratio of the average of Potassium (supplement) Day 1 mg and Potassium (supplement) Day 2 mg over the average of Sodium (total) Day 1 mg and Sodium (total) Day 2 mg",
                  "Average of Percentage of energy from fibre (Day1) and Percentage of energy from fibre (Day2)",
                  "Average of Percentage of energy from carbohydrate (Day1) and Percentage of energy from carbohydrate (Day2)",
                  "Indicates whether the individual is part of the indigenous population"))

variable_table <- variable_data %>%
  gt() %>%
  tab_header(
    title = "All related variables"
  ) %>%
  cols_label(
    Variable = "Variable",
    Description = "Description"
  ) %>%
  tab_caption(
    caption = md("**Table 1:** Variable Descriptions")
  )%>%
  fmt_markdown(columns = everything()) %>%
  tab_options(table.width = pct(100))

variable_table
```


# Correlation

## Scatter plot

```{r message = FALSE}
# First row: FiberEnergy vs Systolic and Diastolic
fiber_syst_plot <- ggplot(selected_data, aes(x = FiberEnergy, y = Systolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.6) +  # Orange scatter points with transparency
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +  # Blue regression line
  labs(title = "Fiber Energy vs Systolic BP",
       x = "Fiber Energy",
       y = "Systolic BP") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

fiber_dias_plot <- ggplot(selected_data, aes(x = FiberEnergy, y = Diastolic)) +
  geom_point(color = "#FF7F0E", alpha = 0.6) +  # Orange scatter points with transparency
  geom_smooth(method = "lm", se = FALSE, color = "#1F77B4", size = 1.5) +  # Blue regression line
  labs(title = "Fiber Energy vs Diastolic BP",
       x = "Fiber Energy",
       y = "Diastolic BP") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

# Second row: CarbohydrateEnergy vs Systolic and Diastolic
carbo_syst_plot <- ggplot(selected_data, aes(x = CarbonhydrateEnergy, y = Systolic)) +
  geom_point(color = "#2CA02C", alpha = 0.6) +  # Green scatter points with transparency
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +  # Red regression line
  labs(title = "Carbohydrate Energy vs Systolic BP",
       x = "Carbohydrate Energy",
       y = "Systolic BP") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

carbo_dias_plot <- ggplot(selected_data, aes(x = CarbonhydrateEnergy, y = Diastolic)) +
  geom_point(color = "#2CA02C", alpha = 0.6) +  # Green scatter points with transparency
  geom_smooth(method = "lm", se = FALSE, color = "#D62728", size = 1.5) +  # Red regression line
  labs(title = "Carbohydrate Energy vs Diastolic BP",
       x = "Carbohydrate Energy",
       y = "Diastolic BP") +
  theme_light(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

# Combine the plots into a 2x2 grid
grid.arrange(fiber_syst_plot, fiber_dias_plot, carbo_syst_plot, carbo_dias_plot, nrow = 2)
```

## linear correlation table


```{r}
# Compute Pearson and Spearman correlations for FiberEnergy and CarbohydrateEnergy vs Systolic and Diastolic BP
fiber_syst_pearson <- cor(selected_data$FiberEnergy, selected_data$Systolic, method = "pearson")
fiber_syst_spearman <- cor(selected_data$FiberEnergy, selected_data$Systolic, method = "spearman")
fiber_dias_pearson <- cor(selected_data$FiberEnergy, selected_data$Diastolic, method = "pearson")
fiber_dias_spearman <- cor(selected_data$FiberEnergy, selected_data$Diastolic, method = "spearman")

carbo_syst_pearson <- cor(selected_data$CarbonhydrateEnergy, selected_data$Systolic, method = "pearson")
carbo_syst_spearman <- cor(selected_data$CarbonhydrateEnergy, selected_data$Systolic, method = "spearman")
carbo_dias_pearson <- cor(selected_data$CarbonhydrateEnergy, selected_data$Diastolic, method = "pearson")
carbo_dias_spearman <- cor(selected_data$CarbonhydrateEnergy, selected_data$Diastolic, method = "spearman")

# Create a data frame to store the correlation results
correlation_table <- data.frame(
  Measure = c("FiberEnergy vs Systolic BP", "FiberEnergy vs Diastolic BP",
              "CarbohydrateEnergy vs Systolic BP", "CarbohydrateEnergy vs Diastolic BP"),
  Pearson_Correlation = c(fiber_syst_pearson, fiber_dias_pearson, carbo_syst_pearson, carbo_dias_pearson),
  Spearman_Correlation = c(fiber_syst_spearman, fiber_dias_spearman, carbo_syst_spearman, carbo_dias_spearman)
)

# Use gt to create a formatted table
correlation_table %>%
  gt() %>%
  tab_header(
    title = "Table: Correlation between Nutrient Energy Intake and Blood Pressure",
    subtitle = "Correlation values for both Systolic and Diastolic Blood Pressure"
  ) %>%
  fmt_number(
    columns = c(Pearson_Correlation, Spearman_Correlation), decimals = 2
  )


```




## Non-linear correlation table


```{r}
# Load necessary packages
library(minerva)

fiber_syst_kendall <- cor(data$FiberEnergy, data$Systolic, method = "kendall")
fiber_syst_mic <- mine(data$FiberEnergy, data$Systolic)$MIC

fiber_dias_kendall <- cor(data$FiberEnergy, data$Diastolic, method = "kendall")
fiber_dias_mic <- mine(data$FiberEnergy, data$Diastolic)$MIC

carbo_syst_kendall <- cor(data$CarbonhydrateEnergy, data$Systolic, method = "kendall")
carbo_syst_mic <- mine(data$CarbonhydrateEnergy, data$Systolic)$MIC

carbo_dias_kendall <- cor(data$CarbonhydrateEnergy, data$Diastolic, method = "kendall")
carbo_dias_mic <- mine(data$CarbonhydrateEnergy, data$Diastolic)$MIC

# Create a data frame to store the correlation results
correlation_table_mic_kendall <- data.frame(
  Measure = c("FiberEnergy vs Systolic BP", "FiberEnergy vs Diastolic BP",
              "CarbohydrateEnergy vs Systolic BP", "CarbohydrateEnergy vs Diastolic BP"),
  Kendall_Correlation = c(fiber_syst_kendall, fiber_dias_kendall, carbo_syst_kendall, carbo_dias_kendall),
  MIC_Correlation = c(fiber_syst_mic, fiber_dias_mic, carbo_syst_mic, carbo_dias_mic)
)

# Create the table using gt
correlation_table_mic_kendall %>%
  gt() %>%
  tab_header(
    title = "Table: Kendall's Tau and MIC Correlation between Nutrient Energy Intake and Blood Pressure"
  ) %>%
  fmt_number(
    columns = c(Kendall_Correlation, MIC_Correlation), decimals = 3
  )

```




# Model comparison

Lasso regression
Multiple Linear regression
GAM
Polynomial regression

```{r}
# Load necessary packages
library(mgcv)  # For GAM
library(MASS)  # For Robust Regression
library(caret)  # For evaluation metrics

# Set seed for reproducibility
set.seed(123)

# Number of folds and repetitions
k <- 5  # Number of folds
repeats <- 3  # Number of repetitions

# Create empty lists to store results for GAM, MLR, and Robust Regression
gam_rmse_values <- numeric()
gam_mae_values <- numeric()
gam_r2_values <- numeric()

mlr_rmse_values <- numeric()
mlr_mae_values <- numeric()
mlr_r2_values <- numeric()

robust_rmse_values <- numeric()
robust_mae_values <- numeric()
robust_r2_values <- numeric()

# Define the number of rows
n <- nrow(normalized_data)

# Perform k-fold cross-validation for `repeats` times
for (r in 1:repeats) {
  # Create random folds
  folds <- sample(rep(1:k, length.out = n))
  
  # Loop through each fold
  for (i in 1:k) {
    # Split the data into training and testing
    train_idx <- which(folds != i)  # All but the ith fold
    test_idx <- which(folds == i)   # Only the ith fold
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Systolic
    
    # 1. Fit a standard GAM model on the training data
    gam_model <- gam(Systolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FiberEnergy) + s(CarbonhydrateEnergy) + 
                       Identity + s(PotassiumSodiumRatio),
                     data = data_train)
    
    # Predict on the test data (GAM)
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    # Evaluate performance for GAM (RMSE, MAE, R2)
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))  # RMSE
    gam_mae <- mean(abs(gam_predictions - actual))  # MAE
    gam_r2 <- cor(gam_predictions, actual)^2  # R2
    
    # Store the GAM results
    gam_rmse_values <- c(gam_rmse_values, gam_rmse)
    gam_mae_values <- c(gam_mae_values, gam_mae)
    gam_r2_values <- c(gam_r2_values, gam_r2)
    
    # 2. Fit a Multiple Linear Regression (MLR) model on the training data
    mlr_model <- lm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + 
                      Identity + PotassiumSodiumRatio, data = data_train)
    
    # Predict on the test data (MLR)
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    # Evaluate performance for MLR (RMSE, MAE, R2)
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))  # RMSE
    mlr_mae <- mean(abs(mlr_predictions - actual))  # MAE
    mlr_r2 <- cor(mlr_predictions, actual)^2  # R2
    
    # Store the MLR results
    mlr_rmse_values <- c(mlr_rmse_values, mlr_rmse)
    mlr_mae_values <- c(mlr_mae_values, mlr_mae)
    mlr_r2_values <- c(mlr_r2_values, mlr_r2)
    
    # 3. Fit a Robust Linear Regression model on the training data using `rlm()` (Robust Regression)
    robust_model <- rlm(Systolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                          AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + 
                          Identity + PotassiumSodiumRatio, data = data_train)
    
    # Predict on the test data (Robust Regression)
    robust_predictions <- predict(robust_model, newdata = data_test)
    
    # Evaluate performance for Robust Regression (RMSE, MAE, R2)
    robust_rmse <- sqrt(mean((robust_predictions - actual)^2))  # RMSE
    robust_mae <- mean(abs(robust_predictions - actual))  # MAE
    robust_r2 <- cor(robust_predictions, actual)^2  # R2
    
    # Store the Robust Regression results
    robust_rmse_values <- c(robust_rmse_values, robust_rmse)
    robust_mae_values <- c(robust_mae_values, robust_mae)
    robust_r2_values <- c(robust_r2_values, robust_r2)
  }
}

# Calculate the average performance across all folds and repetitions for all models
# GAM performance
mean_gam_rmse <- mean(gam_rmse_values)
mean_gam_mae <- mean(gam_mae_values)
mean_gam_r2 <- mean(gam_r2_values)

# Multiple Linear Regression performance
mean_mlr_rmse <- mean(mlr_rmse_values)
mean_mlr_mae <- mean(mlr_mae_values)
mean_mlr_r2 <- mean(mlr_r2_values)

# Robust Regression performance
mean_robust_rmse <- mean(robust_rmse_values)
mean_robust_mae <- mean(robust_mae_values)
mean_robust_r2 <- mean(robust_r2_values)

# Print the results
cat("GAM Performance:\n")
cat("RMSE:", mean_gam_rmse, "\nMAE:", mean_gam_mae, "\nR²:", mean_gam_r2, "\n\n")

cat("Multiple Linear Regression Performance:\n")
cat("RMSE:", mean_mlr_rmse, "\nMAE:", mean_mlr_mae, "\nR²:", mean_mlr_r2, "\n\n")

cat("Robust Regression Performance:\n")
cat("RMSE:", mean_robust_rmse, "\nMAE:", mean_robust_mae, "\nR²:", mean_robust_r2, "\n")

```


```{r fig.width=12}
# Load necessary packages
library(plotly)
library(dplyr)

# Create a data frame with the results for each model
results_df <- data.frame(
    Model = c("GAM", "Polynomial Regression", "Multiple Linear Regression", "Lasso Regression"),
    RMSE = c(mean_gam_rmse, mean_poly_rmse, mean_mlr_rmse, mean_lasso_rmse),
    MAE = c(mean_gam_mae, mean_poly_mae, mean_mlr_mae, mean_lasso_mae),
    R2 = c(mean_gam_r2, mean_poly_r2, mean_mlr_r2, mean_lasso_r2)
)

# Define custom colors for the bar plots
colors <- c('#70d6ff', '#ff70a6', '#ffd670', '#FF9770')

# Manually set Y-axis ranges for each plot
rmse_range <- c(min(results_df$RMSE) * 0.95, max(results_df$RMSE) * 1.05)
mae_range <- c(min(results_df$MAE) * 0.95, max(results_df$MAE) * 1.05)
r2_range <- c(min(results_df$R2) * 0.95, max(results_df$R2) * 1.05)

# Create individual plots for RMSE, MAE, and R2

# RMSE plot
fig_rmse <- plot_ly(results_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "RMSE Comparison",
    yaxis = list(title = 'RMSE', range = rmse_range),  # Manually adjust RMSE range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# MAE plot
fig_mae <- plot_ly(results_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "MAE Comparison",
    yaxis = list(title = 'MAE', range = mae_range),  # Manually adjust MAE range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# R2 plot
fig_r2 <- plot_ly(results_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "R² Comparison",
    yaxis = list(title = 'R²', range = r2_range),  # Manually adjust R² range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# Arrange the three plots into a multi-panel layout
fig <- subplot(fig_rmse, fig_mae, fig_r2, nrows = 1, shareY = FALSE, titleX = TRUE)

# Display the multi-panel plot
fig

```


```{r}
# Load necessary packages
library(mgcv)  # For GAM
library(caret)  # For evaluation metrics
library(glmnet)  # For Lasso regression

# Set seed for reproducibility
set.seed(123)

# Number of folds and repetitions
k <- 5  # Number of folds
repeats <- 3  # Number of repetitions

# Create empty lists to store results for GAM, polynomial regression, multiple linear regression, and Lasso regression
gam_rmse_values <- numeric()
gam_mae_values <- numeric()
gam_r2_values <- numeric()

poly_rmse_values <- numeric()
poly_mae_values <- numeric()
poly_r2_values <- numeric()

mlr_rmse_values <- numeric()
mlr_mae_values <- numeric()
mlr_r2_values <- numeric()

lasso_rmse_values <- numeric()
lasso_mae_values <- numeric()
lasso_r2_values <- numeric()

# Define the number of rows
n <- nrow(normalized_data)

# Perform k-fold cross-validation for `repeats` times
for (r in 1:repeats) {
  # Create random folds
  folds <- sample(rep(1:k, length.out = n))
  
  # Loop through each fold
  for (i in 1:k) {
    # Split the data into training and testing
    train_idx <- which(folds != i)  # All but the ith fold
    test_idx <- which(folds == i)   # Only the ith fold
    
    data_train <- normalized_data[train_idx, ]
    data_test <- normalized_data[test_idx, ]
    
    actual <- data_test$Diastolic
    
    # 1. Fit a GAM model on the training data
    gam_model <- gam(Diastolic ~ s(Age) + s(BMI) + SocialStatus + Gender + SmokeStatus +
                       s(AlcoholPercentage) + s(FiberEnergy) + s(CarbonhydrateEnergy) + 
                       Identity + s(PotassiumSodiumRatio),
                     data = data_train)
    
    # Predict on the test data (GAM)
    gam_predictions <- predict(gam_model, newdata = data_test)
    
    # Evaluate performance for GAM (RMSE, MAE, R2)
    gam_rmse <- sqrt(mean((gam_predictions - actual)^2))  # RMSE
    gam_mae <- mean(abs(gam_predictions - actual))  # MAE
    gam_r2 <- cor(gam_predictions, actual)^2  # R2
    
    # Store the GAM results
    gam_rmse_values <- c(gam_rmse_values, gam_rmse)
    gam_mae_values <- c(gam_mae_values, gam_mae)
    gam_r2_values <- c(gam_r2_values, gam_r2)
    
    # 2. Fit a Polynomial Regression model on the training data
    poly_model <- lm(Diastolic ~ poly(Age, 2) + poly(BMI, 2) + SocialStatus + Gender + SmokeStatus +
                       poly(AlcoholPercentage, 2) + poly(FiberEnergy, 2) + poly(CarbonhydrateEnergy, 2) + 
                       Identity + poly(PotassiumSodiumRatio, 2),
                     data = data_train)
    
    # Predict on the test data (Polynomial Regression)
    poly_predictions <- predict(poly_model, newdata = data_test)
    
    # Evaluate performance for Polynomial Regression (RMSE, MAE, R2)
    poly_rmse <- sqrt(mean((poly_predictions - actual)^2))  # RMSE
    poly_mae <- mean(abs(poly_predictions - actual))  # MAE
    poly_r2 <- cor(poly_predictions, actual)^2  # R2
    
    # Store the Polynomial Regression results
    poly_rmse_values <- c(poly_rmse_values, poly_rmse)
    poly_mae_values <- c(poly_mae_values, poly_mae)
    poly_r2_values <- c(poly_r2_values, poly_r2)
    
    # 3. Fit a Multiple Linear Regression (MLR) model on the training data
    mlr_model <- lm(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                      AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + 
                      Identity + PotassiumSodiumRatio, data = data_train)
    
    # Predict on the test data (Multiple Linear Regression)
    mlr_predictions <- predict(mlr_model, newdata = data_test)
    
    # Evaluate performance for MLR (RMSE, MAE, R2)
    mlr_rmse <- sqrt(mean((mlr_predictions - actual)^2))  # RMSE
    mlr_mae <- mean(abs(mlr_predictions - actual))  # MAE
    mlr_r2 <- cor(mlr_predictions, actual)^2  # R2
    
    # Store the MLR results
    mlr_rmse_values <- c(mlr_rmse_values, mlr_rmse)
    mlr_mae_values <- c(mlr_mae_values, mlr_mae)
    mlr_r2_values <- c(mlr_r2_values, mlr_r2)
    
    # 4. Fit a Lasso Regression model on the training data
    # Prepare the data for glmnet (Lasso requires matrix input)
    x_train <- model.matrix(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                              AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + 
                              Identity + PotassiumSodiumRatio, data = data_train)[, -1]
    y_train <- data_train$Diastolic
    
    x_test <- model.matrix(Diastolic ~ Age + BMI + SocialStatus + Gender + SmokeStatus +
                             AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + 
                             Identity + PotassiumSodiumRatio, data = data_test)[, -1]
    
    # Fit the Lasso model with cross-validated lambda
    lasso_model <- cv.glmnet(x_train, y_train, alpha = 1)
    
    # Predict on the test data (Lasso Regression)
    lasso_predictions <- predict(lasso_model, s = lasso_model$lambda.min, newx = x_test)
    
    # Evaluate performance for Lasso (RMSE, MAE, R2)
    lasso_rmse <- sqrt(mean((lasso_predictions - actual)^2))  # RMSE
    lasso_mae <- mean(abs(lasso_predictions - actual))  # MAE
    lasso_r2 <- cor(lasso_predictions, actual)^2  # R2
    
    # Store the Lasso results
    lasso_rmse_values <- c(lasso_rmse_values, lasso_rmse)
    lasso_mae_values <- c(lasso_mae_values, lasso_mae)
    lasso_r2_values <- c(lasso_r2_values, lasso_r2)
  }
}

# Calculate the average performance across all folds and repetitions for all models
# GAM performance
mean_gam_rmse <- mean(gam_rmse_values)
mean_gam_mae <- mean(gam_mae_values)
mean_gam_r2 <- mean(gam_r2_values)

# Polynomial Regression performance
mean_poly_rmse <- mean(poly_rmse_values)
mean_poly_mae <- mean(poly_mae_values)
mean_poly_r2 <- mean(poly_r2_values)

# Multiple Linear Regression performance
mean_mlr_rmse <- mean(mlr_rmse_values)
mean_mlr_mae <- mean(mlr_mae_values)
mean_mlr_r2 <- mean(mlr_r2_values)

# Lasso Regression performance
mean_lasso_rmse <- mean(lasso_rmse_values)
mean_lasso_mae <- mean(lasso_mae_values)
mean_lasso_r2 <- mean(lasso_r2_values)
```


```{r fig.width=12}
# Load necessary packages
library(plotly)
library(dplyr)

# Create a data frame with the results for each model
results_df <- data.frame(
    Model = c("GAM", "Polynomial Regression", "Multiple Linear Regression", "Lasso Regression"),
    RMSE = c(mean_gam_rmse, mean_poly_rmse, mean_mlr_rmse, mean_lasso_rmse),
    MAE = c(mean_gam_mae, mean_poly_mae, mean_mlr_mae, mean_lasso_mae),
    R2 = c(mean_gam_r2, mean_poly_r2, mean_mlr_r2, mean_lasso_r2)
)

# Define custom colors for the bar plots
colors <- c('#70d6ff', '#ff70a6', '#ffd670', '#FF9770')

# Manually set Y-axis ranges for each plot
rmse_range <- c(min(results_df$RMSE) * 0.95, max(results_df$RMSE) * 1.05)
mae_range <- c(min(results_df$MAE) * 0.95, max(results_df$MAE) * 1.05)
r2_range <- c(min(results_df$R2) * 0.95, max(results_df$R2) * 1.05)

# Create individual plots for RMSE, MAE, and R2

# RMSE plot
fig_rmse <- plot_ly(results_df, x = ~Model, y = ~round(RMSE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(RMSE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "RMSE Comparison",
    yaxis = list(title = 'RMSE', range = rmse_range),  # Manually adjust RMSE range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# MAE plot
fig_mae <- plot_ly(results_df, x = ~Model, y = ~round(MAE, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(MAE, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "MAE Comparison",
    yaxis = list(title = 'MAE', range = mae_range),  # Manually adjust MAE range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# R2 plot
fig_r2 <- plot_ly(results_df, x = ~Model, y = ~round(R2, 3), type = 'bar', color = ~Model, colors = colors,
               text = ~round(R2, 3), textposition = 'outside', textfont = list(color = 'black')) %>%
  layout(
    title = "R² Comparison",
    yaxis = list(title = 'R²', range = r2_range),  # Manually adjust R² range
    xaxis = list(title = ''),
    showlegend = FALSE,  # Remove the legend
    paper_bgcolor = 'rgba(245, 246, 249, 1)',  # Background color
    plot_bgcolor = 'rgba(0, 0, 0, 0)'  # Transparent plot background
  )

# Arrange the three plots into a multi-panel layout
fig <- subplot(fig_rmse, fig_mae, fig_r2, nrows = 1, shareY = FALSE, titleX = TRUE)

# Display the multi-panel plot
fig

```

```{r}
library(MASS)
robust <- rlm (Systolic ~ SocialStatus + Gender + Age + BMI + SmokeStatus + AlcoholPercentage + FiberEnergy + CarbonhydrateEnergy + Identity + PotassiumSodiumRatio, data = normalized_data)
summary(robust)
```



## Systolic






## Diastolic
















